---
title: "CCP_generation_Modul_6_LSA.qmd"
author: "izb."
format: html
editor: visual
execute:
  echo: true
  warning: true
  message: true
---

# Modul 6 - 1

```{r}
# MODUL 6 — LSA Semantic Similarity

#This module builds a fixed 50-dimensional LSA reference space from the L1-Corpus and later projects sentences, paragraphs, titles, and subheadings from the four target texts into that space.

# MODUL 6.1 — Setup


required_packages <- c(
  "dplyr",
  "tidyr",
  "purrr",
  "readr",
  "stringr",
  "tibble",
  "quanteda",
  "Matrix",
  "irlba"
)

missing_packages <- required_packages[
  !vapply(
    required_packages,
    requireNamespace,
    logical(1),
    quietly = TRUE
  )
]

if (length(missing_packages) > 0) {
  stop(
    "Missing packages: ",
    paste(missing_packages, collapse = ", "),
    "\nInstall them before continuing."
  )
}

library(dplyr)
library(tidyr)
library(purrr)
library(readr)
library(stringr)
library(tibble)
library(quanteda)
library(Matrix)
library(irlba)

options(
  dplyr.summarise.inform = FALSE,
  scipen = 999
)

lsa_dimensions <- 50L

# Primary MODUL 6 input categories:
# content words plus coordinating and subordinating connectors
lsa_pos_keep <- c(
  "NOUN",
  "PROPN",
  "VERB",
  "ADJ",
  "ADV",
  "CCONJ",
  "SCONJ"
)

message("MODUL 6 setup complete.")
message("Fixed LSA dimensions: ", lsa_dimensions)
message(
  "Included UPOS categories: ",
  paste(lsa_pos_keep, collapse = ", ")
)
```

# Modul 6.2 — Paths and output folders

```{r}

# MODUL 6.2 — Paths and output folders

# This QMD is located inside:
# CC_Profile_HvsLM/CCP_generation_pipeL/

data_prep_dir <- "../data_prep_CCP_output"
module_6_output_dir <- "../measure_outputs"

dir.create(
  module_6_output_dir,
  showWarnings = FALSE,
  recursive = TRUE
)

if (!dir.exists(data_prep_dir)) {
  stop(
    "Missing frozen input folder: ",
    data_prep_dir
  )
}

message(
  "Working directory: ",
  getwd()
)

message(
  "Frozen input folder: ",
  normalizePath(
    data_prep_dir,
    mustWork = TRUE
  )
)

message(
  "MODUL 6 output folder: ",
  normalizePath(
    module_6_output_dir,
    mustWork = TRUE
  )
)
```

# Modul 6.4 — Core input paths and validation

```{r}
# MODUL 6.4 — Core input paths and validation

core_input_paths <- c(
  current_texts_parsed = file.path(data_prep_dir, "current_texts_parsed.rds"),
  parsed_S_layer = file.path(data_prep_dir, "parsed_S_layer.rds"),
  current_paragraph_layer = file.path(data_prep_dir, "current_paragraph_layer.rds"),
  l1_reference_bundle = file.path(data_prep_dir, "l1_corpus_reference_bundle_corrected.rds")
)

missing_core_inputs <- names(core_input_paths)[!file.exists(core_input_paths)]

if (length(missing_core_inputs) > 0L) {
  stop(
    "Missing required MODUL 6 inputs:\n",
    paste(core_input_paths[missing_core_inputs], collapse = "\n")
  )
}

message("All four core MODUL 6 inputs are available.")
```

# Modul 6.5 — Load frozen core inputs

```{r}
# MODUL 6.5 — Load frozen core inputs

current_texts_parsed <- readRDS(
  core_input_paths[["current_texts_parsed"]]
)

parsed_S_layer <- readRDS(
  core_input_paths[["parsed_S_layer"]]
)

current_paragraph_layer <- readRDS(
  core_input_paths[["current_paragraph_layer"]]
)

l1_reference_bundle <- readRDS(
  core_input_paths[["l1_reference_bundle"]]
)

l1_documents <- l1_reference_bundle$documents
l1_sentences <- l1_reference_bundle$sentences
l1_tokens <- l1_reference_bundle$tokens

stopifnot(
  isTRUE(
    l1_reference_bundle$metadata$final_audit_passed
  )
)

message("Frozen core inputs loaded.")
message("Validated corrected L1 reference bundle loaded.")
```

# Modul 6.7 — Title and subheading source diagnostic

```{r}

# MODUL 6.7 — Load and validate title metadata new

title_source_paths <- c(
  human_headings = "../metadata/text1_hm_headings.csv",
  reconstruction_prompt = "../prompts/prompt_master_reconstruction.txt"
)

missing_title_sources <- names(title_source_paths)[
  !file.exists(title_source_paths)
]

if (length(missing_title_sources) > 0) {
  stop(
    "Missing title source files: ",
    paste(missing_title_sources, collapse = ", ")
  )
}

human_headings <- readr::read_csv(
  title_source_paths[["human_headings"]],
  show_col_types = FALSE
)

prompt_master_reconstruction <- readr::read_file(
  title_source_paths[["reconstruction_prompt"]]
)

required_heading_columns <- c(
  "text_id",
  "heading_id",
  "heading_type",
  "heading_text"
)

missing_heading_columns <- setdiff(
  required_heading_columns,
  names(human_headings)
)

if (length(missing_heading_columns) > 0) {
  stop(
    "The human heading file is missing required columns: ",
    paste(missing_heading_columns, collapse = ", ")
  )
}

human_headings_clean <- human_headings |>
  mutate(
    heading_type = stringr::str_to_lower(
      stringr::str_squish(as.character(heading_type))
    ),
    heading_text = stringr::str_squish(
      as.character(heading_text)
    )
  ) |>
  filter(
    !is.na(heading_text),
    heading_text != ""
  )

main_titles <- human_headings_clean |>
  filter(heading_type == "main_title") |>
  distinct(heading_text) |>
  pull(heading_text)

if (length(main_titles) != 1) {
  stop(
    "Expected exactly one main title, but found: ",
    length(main_titles)
  )
}

common_main_title <- main_titles[[1]]

title_in_prompt <- stringr::str_detect(
  prompt_master_reconstruction,
  stringr::fixed(
    common_main_title,
    ignore_case = TRUE
  )
)

if (!title_in_prompt) {
  stop(
    "The validated main title was not found in the shared ",
    "reconstruction prompt."
  )
}

human_section_titles <- human_headings_clean |>
  filter(heading_type == "section_title") |>
  select(
    text_id,
    heading_id,
    heading_text
  )

message("Validated common main title: ", common_main_title)
message("Main title confirmed in the shared reconstruction prompt.")

message("Human section titles:")
print(
  human_section_titles,
  n = Inf,
  width = Inf
)
```

# Modul 6.8 — Basic target-layer integrity preview

```{r}

# MODUL 6.8 — Basic target-layer integrity preview

required_token_columns <- c(
  "text_id",
  "sentence_id",
  "lemma",
  "pos"
)

missing_token_columns <- setdiff(
  required_token_columns,
  names(current_texts_parsed)
)

if (length(missing_token_columns) > 0) {
  stop(
    "current_texts_parsed is missing required columns: ",
    paste(missing_token_columns, collapse = ", ")
  )
}

target_token_preview <- current_texts_parsed |>
  mutate(
    lemma_clean = str_to_lower(
      str_squish(as.character(lemma))
    ),
    pos = str_to_upper(as.character(pos)),
    included_in_lsa = (
      pos %in% lsa_pos_keep &
      !is.na(lemma_clean) &
      lemma_clean != ""
    ),
    lsa_token_class = case_when(
  pos %in% c("NOUN", "PROPN", "VERB", "ADJ", "ADV") ~ "content",
  pos %in% c("CCONJ", "SCONJ") ~ "connector",
  TRUE ~ "excluded"
    )
  )

target_token_preview |>
  count(
    text_id,
    lsa_token_class,
    name = "n_tokens"
  ) |>
  arrange(text_id, lsa_token_class) |>
  print(n = Inf, width = Inf)

message("Initial MODUL 6 input inspection complete.")

```

# Modul 6.11 — Build and validate the shared prompt-heading grid

```{r}

# MODUL 6.11 — Build and validate the shared prompt-heading grid

# 1. Validate the required preceding objects

required_heading_objects <- c(
  "common_main_title",
  "human_section_titles",
  "prompt_master_reconstruction"
)

missing_heading_objects <- required_heading_objects[
  !vapply(
    required_heading_objects,
    exists,
    logical(1),
    inherits = TRUE
  )
]

if (length(missing_heading_objects) > 0L) {
  stop(
    "Missing required heading objects: ",
    paste(missing_heading_objects, collapse = ", "),
    ". Run the preceding MODUL 6 cells first."
  )
}

# 2. Order and validate the three prescribed subheadings

ordered_section_titles <- human_section_titles |>
  mutate(
    source_heading_number = readr::parse_number(
      as.character(heading_id)
    )
  ) |>
  arrange(
    source_heading_number,
    heading_id
  ) |>
  distinct(
    heading_text,
    .keep_all = TRUE
  )

if (nrow(ordered_section_titles) != 3L) {
  stop(
    "Expected exactly three distinct prescribed subheadings, ",
    "but found: ",
    nrow(ordered_section_titles)
  )
}

if (any(is.na(ordered_section_titles$source_heading_number))) {
  stop(
    "At least one source heading ID has no usable numeric order."
  )
}

# 3. Build the fixed shared heading grid

shared_prompt_headings <- tibble::tibble(
  heading_order = 0:3,
  heading_id = c(
    "main_title",
    "subheading_1",
    "subheading_2",
    "subheading_3"
  ),
  heading_type = c(
    "main_title",
    rep("section_title", 3)
  ),
  heading_text = c(
    common_main_title,
    ordered_section_titles$heading_text
  )
)

# 4. Confirm that every heading occurs exactly once in the prompt

shared_prompt_headings <- shared_prompt_headings |>
  mutate(
    prompt_locations = purrr::map(
      heading_text,
      function(current_heading) {
        stringr::str_locate_all(
          prompt_master_reconstruction,
          stringr::fixed(
            current_heading,
            ignore_case = TRUE
          )
        )[[1]]
      }
    ),
    n_prompt_matches = purrr::map_int(
      prompt_locations,
      nrow
    ),
    prompt_start = purrr::map_int(
      prompt_locations,
      function(current_locations) {
        if (nrow(current_locations) == 1L) {
          return(
            as.integer(
              current_locations[1, "start"]
            )
          )
        }

        NA_integer_
      }
    )
  )

if (any(shared_prompt_headings$n_prompt_matches != 1L)) {
  invalid_prompt_headings <- shared_prompt_headings |>
    filter(
      n_prompt_matches != 1L
    )

  stop(
    "Each prescribed heading must occur exactly once in the ",
    "shared reconstruction prompt. Problematic headings: ",
    paste(
      invalid_prompt_headings$heading_text,
      collapse = " | "
    )
  )
}

# 5. Validate the prescribed heading order

heading_order_correct <- all(
  diff(shared_prompt_headings$prompt_start) > 0L
)

if (!heading_order_correct) {
  stop(
    "The main title and three prescribed subheadings do not ",
    "occur in the expected order in the reconstruction prompt."
  )
}

shared_prompt_headings <- shared_prompt_headings |>
  select(
    heading_order,
    heading_id,
    heading_type,
    heading_text,
    n_prompt_matches,
    prompt_start
  )

# 6. Display the validated shared heading grid

message(
  "Validated one shared main title and three prescribed ",
  "subheadings in the correct prompt order."
)

print(
  shared_prompt_headings,
  n = Inf,
  width = Inf
)
```

# Modul 6.12 — Prepare structural mapping inputs

```{r}

# MODUL 6.12 — Prepare heading and target-text inputs for structural mapping

# 1. Validate required upstream objects

required_objects <- c(
  "current_paragraph_layer",
  "shared_prompt_headings"
)

missing_objects <- required_objects[
  !vapply(
    required_objects,
    exists,
    logical(1),
    inherits = TRUE
  )
]

if (length(missing_objects) > 0L) {
  stop(
    "Missing required objects: ",
    paste(missing_objects, collapse = ", "),
    ". Run the preceding MODUL 6 cells first."
  )
}

required_paragraph_columns <- c(
  "text_id",
  "model",
  "paragraph_id",
  "paragraph_number",
  "paragraph_text"
)

missing_paragraph_columns <- setdiff(
  required_paragraph_columns,
  names(current_paragraph_layer)
)

if (length(missing_paragraph_columns) > 0L) {
  stop(
    "current_paragraph_layer is missing required columns: ",
    paste(missing_paragraph_columns, collapse = ", ")
  )
}

# 2. Prepare the three prescribed subheadings

prescribed_subheadings <- shared_prompt_headings |>
  filter(
    heading_id %in% paste0(
      "subheading_",
      1:3
    )
  ) |>
  transmute(
    section_id = as.integer(
      readr::parse_number(heading_id)
    ),
    subheading_id = heading_id,
    subheading_text = heading_text
  ) |>
  arrange(section_id)

if (
  nrow(prescribed_subheadings) != 3L ||
  !identical(
    prescribed_subheadings$section_id,
    1:3
  )
) {
  stop(
    "The shared heading grid must contain exactly ",
    "subheading_1, subheading_2, and subheading_3."
  )
}

# 3. Register and validate the four original target texts

target_text_sources <- tibble::tribble(
  ~text_id,                    ~file_path,
  "HUMAN_BASELINE_001", "../texte_input/text1_hm_annotated.txt",
  "BASIC_CHATGPT_NATIVE_001",  "../outputs_basic/BASIC_CHATGPT_NATIVE_001.txt",
  "BASIC_CLAUDE_NATIVE_001",   "../outputs_basic/BASIC_CLAUDE_NATIVE_001.txt",
  "BASIC_GEMINI_NATIVE_001",   "../outputs_basic/BASIC_GEMINI_NATIVE_001.txt"
)

if (
  !setequal(
    current_paragraph_layer$text_id,
    target_text_sources$text_id
  )
) {
  stop(
    "The paragraph-layer text IDs do not match ",
    "the four registered target texts."
  )
}

missing_target_files <- target_text_sources$file_path[
  !file.exists(target_text_sources$file_path)
]

if (length(missing_target_files) > 0L) {
  stop(
    "Missing original target-text files: ",
    paste(missing_target_files, collapse = ", ")
  )
}

# 4. Define structural normalization and location helpers

normalize_structure_text <- function(text_vector) {
  text_vector |>
    as.character() |>
    stringr::str_replace_all(
      "[\u2010\u2011\u2012\u2013\u2014\u2212]",
      "-"
    ) |>
    stringr::str_to_lower() |>
    stringr::str_replace_all(
      "[^\\p{L}\\p{N}]+",
      " "
    ) |>
    stringr::str_squish()
}

locate_once <- function(text, pattern) {
  locations <- stringr::str_locate_all(
    text,
    stringr::fixed(pattern)
  )[[1]]

  n_matches <- nrow(locations)

  tibble::tibble(
    n_matches = as.integer(n_matches),
    match_start = if (n_matches == 1L) {
      as.integer(locations[1, "start"])
    } else {
      NA_integer_
    },
    match_end = if (n_matches == 1L) {
      as.integer(locations[1, "end"])
    } else {
      NA_integer_
    }
  )
}

# 5. Load and normalize the structural reference texts

target_text_sources <- target_text_sources |>
  mutate(
    raw_text = purrr::map_chr(
      file_path,
      readr::read_file
    ),
    raw_text_normalized =
      normalize_structure_text(raw_text)
  )

prescribed_subheadings <- prescribed_subheadings |>
  mutate(
    subheading_normalized =
      normalize_structure_text(
        subheading_text
      )
  )

# 6. Display the prepared structural inputs

message(
  "Structural mapping inputs prepared for four texts ",
  "and three prescribed subheadings."
)

print(
  target_text_sources |>
    transmute(
      text_id,
      file_path,
      file_exists = file.exists(file_path),
      normalized_character_count =
        stringr::str_length(
          raw_text_normalized
        )
    ),
  n = Inf,
  width = Inf
)

print(
  prescribed_subheadings,
  n = Inf,
  width = Inf
)
```

# Modul 6.13 — Locate and audit prescribed subheadings

```{r}

# MODUL 6.13 — Locate the three prescribed subheadings in each target text

# 1. Locate each normalized subheading in each normalized target text

heading_locations <- tidyr::crossing(
  target_text_sources |>
    select(
      text_id,
      raw_text_normalized
    ),
  prescribed_subheadings
) |>
  mutate(
    location_result = purrr::map2(
      raw_text_normalized,
      subheading_normalized,
      locate_once
    )
  ) |>
  tidyr::unnest(location_result) |>
  rename(
    n_heading_matches = n_matches,
    heading_start = match_start,
    heading_end = match_end
  ) |>
  group_by(text_id) |>
  arrange(
    section_id,
    .by_group = TRUE
  ) |>
  mutate(
    next_heading_start = lead(
      heading_start
    )
  ) |>
  ungroup()

# 2. Build the heading-level structural audit

heading_audit <- heading_locations |>
  group_by(text_id) |>
  summarise(
    n_detected_sections = sum(
      n_heading_matches == 1L
    ),
    n_missing_sections = sum(
      n_heading_matches == 0L
    ),
    n_duplicated_sections = sum(
      n_heading_matches > 1L
    ),
    all_sections_present = all(
      n_heading_matches == 1L
    ),
    .groups = "drop"
  ) |>
  left_join(
    heading_locations |>
      group_by(text_id) |>
      summarise(
        section_order_correct = if (
          all(n_heading_matches == 1L)
        ) {
          all(diff(heading_start) > 0L)
        } else {
          FALSE
        },
        .groups = "drop"
      ),
    by = "text_id"
  )

# 3. Display the heading locations and audit

message("Detected prescribed subheadings:")

print(
  heading_locations |>
    select(
      text_id,
      section_id,
      subheading_id,
      n_heading_matches,
      heading_start,
      heading_end,
      next_heading_start
    ),
  n = Inf,
  width = Inf
)

message("Heading audit:")

print(
  heading_audit,
  n = Inf,
  width = Inf
)

# 4. Warn without silently repairing structural anomalies

if (
  any(
    !heading_audit$all_sections_present |
      !heading_audit$section_order_correct
  )
) {
  warning(
    "At least one target text has a missing, duplicated, ",
    "or misordered prescribed subheading."
  )
} else {
  message(
    "All four texts contain the three prescribed ",
    "subheadings exactly once and in the correct order."
  )
}
```

# Modul 6.14 — Locate and audit body paragraphs

```{r}

# MODUL 6.14 — Locate each frozen body paragraph in its original target text

# 1. Validate the required upstream objects

required_objects <- c(
  "current_paragraph_layer",
  "target_text_sources",
  "normalize_structure_text",
  "locate_once"
)

missing_objects <- required_objects[
  !vapply(
    required_objects,
    exists,
    logical(1),
    inherits = TRUE
  )
]

if (length(missing_objects) > 0L) {
  stop(
    "Missing required objects: ",
    paste(missing_objects, collapse = ", "),
    ". Run MODUL 6.12 and MODUL 6.13 first."
  )
}

# 2. Locate every normalized paragraph in its original text

paragraph_locations <- current_paragraph_layer |>
  left_join(
    target_text_sources |>
      select(
        text_id,
        raw_text_normalized
      ),
    by = "text_id"
  ) |>
  mutate(
    paragraph_text_normalized =
      normalize_structure_text(
        paragraph_text
      ),
    location_result = purrr::map2(
      raw_text_normalized,
      paragraph_text_normalized,
      locate_once
    )
  ) |>
  tidyr::unnest(location_result) |>
  rename(
    n_paragraph_matches = n_matches,
    paragraph_start = match_start,
    paragraph_end = match_end
  ) |>
  arrange(
    text_id,
    paragraph_number
  )

# 3. Build the paragraph-location audit

paragraph_location_audit <- paragraph_locations |>
  group_by(text_id) |>
  summarise(
    n_paragraphs = dplyr::n(),
    n_located_once = sum(
      n_paragraph_matches == 1L
    ),
    n_missing_paragraphs = sum(
      n_paragraph_matches == 0L
    ),
    n_duplicated_paragraph_matches = sum(
      n_paragraph_matches > 1L
    ),
    all_paragraphs_located_once = all(
      n_paragraph_matches == 1L
    ),
    .groups = "drop"
  ) |>
  arrange(text_id)

# 4. Display and validate the paragraph locations

message("Paragraph locations:")

print(
  paragraph_locations |>
    select(
      text_id,
      paragraph_id,
      paragraph_number,
      n_paragraph_matches,
      paragraph_start,
      paragraph_end
    ),
  n = Inf,
  width = Inf
)

message("Paragraph-location audit:")

print(
  paragraph_location_audit,
  n = Inf,
  width = Inf
)

if (
  any(
    !paragraph_location_audit$
      all_paragraphs_located_once
  )
) {
  warning(
    "At least one frozen paragraph was missing or matched ",
    "more than once in its original target text."
  )
} else {
  message(
    "All 15 frozen body paragraphs were located exactly once."
  )
}
```

# Modul 6.15 — Assign paragraphs to prescribed sections

```{r}
# MODUL 6.15 — Assign each body paragraph to its governing subheading

# 1. Create paragraph-to-section candidates

paragraph_section_candidates <- paragraph_locations |>
  select(
    text_id,
    model,
    paragraph_id,
    paragraph_number,
    paragraph_text,
    n_paragraph_matches,
    paragraph_start,
    paragraph_end
  ) |>
  left_join(
    heading_locations |>
      select(
        text_id,
        section_id,
        subheading_id,
        subheading_text,
        n_heading_matches,
        heading_end,
        next_heading_start
      ),
    by = "text_id",
    relationship = "many-to-many"
  ) |>
  mutate(
    paragraph_in_section = (
      n_paragraph_matches == 1L &
      n_heading_matches == 1L &
      paragraph_start > heading_end &
      (
        is.na(next_heading_start) |
        paragraph_start < next_heading_start
      )
    )
  )

# 2. Retain exactly one governing section per paragraph

paragraph_section_map <- paragraph_section_candidates |>
  group_by(
    text_id,
    model,
    paragraph_id,
    paragraph_number,
    paragraph_text,
    n_paragraph_matches,
    paragraph_start,
    paragraph_end
  ) |>
  summarise(
    n_section_matches = sum(
      paragraph_in_section,
      na.rm = TRUE
    ),
    section_id = if (
      n_section_matches == 1L
    ) {
      section_id[
        which(paragraph_in_section)[1]
      ]
    } else {
      NA_integer_
    },
    .groups = "drop"
  ) |>
  mutate(
    mapping_status = case_when(
      n_paragraph_matches != 1L ~
        "invalid_paragraph_location",
      n_section_matches == 1L ~
        "mapped",
      n_section_matches == 0L ~
        "outside_section_boundaries",
      n_section_matches > 1L ~
        "ambiguous_section_match",
      TRUE ~
        "unresolved"
    )
  ) |>
  left_join(
    prescribed_subheadings |>
      select(
        section_id,
        subheading_id,
        subheading_text
      ),
    by = "section_id"
  ) |>
  arrange(
    text_id,
    paragraph_number
  )

# 3. Display the paragraph-to-section mapping

print(
  paragraph_section_map |>
    select(
      text_id,
      paragraph_id,
      paragraph_number,
      section_id,
      subheading_id,
      mapping_status
    ),
  n = Inf,
  width = Inf
)

# 4. Validate the mapping result

if (
  any(
    paragraph_section_map$mapping_status != "mapped"
  )
) {
  warning(
    "At least one paragraph could not be assigned ",
    "unambiguously to a prescribed section."
  )
} else {
  message(
    "All 15 body paragraphs were assigned to exactly ",
    "one heading-defined section."
  )
}
```

# Modul 6.16 — Structural audit

```{r}
# MODUL 6.16 — Build and validate the final structural audit

# 1. Apply the updated subheading-object name

if (
  !exists("introduced_subheadings", inherits = TRUE) &&
  exists("prescribed_subheadings", inherits = TRUE)
) {
  introduced_subheadings <- prescribed_subheadings
}

required_objects <- c(
  "introduced_subheadings",
  "heading_audit",
  "paragraph_section_map",
  "target_text_sources"
)

missing_objects <- required_objects[
  !vapply(
    required_objects,
    exists,
    logical(1),
    inherits = TRUE
  )
]

if (length(missing_objects) > 0L) {
  stop(
    "Missing required objects: ",
    paste(missing_objects, collapse = ", ")
  )
}

# 2. Count mapped paragraphs in each introduced section

section_counts <- tidyr::crossing(
  text_id = target_text_sources$text_id,
  section_id = introduced_subheadings$section_id
) |>
  left_join(
    paragraph_section_map |>
      filter(mapping_status == "mapped") |>
      count(
        text_id,
        section_id,
        name = "n_paragraphs"
      ),
    by = c(
      "text_id",
      "section_id"
    )
  ) |>
  mutate(
    n_paragraphs = tidyr::replace_na(
      n_paragraphs,
      0L
    ),
    count_name = paste0(
      "n_paragraphs_section_",
      section_id
    )
  ) |>
  select(
    text_id,
    count_name,
    n_paragraphs
  ) |>
  tidyr::pivot_wider(
    names_from = count_name,
    values_from = n_paragraphs
  )

# 3. Audit paragraph mapping outcomes

paragraph_mapping_audit <- paragraph_section_map |>
  group_by(text_id) |>
  summarise(
    n_paragraphs_total = dplyr::n(),
    n_mapped_paragraphs = sum(
      mapping_status == "mapped"
    ),
    n_unmapped_paragraphs = sum(
      mapping_status != "mapped"
    ),
    n_ambiguous_paragraphs = sum(
      stringr::str_detect(
        mapping_status,
        "ambiguous"
      )
    ),
    n_outside_section_paragraphs = sum(
      mapping_status ==
        "outside_section_boundaries"
    ),
    .groups = "drop"
  )

# 4. Assemble the final structural audit

structural_audit <- target_text_sources |>
  select(text_id) |>
  left_join(
    heading_audit,
    by = "text_id"
  ) |>
  left_join(
    section_counts,
    by = "text_id"
  ) |>
  left_join(
    paragraph_mapping_audit,
    by = "text_id"
  ) |>
  mutate(
    n_prescribed_sections =
      nrow(introduced_subheadings),

    n_empty_sections =
      as.integer(
        n_paragraphs_section_1 == 0L
      ) +
      as.integer(
        n_paragraphs_section_2 == 0L
      ) +
      as.integer(
        n_paragraphs_section_3 == 0L
      ),

    section_mapping_valid = (
      n_prescribed_sections == 3L &
      n_detected_sections == 3L &
      all_sections_present &
      section_order_correct &
      n_missing_sections == 0L &
      n_duplicated_sections == 0L &
      n_empty_sections == 0L &
      n_mapped_paragraphs ==
        n_paragraphs_total &
      n_unmapped_paragraphs == 0L &
      n_ambiguous_paragraphs == 0L &
      n_outside_section_paragraphs == 0L
    )
  ) |>
  select(
    text_id,
    n_prescribed_sections,
    n_detected_sections,
    all_sections_present,
    section_order_correct,
    n_paragraphs_section_1,
    n_paragraphs_section_2,
    n_paragraphs_section_3,
    n_empty_sections,
    n_paragraphs_total,
    n_mapped_paragraphs,
    n_unmapped_paragraphs,
    n_ambiguous_paragraphs,
    n_outside_section_paragraphs,
    n_missing_sections,
    n_duplicated_sections,
    section_mapping_valid
  ) |>
  arrange(text_id)

# 5. Display and validate the audit

print(
  structural_audit,
  n = Inf,
  width = Inf
)

if (any(!structural_audit$section_mapping_valid)) {
  warning(
    "At least one target text has an invalid structural mapping. ",
    "Its later paragraph-to-subheading index must be NA."
  )
} else {
  message(
    "Structural audit passed: all four texts contain three ",
    "ordered, non-empty sections and all 15 paragraphs are ",
    "mapped unambiguously."
  )
}
```

# Modul 6.17 — Reconstruct sentence-to-paragraph mapping

```{r}
# MODUL 6.17 — Separate heading rows and map body sentences to validated paragraphs

# 1. Validate required inputs

required_objects <- c(
  "parsed_S_layer",
  "paragraph_section_map",
  "shared_prompt_headings",
  "normalize_structure_text"
)

missing_objects <- required_objects[
  !vapply(
    required_objects,
    exists,
    logical(1),
    inherits = TRUE
  )
]

if (length(missing_objects) > 0L) {
  stop(
    "Missing required objects: ",
    paste(missing_objects, collapse = ", ")
  )
}

required_sentence_columns <- c(
  "text_id",
  "sentence_id",
  "sentence_number",
  "sentence_text"
)

missing_sentence_columns <- setdiff(
  required_sentence_columns,
  names(parsed_S_layer)
)

if (length(missing_sentence_columns) > 0L) {
  stop(
    "parsed_S_layer is missing required columns: ",
    paste(missing_sentence_columns, collapse = ", ")
  )
}

# 2. Prepare normalized shared heading material

normalized_shared_headings <- shared_prompt_headings |>
  arrange(heading_order) |>
  mutate(
    heading_text_normalized =
      normalize_structure_text(heading_text)
  ) |>
  pull(heading_text_normalized)

strip_shared_heading_material <- function(
  normalized_sentence,
  normalized_headings
) {
  cleaned_sentence <- normalized_sentence

  for (current_heading in normalized_headings) {
    cleaned_sentence <- stringr::str_replace_all(
      cleaned_sentence,
      stringr::fixed(current_heading),
      " "
    )
  }

  stringr::str_squish(cleaned_sentence)
}

# 3. Separate heading material from body-sentence text

sentence_mapping_base <- parsed_S_layer |>
  transmute(
    text_id,
    sentence_id,
    sentence_number,
    sentence_text,

    sentence_text_normalized =
      normalize_structure_text(sentence_text),

    sentence_body_normalized =
      purrr::map_chr(
        sentence_text_normalized,
        strip_shared_heading_material,
        normalized_headings =
          normalized_shared_headings
      )
  ) |>
  mutate(
    contains_heading_material = (
      sentence_text_normalized !=
        sentence_body_normalized
    ),

    heading_only = (
      sentence_body_normalized == ""
    )
  )

# 4. Prepare the validated paragraph reference

paragraph_mapping_reference <- paragraph_section_map |>
  filter(
    mapping_status == "mapped"
  ) |>
  transmute(
    text_id,
    model,
    paragraph_id,
    paragraph_number,
    section_id,
    subheading_id,
    subheading_text,

    paragraph_text_normalized =
      normalize_structure_text(paragraph_text)
  )

# 5. Compare each non-empty body sentence with paragraphs from its text

sentence_paragraph_candidates <- sentence_mapping_base |>
  left_join(
    paragraph_mapping_reference,
    by = "text_id",
    relationship = "many-to-many"
  ) |>
  mutate(
    sentence_found_in_paragraph = purrr::map2_lgl(
      paragraph_text_normalized,
      sentence_body_normalized,
      function(
        current_paragraph,
        current_sentence
      ) {
        if (
          is.na(current_paragraph) ||
          is.na(current_sentence) ||
          current_sentence == ""
        ) {
          return(FALSE)
        }

        stringr::str_detect(
          current_paragraph,
          stringr::fixed(current_sentence)
        )
      }
    )
  )

# 6. Retain one mapping result per parsed sentence

sentence_paragraph_map <- sentence_paragraph_candidates |>
  group_by(
    text_id,
    sentence_id,
    sentence_number,
    sentence_text,
    sentence_text_normalized,
    sentence_body_normalized,
    contains_heading_material,
    heading_only
  ) |>
  summarise(
    n_paragraph_matches = sum(
      sentence_found_in_paragraph,
      na.rm = TRUE
    ),

    paragraph_number = if (
      n_paragraph_matches == 1L
    ) {
      paragraph_number[
        sentence_found_in_paragraph
      ][[1]]
    } else {
      NA_integer_
    },

    .groups = "drop"
  ) |>
  mutate(
    mapping_status = case_when(
      heading_only ~
        "heading_only",

      n_paragraph_matches == 1L ~
        "mapped",

      n_paragraph_matches == 0L ~
        "unmapped_body_sentence",

      n_paragraph_matches > 1L ~
        "ambiguous_body_sentence",

      TRUE ~
        "unresolved"
    )
  ) |>
  left_join(
    paragraph_mapping_reference |>
      select(
        text_id,
        model,
        paragraph_id,
        paragraph_number,
        section_id,
        subheading_id,
        subheading_text
      ),
    by = c(
      "text_id",
      "paragraph_number"
    ),
    relationship = "many-to-one"
  ) |>
  arrange(
    text_id,
    sentence_number
  )
```

# MODUL 6.18 — Resolve structural markers and high-confidence sentence mappings

```{r}

# 1. Validate required upstream objects

required_objects <- c(
  "sentence_paragraph_map",
  "paragraph_mapping_reference"
)

missing_objects <- required_objects[
  !vapply(
    required_objects,
    exists,
    logical(1),
    inherits = TRUE
  )
]

if (length(missing_objects) > 0L) {
  stop(
    "Missing required objects: ",
    paste(missing_objects, collapse = ", ")
  )
}

# 2. Identify numeric structural markers left after heading removal

structural_marker_keys <- sentence_paragraph_map |>
  filter(
    mapping_status == "unmapped_body_sentence",
    stringr::str_detect(
      sentence_body_normalized,
      "^\\d+$"
    )
  ) |>
  distinct(
    text_id,
    sentence_id
  ) |>
  mutate(
    structural_marker = TRUE
  )

# 3. Define a helper for controlled term-overlap matching

split_unique_terms <- function(text_value) {
  text_value <- stringr::str_squish(
    as.character(text_value)
  )

  if (
    is.na(text_value) ||
    text_value == ""
  ) {
    return(character())
  }

  unique(
    stringr::str_split(
      text_value,
      "\\s+"
    )[[1]]
  )
}

# 4. Score the remaining unmapped body sentences against paragraphs

fallback_candidate_scores <- sentence_paragraph_map |>
  filter(
    mapping_status == "unmapped_body_sentence"
  ) |>
  anti_join(
    structural_marker_keys,
    by = c(
      "text_id",
      "sentence_id"
    )
  ) |>
  select(
    text_id,
    sentence_id,
    sentence_number,
    sentence_body_normalized
  ) |>
  left_join(
    paragraph_mapping_reference |>
      select(
        text_id,
        paragraph_id,
        paragraph_number,
        section_id,
        subheading_id,
        paragraph_text_normalized
      ),
    by = "text_id",
    relationship = "many-to-many"
  ) |>
  mutate(
    sentence_terms = purrr::map(
      sentence_body_normalized,
      split_unique_terms
    ),

    paragraph_terms = purrr::map(
      paragraph_text_normalized,
      split_unique_terms
    ),

    n_sentence_terms = purrr::map_int(
      sentence_terms,
      length
    ),

    n_shared_terms = purrr::map2_int(
      sentence_terms,
      paragraph_terms,
      function(
        current_sentence_terms,
        current_paragraph_terms
      ) {
        length(
          intersect(
            current_sentence_terms,
            current_paragraph_terms
          )
        )
      }
    ),

    sentence_term_coverage = if_else(
      n_sentence_terms > 0L,
      n_shared_terms / n_sentence_terms,
      NA_real_
    )
  ) |>
  group_by(
    text_id,
    sentence_id
  ) |>
  arrange(
    desc(sentence_term_coverage),
    paragraph_number,
    .by_group = TRUE
  ) |>
  mutate(
    candidate_rank = dplyr::row_number(),

    best_coverage = dplyr::first(
      sentence_term_coverage
    ),

    second_best_coverage = dplyr::nth(
      sentence_term_coverage,
      2L,
      default = 0
    ),

    coverage_margin = (
      best_coverage -
      second_best_coverage
    )
  ) |>
  ungroup()

# 5. Accept only clear high-confidence fallback assignments

fallback_assignments <- fallback_candidate_scores |>
  filter(
    candidate_rank == 1L,
    best_coverage >= 0.70,
    coverage_margin >= 0.20
  ) |>
  transmute(
    text_id,
    sentence_id,

    fallback_paragraph_number =
      paragraph_number,

    fallback_coverage =
      best_coverage,

    fallback_margin =
      coverage_margin
  )

# 6. Apply structural-marker and fallback classifications

sentence_paragraph_map <- sentence_paragraph_map |>
  select(
    -any_of(
      c(
        "model",
        "paragraph_id",
        "section_id",
        "subheading_id",
        "subheading_text"
      )
    )
  ) |>
  left_join(
    structural_marker_keys,
    by = c(
      "text_id",
      "sentence_id"
    )
  ) |>
  left_join(
    fallback_assignments,
    by = c(
      "text_id",
      "sentence_id"
    )
  ) |>
  mutate(
    structural_marker = tidyr::replace_na(
      structural_marker,
      FALSE
    ),

    paragraph_number = dplyr::coalesce(
      fallback_paragraph_number,
      paragraph_number
    ),

    mapping_status = case_when(
      structural_marker ~
        "structural_marker",

      !is.na(fallback_paragraph_number) ~
        "mapped_by_high_overlap",

      TRUE ~
        mapping_status
    )
  ) |>
  select(
    -fallback_paragraph_number
  ) |>
  left_join(
    paragraph_mapping_reference |>
      select(
        text_id,
        model,
        paragraph_id,
        paragraph_number,
        section_id,
        subheading_id,
        subheading_text
      ),
    by = c(
      "text_id",
      "paragraph_number"
    ),
    relationship = "many-to-one"
  ) |>
  arrange(
    text_id,
    sentence_number
  )

# 7. Create the finalized body-sentence map

body_sentence_map <- sentence_paragraph_map |>
  filter(
    mapping_status %in% c(
      "mapped",
      "mapped_by_high_overlap"
    )
  )

# 8. Build the finalized mapping audit

sentence_mapping_audit <- sentence_paragraph_map |>
  group_by(text_id) |>
  summarise(
    n_parsed_sentence_rows = dplyr::n(),

    n_heading_only_rows = sum(
      mapping_status == "heading_only"
    ),

    n_structural_marker_rows = sum(
      mapping_status == "structural_marker"
    ),

    n_exact_mapped_body_sentences = sum(
      mapping_status == "mapped"
    ),

    n_overlap_mapped_body_sentences = sum(
      mapping_status ==
        "mapped_by_high_overlap"
    ),

    n_unmapped_body_sentences = sum(
      mapping_status ==
        "unmapped_body_sentence"
    ),

    n_ambiguous_body_sentences = sum(
      mapping_status ==
        "ambiguous_body_sentence"
    ),

    all_body_sentences_mapped_once = (
      n_unmapped_body_sentences == 0L &
      n_ambiguous_body_sentences == 0L
    ),

    .groups = "drop"
  ) |>
  arrange(text_id)

# 9. Display and validate the finalized mapping

message("Accepted high-confidence fallback assignments:")

print(
  fallback_assignments,
  n = Inf,
  width = Inf
)

message("Final sentence-mapping audit:")

print(
  sentence_mapping_audit,
  n = Inf,
  width = Inf
)

if (
  any(
    !sentence_mapping_audit$
      all_body_sentences_mapped_once
  )
) {
  stop(
    "Residual sentence mapping remains incomplete. ",
    "The segment tables must not yet be frozen."
  )
} else {
  message(
    "Sentence mapping finalized: all body sentences are ",
    "assigned once, while heading-only and structural-marker ",
    "rows remain excluded."
  )
}
```

# Modul 6.19 — Build validated body-token layer

```{r}

# MODUL 6.19 — Build the validated heading-free body-token layer

# 1. Validate required inputs

required_objects <- c(
  "current_texts_parsed",
  "body_sentence_map",
  "sentence_mapping_audit",
  "structural_audit",
  "lsa_pos_keep",
  "normalize_structure_text"
)

missing_objects <- required_objects[
  !vapply(
    required_objects,
    exists,
    logical(1),
    inherits = TRUE
  )
]

if (length(missing_objects) > 0L) {
  stop(
    "Missing required objects: ",
    paste(missing_objects, collapse = ", ")
  )
}

if (
  !all(structural_audit$section_mapping_valid) ||
  !all(
    sentence_mapping_audit$
      all_body_sentences_mapped_once
  )
) {
  stop(
    "The finalized structural mapping is incomplete."
  )
}

# 2. Prepare the finalized sentence-to-structure bridge

sentence_structure_bridge <- body_sentence_map |>
  transmute(
    text_id,
    sentence_id,
    sentence_number,
    sentence_mapping_status = mapping_status,
    sentence_body_normalized,
    contains_heading_material,
    model,
    paragraph_id,
    paragraph_number,
    section_id,
    subheading_id,
    subheading_text
  )

duplicate_sentence_keys <- sentence_structure_bridge |>
  count(
    text_id,
    sentence_id
  ) |>
  filter(n > 1L)

if (nrow(duplicate_sentence_keys) > 0L) {
  stop(
    "Duplicate finalized sentence keys were found."
  )
}

# 3. Join parsed tokens to the validated structure

joined_body_tokens <- current_texts_parsed |>
  rename(
    parsed_paragraph_id = paragraph_id
  ) |>
  mutate(
    token_id_numeric = suppressWarnings(
      as.integer(token_id)
    )
  )

if (any(is.na(joined_body_tokens$token_id_numeric))) {
  stop(
    "At least one token_id could not be converted ",
    "to numeric order."
  )
}

joined_body_tokens <- joined_body_tokens |>
  inner_join(
    sentence_structure_bridge,
    by = c(
      "text_id",
      "sentence_id"
    ),
    relationship = "many-to-one"
  ) |>
  arrange(
    text_id,
    sentence_number,
    token_id_numeric
  ) |>
  group_by(
    text_id,
    sentence_id
  ) |>
  mutate(
    token_position = dplyr::row_number(),

    token_normalized =
      normalize_structure_text(token)
  ) |>
  ungroup()

# 4. Locate the first body token in the mixed sentence

mixed_sentence_start <- joined_body_tokens |>
  filter(
    contains_heading_material
  ) |>
  group_by(
    text_id,
    sentence_id
  ) |>
  summarise(
    first_body_word = stringr::str_extract(
      dplyr::first(
        sentence_body_normalized
      ),
      "^\\S+"
    ),

    body_start_position = match(
      first_body_word,
      token_normalized
    ),

    .groups = "drop"
  )

print(
  mixed_sentence_start,
  n = Inf,
  width = Inf
)

if (
  nrow(mixed_sentence_start) != 1L ||
  any(is.na(
    mixed_sentence_start$
      body_start_position
  ))
) {
  stop(
    "The beginning of the mixed body sentence could not be located."
  )
}

# 5. Retain ordinary sentences and trim only the mixed sentence

body_token_layer <- joined_body_tokens |>
  left_join(
    mixed_sentence_start |>
      select(
        text_id,
        sentence_id,
        body_start_position
      ),
    by = c(
      "text_id",
      "sentence_id"
    ),
    relationship = "many-to-one"
  ) |>
  filter(
    !contains_heading_material |
      token_position >= body_start_position
  ) |>
  mutate(
    lemma_clean = stringr::str_to_lower(
      stringr::str_squish(
        as.character(lemma)
      )
    ),

    pos = stringr::str_to_upper(
      as.character(pos)
    ),

    included_in_lsa = (
      pos %in% lsa_pos_keep &
      !is.na(lemma_clean) &
      lemma_clean != ""
    )
  ) |>
  arrange(
    text_id,
    sentence_number,
    token_id
  )

# 7. Validate final structural counts

body_token_audit <- body_token_layer |>
  summarise(
    n_body_sentences = n_distinct(
      interaction(
        text_id,
        sentence_id,
        drop = TRUE
      )
    ),

    n_paragraphs = n_distinct(
      interaction(
        text_id,
        paragraph_number,
        drop = TRUE
      )
    ),

    n_body_tokens = dplyr::n(),

    n_lsa_tokens = sum(
      included_in_lsa
    )
  )

print(
  body_token_audit,
  n = Inf,
  width = Inf
)

if (
  body_token_audit$n_body_sentences != 106L ||
  body_token_audit$n_paragraphs != 15L
) {
  stop(
    "The final body-token structure has unexpected counts."
  )
}

message(
  "Validated heading-free body-token layer created successfully."
)
```

# MODUL 6.20 — Build ordered sentence and paragraph segment tables

```{r}

# MODUL 6.20 — Build ordered sentence and paragraph segment tables

# 1. Validate required inputs

required_objects <- c(
  "body_token_layer",
  "paragraph_section_map"
)

missing_objects <- required_objects[
  !vapply(
    required_objects,
    exists,
    logical(1),
    inherits = TRUE
  )
]

if (length(missing_objects) > 0L) {
  stop(
    "Missing required objects: ",
    paste(missing_objects, collapse = ", ")
  )
}

required_body_columns <- c(
  "text_id",
  "model",
  "sentence_id",
  "sentence_number",
  "sentence_mapping_status",
  "sentence_body_normalized",
  "paragraph_id",
  "paragraph_number",
  "section_id",
  "subheading_id",
  "subheading_text",
  "token_id_numeric",
  "token",
  "lemma_clean",
  "included_in_lsa"
)

missing_body_columns <- setdiff(
  required_body_columns,
  names(body_token_layer)
)

if (length(missing_body_columns) > 0L) {
  stop(
    "body_token_layer is missing required columns: ",
    paste(missing_body_columns, collapse = ", ")
  )
}

# 2. Establish correct numeric token order

ordered_body_tokens <- body_token_layer |>
  arrange(
    text_id,
    paragraph_number,
    sentence_number,
    token_id_numeric
  )

# 3. Build the sentence segment table

sentence_segment_table <- ordered_body_tokens |>
  group_by(
    text_id,
    model,
    sentence_id,
    sentence_number,
    sentence_mapping_status,
    sentence_body_normalized,
    paragraph_id,
    paragraph_number,
    section_id,
    subheading_id,
    subheading_text
  ) |>
  summarise(
    n_tokens_total = dplyr::n(),

    n_lsa_tokens = sum(
      included_in_lsa,
      na.rm = TRUE
    ),

    body_token_text = paste(
      token[
        !is.na(token) &
        token != ""
      ],
      collapse = " "
    ),

    lsa_lemma_text = paste(
      lemma_clean[
        included_in_lsa %in% TRUE &
        !is.na(lemma_clean) &
        lemma_clean != ""
      ],
      collapse = " "
    ),

    .groups = "drop"
  ) |>
  mutate(
    segment_id = paste0(
      text_id,
      "_S",
      stringr::str_pad(
        sentence_number,
        width = 3,
        pad = "0"
      )
    )
  ) |>
  relocate(
    segment_id,
    .after = model
  ) |>
  arrange(
    text_id,
    sentence_number
  )

# 4. Summarize ordered body tokens by paragraph

paragraph_token_summary <- ordered_body_tokens |>
  group_by(
    text_id,
    paragraph_id,
    paragraph_number
  ) |>
  summarise(
    n_sentences = dplyr::n_distinct(
      sentence_id
    ),

    n_tokens_total = dplyr::n(),

    n_lsa_tokens = sum(
      included_in_lsa,
      na.rm = TRUE
    ),

    body_token_text = paste(
      token[
        !is.na(token) &
        token != ""
      ],
      collapse = " "
    ),

    lsa_lemma_text = paste(
      lemma_clean[
        included_in_lsa %in% TRUE &
        !is.na(lemma_clean) &
        lemma_clean != ""
      ],
      collapse = " "
    ),

    .groups = "drop"
  )

# 5. Build the paragraph segment table

paragraph_segment_table <- paragraph_section_map |>
  filter(
    mapping_status == "mapped"
  ) |>
  select(
    text_id,
    model,
    paragraph_id,
    paragraph_number,
    section_id,
    subheading_id,
    subheading_text,
    paragraph_text
  ) |>
  left_join(
    paragraph_token_summary,
    by = c(
      "text_id",
      "paragraph_id",
      "paragraph_number"
    ),
    relationship = "one-to-one"
  ) |>
  mutate(
    segment_id = paste0(
      text_id,
      "_P",
      stringr::str_pad(
        paragraph_number,
        width = 2,
        pad = "0"
      )
    )
  ) |>
  relocate(
    segment_id,
    .after = model
  ) |>
  arrange(
    text_id,
    paragraph_number
  )

# 6. Validate segment counts and structural completeness

segment_table_audit <- tibble::tibble(
  segment_level = c(
    "sentence",
    "paragraph"
  ),

  expected_segments = c(
    106L,
    15L
  ),

  observed_segments = c(
    nrow(sentence_segment_table),
    nrow(paragraph_segment_table)
  ),

  n_empty_lsa_segments = c(
    sum(
      sentence_segment_table$n_lsa_tokens == 0L
    ),

    sum(
      paragraph_segment_table$n_lsa_tokens == 0L
    )
  )
) |>
  mutate(
    segment_count_valid = (
      expected_segments ==
      observed_segments
    )
  )

duplicate_sentence_segments <- sentence_segment_table |>
  count(segment_id) |>
  filter(n > 1L)

duplicate_paragraph_segments <- paragraph_segment_table |>
  count(segment_id) |>
  filter(n > 1L)

if (
  any(!segment_table_audit$segment_count_valid) ||
  nrow(duplicate_sentence_segments) > 0L ||
  nrow(duplicate_paragraph_segments) > 0L ||
  any(is.na(paragraph_segment_table$n_tokens_total))
) {
  stop(
    "Sentence or paragraph segment validation failed."
  )
}

print(
  segment_table_audit,
  n = Inf,
  width = Inf
)

message(
  "Ordered segment tables created: ",
  nrow(sentence_segment_table),
  " sentences and ",
  nrow(paragraph_segment_table),
  " paragraphs."
)
```

# Modul 6.21 — Build the ordered whole-body-text table and final audit

```{r}

# MODUL 6.21 — Build the ordered whole-body-text table and final audit

# 1. Validate required inputs

required_objects <- c(
  "ordered_body_tokens",
  "sentence_segment_table",
  "paragraph_segment_table",
  "segment_table_audit"
)

missing_objects <- required_objects[
  !vapply(
    required_objects,
    exists,
    logical(1),
    inherits = TRUE
  )
]

if (length(missing_objects) > 0L) {
  stop(
    "Missing required objects: ",
    paste(missing_objects, collapse = ", ")
  )
}

# 2. Build one ordered whole-body-text segment per target text

whole_body_text_segment_table <- ordered_body_tokens |>
  arrange(
    text_id,
    paragraph_number,
    sentence_number,
    token_id_numeric
  ) |>
  group_by(
    text_id,
    model
  ) |>
  summarise(
    n_paragraphs = dplyr::n_distinct(
      paragraph_id
    ),

    n_sentences = dplyr::n_distinct(
      sentence_id
    ),

    n_tokens_total = dplyr::n(),

    n_lsa_tokens = sum(
      included_in_lsa,
      na.rm = TRUE
    ),

    body_token_text = paste(
      token[
        !is.na(token) &
        token != ""
      ],
      collapse = " "
    ),

    lsa_lemma_text = paste(
      lemma_clean[
        included_in_lsa %in% TRUE &
        !is.na(lemma_clean) &
        lemma_clean != ""
      ],
      collapse = " "
    ),

    .groups = "drop"
  ) |>
  mutate(
    segment_id = paste0(
      text_id,
      "_BODY"
    )
  ) |>
  relocate(
    segment_id,
    .after = model
  ) |>
  arrange(text_id)

# 3. Compare whole-text counts with sentence and paragraph tables

expected_text_counts <- paragraph_segment_table |>
  group_by(
    text_id,
    model
  ) |>
  summarise(
    expected_paragraphs = dplyr::n(),

    expected_sentences = sum(
      n_sentences,
      na.rm = TRUE
    ),

    expected_tokens = sum(
      n_tokens_total,
      na.rm = TRUE
    ),

    expected_lsa_tokens = sum(
      n_lsa_tokens,
      na.rm = TRUE
    ),

    .groups = "drop"
  )

whole_text_count_audit <- whole_body_text_segment_table |>
  select(
    text_id,
    model,
    n_paragraphs,
    n_sentences,
    n_tokens_total,
    n_lsa_tokens
  ) |>
  left_join(
    expected_text_counts,
    by = c(
      "text_id",
      "model"
    ),
    relationship = "one-to-one"
  ) |>
  mutate(
    paragraph_count_valid = (
      n_paragraphs ==
      expected_paragraphs
    ),

    sentence_count_valid = (
      n_sentences ==
      expected_sentences
    ),

    token_count_valid = (
      n_tokens_total ==
      expected_tokens
    ),

    lsa_token_count_valid = (
      n_lsa_tokens ==
      expected_lsa_tokens
    ),

    all_counts_valid = (
      paragraph_count_valid &
      sentence_count_valid &
      token_count_valid &
      lsa_token_count_valid
    )
  ) |>
  arrange(text_id)

# 4. Extend the segment-level audit

final_segment_table_audit <- bind_rows(
  segment_table_audit,

  tibble::tibble(
    segment_level = "whole_body_text",
    expected_segments = 4L,
    observed_segments =
      nrow(whole_body_text_segment_table),
    n_empty_lsa_segments = sum(
      whole_body_text_segment_table$
        n_lsa_tokens == 0L
    ),
    segment_count_valid = (
      nrow(
        whole_body_text_segment_table
      ) == 4L
    )
  )
)

# 5. Validate the final preparation structure

duplicate_whole_text_segments <-
  whole_body_text_segment_table |>
  count(segment_id) |>
  filter(n > 1L)

if (
  any(
    !final_segment_table_audit$
      segment_count_valid
  ) ||
  any(
    !whole_text_count_audit$
      all_counts_valid
  ) ||
  nrow(
    duplicate_whole_text_segments
  ) > 0L ||
  any(
    whole_body_text_segment_table$
      n_lsa_tokens == 0L
  )
) {
  stop(
    "Final whole-body-text or cross-level validation failed."
  )
}

# 6. Display the final audits

message("Whole-body-text count audit:")

print(
  whole_text_count_audit,
  n = Inf,
  width = Inf
)

message("Final segment-table audit:")

print(
  final_segment_table_audit,
  n = Inf,
  width = Inf
)

message(
  "Final preparation tables validated: ",
  nrow(sentence_segment_table),
  " sentences, ",
  nrow(paragraph_segment_table),
  " paragraphs, and ",
  nrow(whole_body_text_segment_table),
  " whole body texts."
)
```

# Modul 6 - 6.22 — Freeze and export the validated MODUL 6 preparation inputs

```{r}
# MODUL 6.22 — Freeze and export the validated MODUL 6 preparation inputs

# 1. Validate all objects required for the frozen bundle

required_objects <- c(
  "lsa_dimensions",
  "lsa_pos_keep",
  "module_6_output_dir",
  "shared_prompt_headings",
  "introduced_subheadings",
  "paragraph_section_map",
  "structural_audit",
  "sentence_paragraph_map",
  "body_sentence_map",
  "sentence_mapping_audit",
  "fallback_assignments",
  "sentence_segment_table",
  "paragraph_segment_table",
  "whole_body_text_segment_table",
  "final_segment_table_audit",
  "whole_text_count_audit"
)

missing_objects <- required_objects[
  !vapply(
    required_objects,
    exists,
    logical(1),
    inherits = TRUE
  )
]

if (length(missing_objects) > 0L) {
  stop(
    "Missing required objects: ",
    paste(missing_objects, collapse = ", ")
  )
}

# 2. Confirm that all final audits passed

if (
  any(!structural_audit$section_mapping_valid) ||
  any(
    !sentence_mapping_audit$
      all_body_sentences_mapped_once
  ) ||
  any(
    !final_segment_table_audit$
      segment_count_valid
  ) ||
  any(
    !whole_text_count_audit$
      all_counts_valid
  )
) {
  stop(
    "At least one structural or segment-table audit failed. ",
    "The MODUL 6 inputs must not be frozen."
  )
}

if (
  nrow(sentence_segment_table) != 106L ||
  nrow(paragraph_segment_table) != 15L ||
  nrow(whole_body_text_segment_table) != 4L
) {
  stop(
    "Unexpected final segment counts."
  )
}

# 3. Define the frozen-output paths

m6_prep_output_dir <- "M6_prep_outputs"

dir.create(
  m6_prep_output_dir,
  showWarnings = FALSE,
  recursive = TRUE
)

dir.create(
  module_6_output_dir,
  showWarnings = FALSE,
  recursive = TRUE
)

lsa_input_bundle_path <- file.path(
  m6_prep_output_dir,
  "modul_6_LSA_inputs.rds"
)

paragraph_map_csv_path <- file.path(
  module_6_output_dir,
  "modul_6_paragraph_section_map.csv"
)

structural_audit_csv_path <- file.path(
  module_6_output_dir,
  "modul_6_structural_audit.csv"
)

# 4. Assemble the frozen MODUL 6 input bundle

modul_6_lsa_inputs <- list(
  metadata = list(
    module =
      "MODUL 6 LSA structural preparation",

    source_qmd =
      "CCP_generation_Modul_6_LSA.qmd",

    lsa_dimensions =
      lsa_dimensions,

    lsa_pos_keep =
      lsa_pos_keep,

    n_target_texts =
      nrow(
        whole_body_text_segment_table
      ),

    n_sentence_segments =
      nrow(
        sentence_segment_table
      ),

    n_paragraph_segments =
      nrow(
        paragraph_segment_table
      ),

    n_whole_body_text_segments =
      nrow(
        whole_body_text_segment_table
      )
  ),

  shared_prompt_headings =
    shared_prompt_headings,

  introduced_subheadings =
    introduced_subheadings,

  paragraph_section_map =
    paragraph_section_map,

  structural_audit =
    structural_audit,

  sentence_paragraph_map =
    sentence_paragraph_map,

  body_sentence_map =
    body_sentence_map,

  sentence_mapping_audit =
    sentence_mapping_audit,

  fallback_assignments =
    fallback_assignments,

  sentence_segment_table =
    sentence_segment_table,

  paragraph_segment_table =
    paragraph_segment_table,

  whole_body_text_segment_table =
    whole_body_text_segment_table,

  final_segment_table_audit =
    final_segment_table_audit,

  whole_text_count_audit =
    whole_text_count_audit
)

# 5. Write the frozen bundle and inspection CSVs

saveRDS(
  modul_6_lsa_inputs,
  lsa_input_bundle_path,
  version = 3
)

readr::write_csv(
  paragraph_section_map,
  paragraph_map_csv_path
)

readr::write_csv(
  structural_audit,
  structural_audit_csv_path
)

# 6. Reload and validate the frozen bundle

frozen_bundle_check <- readRDS(
  lsa_input_bundle_path
)

required_bundle_entries <- c(
  "metadata",
  "shared_prompt_headings",
  "introduced_subheadings",
  "paragraph_section_map",
  "structural_audit",
  "sentence_paragraph_map",
  "body_sentence_map",
  "sentence_mapping_audit",
  "fallback_assignments",
  "sentence_segment_table",
  "paragraph_segment_table",
  "whole_body_text_segment_table",
  "final_segment_table_audit",
  "whole_text_count_audit"
)

missing_bundle_entries <- setdiff(
  required_bundle_entries,
  names(frozen_bundle_check)
)

if (length(missing_bundle_entries) > 0L) {
  stop(
    "Frozen bundle is missing entries: ",
    paste(
      missing_bundle_entries,
      collapse = ", "
    )
  )
}

if (
  nrow(
    frozen_bundle_check$
      sentence_segment_table
  ) != 106L ||
  nrow(
    frozen_bundle_check$
      paragraph_segment_table
  ) != 15L ||
  nrow(
    frozen_bundle_check$
      whole_body_text_segment_table
  ) != 4L
) {
  stop(
    "Frozen bundle validation failed."
  )
}

# 7. Display the exported-file manifest

freeze_manifest <- tibble::tibble(
  output_type = c(
    "Frozen RDS bundle",
    "Paragraph-section audit CSV",
    "Structural audit CSV"
  ),

  file_path = c(
    normalizePath(
      lsa_input_bundle_path,
      mustWork = TRUE
    ),

    normalizePath(
      paragraph_map_csv_path,
      mustWork = TRUE
    ),

    normalizePath(
      structural_audit_csv_path,
      mustWork = TRUE
    )
  ),

  size_bytes = file.info(
    c(
      lsa_input_bundle_path,
      paragraph_map_csv_path,
      structural_audit_csv_path
    )
  )$size
)

print(
  freeze_manifest,
  n = Inf,
  width = Inf
)

message(
  "MODUL 6 structural preparation frozen successfully."
)
```
# ---- END OF MODUL 6_PREP_FOR_LSA-------------------


# Note: source_mapping_position

= position inferred from the order of mappings in inv_so.txt

segment_id = order among the retained topic-filtered sentences from that document


# Independent check 1

```{r}
bundle_path <- file.path(
  "../data_prep_CCP_output",
  "l1_corpus_reference_bundle_corrected.rds"
)

l1_bundle <- readRDS(bundle_path)

doc_tab <- l1_bundle$documents
sent_tab <- l1_bundle$sentences
tok_tab <- l1_bundle$tokens

summary_table <- tibble::tibble(
  metric   = c(
    "documents",
    "sentences",
    "tokens",
    "LSA-empty sentences",
    "LSA-empty documents",
    "UDPipe-resplit sentences"
  ),
  observed = c(
    nrow(doc_tab),
    nrow(sent_tab),
    nrow(tok_tab),
    sum(!sent_tab$has_selected_lsa_terms),
    sum(!doc_tab$has_selected_lsa_terms),
    sum(sent_tab$udpipe_resplit)
  )
)

print(summary_table, n = Inf)
```

# Independent check 2

```{r}
bundle_path <- file.path(
  "../data_prep_CCP_output",
  "l1_corpus_reference_bundle_corrected.rds"
)

l1_bundle <- readRDS(bundle_path)

set.seed(20240714)  # for reproducibility; change or drop if you like

sent_sample <- l1_bundle$sentences |>
  dplyr::select(
    document_uid,
    sentence_uid,
    sentence_text = dplyr::any_of(c(
      "sentence_text",
      "source_sentence_text",
      "text_unit",
      "text"
    ))
  ) |>
  dplyr::distinct() |>
  dplyr::slice_sample(n = 10)

print(sent_sample, n = Inf, width = Inf)
```

# # Independent check 3

```{r}
doc_sample <- l1_bundle$documents |>
  dplyr::semi_join(
    sent_sample,
    by = "document_uid"
  )

print(doc_sample, n = Inf, width = Inf)
```

# # Independent check 4

```{r}
bundle_path <- file.path(
  "../data_prep_CCP_output",
  "l1_corpus_reference_bundle_corrected.rds"
)

l1_bundle <- readRDS(bundle_path)
doc_tab   <- l1_bundle$documents

na_by_corpus <- function(df, corpus_col = "corpus_name") {
  corpus_vals <- df[[corpus_col]]

  purrr::map_dfr(
    unique(corpus_vals),
    function(corpus) {
      sub <- df[corpus_vals == corpus, , drop = FALSE]

      tibble::tibble(
        corpus = corpus,
        column = names(sub),
        n_na   = vapply(sub, function(x) sum(is.na(x)), integer(1)),
        n_row  = nrow(sub)
      ) |>
        dplyr::filter(n_na > 0L)
    }
  ) |>
    dplyr::arrange(corpus, dplyr::desc(n_na))
}

na_doc_summary <- na_by_corpus(doc_tab, corpus_col = "corpus_name")

print(na_doc_summary, n = Inf)
```



