---
title: "CCP_generation_Modul_6_LSA_actual"
format: html
editor: visual
execute:
  echo: true
  warning: true
  message: true
---

# Intro: # MODUL 6 — Actual Latent Semantic Analysis

This document constructs the fixed L1-Corpus semantic reference space, folds the prepared target segments into that space without refitting it, and calculates the x3 and x4 semantic-similarity measures.

The document reads only frozen inputs. It does not reconstruct the L1 corpus, rerun UDPipe, reread the raw target texts, relocate headings, or rebuild the validated sentence and paragraph structure.


#

```{r}
# MODUL 6.23 — Set up the numerical LSA workflow

# 1. Validate the required packages

required_packages <- c(
  "dplyr",
  "tidyr",
  "purrr",
  "readr",
  "stringr",
  "tibble",
  "quanteda",
  "Matrix",
  "irlba"
)

missing_packages <- required_packages[
  !vapply(
    required_packages,
    requireNamespace,
    logical(1),
    quietly = TRUE
  )
]

if (length(missing_packages) > 0L) {
  stop(
    "Missing required packages: ",
    paste(
      missing_packages,
      collapse = ", "
    ),
    ". Install them before running MODUL 6."
  )
}

# 2. Attach the required packages

library(dplyr)
library(tidyr)
library(purrr)
library(readr)
library(stringr)
library(tibble)
library(quanteda)
library(Matrix)
library(irlba)

# 3. Set general options

options(
  dplyr.summarise.inform = FALSE,
  scipen = 999
)

# 4. Define the fixed LSA configuration

lsa_dimensions <- 50L

lsa_pos_keep <- c(
  "NOUN",
  "PROPN",
  "VERB",
  "ADJ",
  "ADV",
  "CCONJ",
  "SCONJ"
)

lsa_tf_method <- "log1p"

lsa_idf_method <- "log_N_over_df"

lsa_random_seed <- 20260714L

# 5. Validate the fixed configuration

if (
  lsa_dimensions != 50L ||
  length(lsa_pos_keep) != 7L ||
  anyDuplicated(lsa_pos_keep) > 0L
) {
  stop(
    "The fixed MODUL 6 LSA configuration is invalid."
  )
}

# 6. Display the active configuration

lsa_configuration <- tibble::tibble(
  parameter = c(
    "Dimensions",
    "TF weighting",
    "IDF weighting",
    "Included UPOS categories",
    "SVD random seed"
  ),

  value = c(
    as.character(lsa_dimensions),
    "log(1 + tf)",
    "log(N / df)",
    paste(
      lsa_pos_keep,
      collapse = ", "
    ),
    as.character(lsa_random_seed)
  )
)

print(
  lsa_configuration,
  n = Inf,
  width = Inf
)

message(
  "MODUL 6 actual LSA setup complete."
)
```
# MODUL 6.24 — Register and validate frozen paths

```{r}

# MODUL 6.24 — Register and validate frozen paths

project_root <- normalizePath(
  "..",
  winslash = "/",
  mustWork = TRUE
)

actual_lsa_input_paths <- c(
  l1_reference_bundle = file.path(
    project_root,
    "data_prep_CCP_output",
    "l1_corpus_reference_bundle_corrected.rds"
  ),
  target_input_bundle = file.path(
    project_root,
    "CCP_generation_pipeL",
    "M6_prep_outputs",
    "modul_6_LSA_inputs.rds"
  )
)

actual_lsa_output_dir <- file.path(
  project_root,
  "measure_outputs"
)

dir.create(
  actual_lsa_output_dir,
  recursive = TRUE,
  showWarnings = FALSE
)

missing_inputs <- actual_lsa_input_paths[
  !file.exists(actual_lsa_input_paths)
]

if (length(missing_inputs) > 0L) {
  stop(
    "Missing frozen input files: ",
    paste(
      names(missing_inputs),
      collapse = ", "
    )
  )
}

actual_lsa_path_audit <- tibble::tibble(
  input = names(actual_lsa_input_paths),
  path = unname(actual_lsa_input_paths),
  exists = file.exists(actual_lsa_input_paths)
)

print(
  actual_lsa_path_audit,
  n = Inf,
  width = Inf
)

message(
  "Frozen actual LSA paths validated successfully."
)
```

# # MODUL 6.25 — Load the frozen L1 and target bundles

```{r}
# MODUL 6.25 — Load the frozen L1 and target bundles

l1_reference_bundle <- readRDS(
  actual_lsa_input_paths[["l1_reference_bundle"]]
)

target_input_bundle <- readRDS(
  actual_lsa_input_paths[["target_input_bundle"]]
)

required_l1_objects <- c(
  "documents",
  "sentences",
  "tokens",
  "metadata"
)

required_target_objects <- c(
  "sentence_segment_table",
  "paragraph_segment_table",
  "whole_body_text_segment_table",
  "shared_prompt_headings",
  "introduced_subheadings",
  "paragraph_section_map",
  "structural_audit",
  "sentence_mapping_audit"
)

missing_l1_objects <- setdiff(
  required_l1_objects,
  names(l1_reference_bundle)
)

missing_target_objects <- setdiff(
  required_target_objects,
  names(target_input_bundle)
)

if (length(missing_l1_objects) > 0L) {
  stop(
    "Missing L1 bundle objects: ",
    paste(missing_l1_objects, collapse = ", ")
  )
}

if (length(missing_target_objects) > 0L) {
  stop(
    "Missing target bundle objects: ",
    paste(missing_target_objects, collapse = ", ")
  )
}

if (!isTRUE(
  l1_reference_bundle$metadata$final_audit_passed
)) {
  stop(
    "The frozen L1 reference bundle did not pass its final audit."
  )
}

bundle_audit <- tibble::tibble(
  bundle = c(
    "L1 reference bundle",
    "Target preparation bundle"
  ),
  n_objects = c(
    length(l1_reference_bundle),
    length(target_input_bundle)
  ),
  valid = TRUE
)

print(
  bundle_audit,
  n = Inf
)

message(
  "Frozen L1 and target bundles loaded successfully."
)
```
# MODUL 6.26 — Extract and validate the required objects

```{r}


# MODUL 6.26 — Extract and validate the required objects

# Extract the frozen L1 objects

l1_documents <- l1_reference_bundle$documents
l1_sentences <- l1_reference_bundle$sentences
l1_tokens <- l1_reference_bundle$tokens
l1_metadata <- l1_reference_bundle$metadata

# Extract the frozen target objects

sentence_segment_table <-
  target_input_bundle$sentence_segment_table

paragraph_segment_table <-
  target_input_bundle$paragraph_segment_table

whole_body_text_segment_table <-
  target_input_bundle$whole_body_text_segment_table

shared_prompt_headings <-
  target_input_bundle$shared_prompt_headings

introduced_subheadings <-
  target_input_bundle$introduced_subheadings

paragraph_section_map <-
  target_input_bundle$paragraph_section_map

structural_audit <-
  target_input_bundle$structural_audit

sentence_mapping_audit <-
  target_input_bundle$sentence_mapping_audit

# Validate the frozen inputs

stopifnot(
  isTRUE(l1_metadata$final_audit_passed),
  nrow(l1_documents) == 2004L,
  nrow(l1_sentences) == 5252L,
  nrow(l1_tokens) == 108519L,
  nrow(sentence_segment_table) == 106L,
  nrow(paragraph_segment_table) == 15L,
  nrow(whole_body_text_segment_table) == 4L
)

# Display the validated inventory

frozen_object_audit <- tibble::tibble(
  object = c(
    "L1 documents",
    "L1 source sentences",
    "L1 parsed tokens",
    "Target sentence segments",
    "Target paragraph segments",
    "Target whole-body-text segments",
    "Shared prompt headings",
    "Paragraph-section mappings"
  ),
  n_rows = c(
    nrow(l1_documents),
    nrow(l1_sentences),
    nrow(l1_tokens),
    nrow(sentence_segment_table),
    nrow(paragraph_segment_table),
    nrow(whole_body_text_segment_table),
    nrow(shared_prompt_headings),
    nrow(paragraph_section_map)
  )
)

print(
  frozen_object_audit,
  n = Inf
)

message(
  "Frozen actual LSA objects extracted and validated successfully."
)
```

# MODUL 6.27 — Prepare the selected L1 token layer

```{r}

# MODUL 6.27 — Prepare the selected L1 token layer

required_token_columns <- c(
  "document_uid",
  "sentence_uid",
  "token_uid",
  "lemma",
  "lemma_clean",
  "pos",
  "lsa_selected_final"
)

missing_token_columns <- setdiff(
  required_token_columns,
  names(l1_tokens)
)

if (length(missing_token_columns) > 0L) {
  stop(
    "Missing required token columns: ",
    paste(missing_token_columns, collapse = ", ")
  )
}

clean_lsa_term <- function(x) {
  x |>
    as.character() |>
    stringr::str_squish() |>
    stringr::str_to_lower() |>
    dplyr::na_if("") |>
    dplyr::na_if("_")
}

l1_lsa_tokens <- l1_tokens |>
  dplyr::mutate(
    token_order = dplyr::row_number(),

    lemma_primary = clean_lsa_term(
      lemma
    ),

    lemma_fallback = clean_lsa_term(
      lemma_clean
    ),

    lsa_term = dplyr::coalesce(
      lemma_primary,
      lemma_fallback
    ),

    lsa_pos = stringr::str_to_upper(
      as.character(pos)
    )
  ) |>
  dplyr::filter(
    lsa_selected_final
  )

missing_lsa_terms <- l1_lsa_tokens |>
  dplyr::filter(
    is.na(lsa_term)
  )

if (nrow(missing_lsa_terms) > 0L) {
  print(
    missing_lsa_terms |>
      dplyr::select(
        document_uid,
        sentence_uid,
        token_uid,
        lemma,
        lemma_clean,
        pos
      ),
    n = Inf
  )

  stop(
    nrow(missing_lsa_terms),
    " selected tokens still lack a usable lemma."
  )
}

l1_lsa_tokens <- l1_lsa_tokens |>
  dplyr::select(
    document_uid,
    sentence_uid,
    token_uid,
    token_order,
    lsa_term,
    lsa_pos
  )

stopifnot(
  nrow(l1_lsa_tokens) == 56289L,
  dplyr::n_distinct(
    l1_lsa_tokens$document_uid
  ) == 2004L,
  all(
    l1_lsa_tokens$lsa_pos %in%
      lsa_pos_keep
  ),
  !anyDuplicated(
    l1_lsa_tokens$token_uid
  )
)

l1_lsa_token_audit <- l1_lsa_tokens |>
  dplyr::count(
    lsa_pos,
    name = "n_tokens"
  ) |>
  dplyr::arrange(
    match(
      lsa_pos,
      lsa_pos_keep
    )
  )

print(
  l1_lsa_token_audit,
  n = Inf
)

message(
  "Selected L1 token layer prepared: ",
  format(
    nrow(l1_lsa_tokens),
    big.mark = ","
  ),
  " tokens across ",
  format(
    dplyr::n_distinct(
      l1_lsa_tokens$document_uid
    ),
    big.mark = ","
  ),
  " documents."
)
 

  
    
      

```


```{r}
# MODUL 6.27 — Prepare and freeze the selected L1 token layer

l1_lsa_token_bundle_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_l1_token_layer.rds"
)

if (file.exists(l1_lsa_token_bundle_path)) {

  # Load the previously frozen layer

  l1_lsa_token_bundle <- readRDS(
    l1_lsa_token_bundle_path
  )

  l1_lsa_tokens <-
    l1_lsa_token_bundle$l1_lsa_tokens

  excluded_lsa_tokens <-
    l1_lsa_token_bundle$excluded_lsa_tokens

  l1_lsa_token_audit <-
    l1_lsa_token_bundle$l1_lsa_token_audit

  message(
    "Saved actual-LSA L1 token layer loaded."
  )

} else {

  # Validate the required source columns

  required_token_columns <- c(
    "document_uid",
    "sentence_uid",
    "token_uid",
    "lemma",
    "lemma_clean",
    "pos",
    "lsa_selected_final"
  )

  missing_token_columns <- setdiff(
    required_token_columns,
    names(l1_tokens)
  )

  if (length(missing_token_columns) > 0L) {
    stop(
      "Missing required token columns: ",
      paste(missing_token_columns, collapse = ", ")
    )
  }

  # Clean and select the frozen LSA tokens

  clean_lsa_term <- function(x) {
    x |>
      as.character() |>
      stringr::str_squish() |>
      stringr::str_to_lower() |>
      dplyr::na_if("") |>
      dplyr::na_if("_")
  }

  l1_lsa_tokens_selected <- l1_tokens |>
    dplyr::mutate(
      token_order = dplyr::row_number(),

      lsa_term = dplyr::coalesce(
        clean_lsa_term(lemma),
        clean_lsa_term(lemma_clean)
      ),

      lsa_pos = stringr::str_to_upper(
        stringr::str_squish(
          as.character(pos)
        )
      )
    ) |>
    dplyr::filter(
      lsa_selected_final
    )

  # Preserve but exclude two selected tokens without usable lemmas

  excluded_lsa_tokens <- l1_lsa_tokens_selected |>
    dplyr::filter(
      is.na(lsa_term)
    ) |>
    dplyr::select(
      document_uid,
      sentence_uid,
      token_uid,
      lemma,
      lemma_clean,
      lsa_pos
    )

  l1_lsa_tokens <- l1_lsa_tokens_selected |>
    dplyr::filter(
      !is.na(lsa_term)
    ) |>
    dplyr::select(
      document_uid,
      sentence_uid,
      token_uid,
      token_order,
      lsa_term,
      lsa_pos
    )

  # Validate the final layer

  stopifnot(
    nrow(l1_lsa_tokens_selected) == 56289L,
    nrow(excluded_lsa_tokens) == 2L,
    nrow(l1_lsa_tokens) == 56287L,
    dplyr::n_distinct(
      l1_lsa_tokens$document_uid
    ) == 2004L,
    all(
      l1_lsa_tokens$lsa_pos %in%
        lsa_pos_keep
    ),
    !anyDuplicated(
      l1_lsa_tokens$token_uid
    )
  )

  # Create a compact audit

  l1_lsa_token_audit <- l1_lsa_tokens |>
    dplyr::count(
      lsa_pos,
      name = "n_tokens"
    )

  # Freeze the reusable layer

  l1_lsa_token_bundle <- list(
    l1_lsa_tokens = l1_lsa_tokens,
    excluded_lsa_tokens = excluded_lsa_tokens,
    l1_lsa_token_audit = l1_lsa_token_audit,
    metadata = list(
      n_selected = 56289L,
      n_excluded = 2L,
      n_usable = 56287L,
      n_documents = 2004L,
      lsa_pos_keep = lsa_pos_keep
    )
  )

  saveRDS(
    l1_lsa_token_bundle,
    l1_lsa_token_bundle_path
  )

  message(
    "Actual-LSA L1 token layer prepared and frozen."
  )
}

print(
  l1_lsa_token_audit,
  n = Inf
)

message(
  "Usable actual-LSA tokens: ",
  format(nrow(l1_lsa_tokens), big.mark = ","),
  " across ",
  dplyr::n_distinct(l1_lsa_tokens$document_uid),
  " documents."
)
```
# MODUL 6.28 — Construct and freeze L1 document term sequences

```{r}

# MODUL 6.28 — Construct and freeze L1 document term sequences

l1_document_terms_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_l1_document_terms.rds"
)

if (file.exists(l1_document_terms_path)) {

  l1_document_terms <- readRDS(
    l1_document_terms_path
  )

  message(
    "Saved L1 document term sequences loaded."
  )

} else {

  l1_document_terms <- l1_lsa_tokens |>
    dplyr::arrange(
      document_uid,
      token_order
    ) |>
    dplyr::group_by(
      document_uid
    ) |>
    dplyr::summarise(
      lsa_text = stringr::str_c(
        lsa_term,
        collapse = " "
      ),
      n_lsa_tokens = dplyr::n(),
      n_lsa_types = dplyr::n_distinct(
        lsa_term
      ),
      .groups = "drop"
    )

  stopifnot(
    nrow(l1_document_terms) == 2004L,
    sum(l1_document_terms$n_lsa_tokens) == 56287L,
    !anyDuplicated(
      l1_document_terms$document_uid
    ),
    all(
      !is.na(l1_document_terms$lsa_text) &
        l1_document_terms$lsa_text != ""
    )
  )

  saveRDS(
    l1_document_terms,
    l1_document_terms_path
  )

  message(
    "L1 document term sequences constructed and frozen."
  )
}

stopifnot(
  nrow(l1_document_terms) == 2004L,
  sum(l1_document_terms$n_lsa_tokens) == 56287L
)

print(
  l1_document_terms |>
    dplyr::summarise(
      n_documents = dplyr::n(),
      n_tokens = sum(n_lsa_tokens),
      median_tokens = stats::median(
        n_lsa_tokens
      ),
      min_tokens = min(
        n_lsa_tokens
      ),
      max_tokens = max(
        n_lsa_tokens
      )
    )
)
```
# # MODUL 6.29 — Build and freeze the L1 document-feature matrix

```{r}
# MODUL 6.29 — Build and freeze the L1 document-feature matrix

l1_dfm_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_l1_dfm.rds"
)

if (file.exists(l1_dfm_path)) {

  l1_dfm <- readRDS(
    l1_dfm_path
  )

  message(
    "Saved L1 document-feature matrix loaded."
  )

} else {

  l1_tokens_quanteda <- quanteda::tokens(
    l1_document_terms$lsa_text,
    what = "fastestword"
  )

  quanteda::docnames(
    l1_tokens_quanteda
  ) <- l1_document_terms$document_uid

  l1_dfm <- quanteda::dfm(
    l1_tokens_quanteda
  )

  stopifnot(
    quanteda::ndoc(l1_dfm) == 2004L,
    sum(l1_dfm) == 56287L,
    quanteda::nfeat(l1_dfm) > lsa_dimensions,
    identical(
      quanteda::docnames(l1_dfm),
      l1_document_terms$document_uid
    )
  )

  saveRDS(
    l1_dfm,
    l1_dfm_path
  )

  message(
    "L1 document-feature matrix built and frozen."
  )
}

l1_dfm_audit <- tibble::tibble(
  n_documents = quanteda::ndoc(l1_dfm),
  n_features = quanteda::nfeat(l1_dfm),
  n_tokens = sum(l1_dfm),
  sparsity = quanteda::sparsity(l1_dfm)
)

print(
  l1_dfm_audit,
  n = Inf
)
```
# MODUL 6.30 — Calculate and freeze L1 document frequencies and IDF

```{r}

# MODUL 6.30 — Calculate and freeze L1 document frequencies and IDF

l1_idf_bundle_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_l1_idf.rds"
)

if (file.exists(l1_idf_bundle_path)) {

  l1_idf_bundle <- readRDS(
    l1_idf_bundle_path
  )

  l1_vocabulary <- l1_idf_bundle$vocabulary
  l1_document_frequency <- l1_idf_bundle$document_frequency
  l1_idf <- l1_idf_bundle$idf

  message(
    "Saved L1 vocabulary and IDF loaded."
  )

} else {

  n_l1_documents <- quanteda::ndoc(
    l1_dfm
  )

  l1_vocabulary <- quanteda::featnames(
    l1_dfm
  )

  l1_document_frequency <- as.numeric(
    Matrix::colSums(
      l1_dfm > 0
    )
  )

  names(l1_document_frequency) <-
    l1_vocabulary

  l1_idf <- log(
    n_l1_documents /
      l1_document_frequency
  )

  names(l1_idf) <- l1_vocabulary

  stopifnot(
    length(l1_vocabulary) == 9109L,
    length(l1_document_frequency) == 9109L,
    length(l1_idf) == 9109L,
    all(l1_document_frequency >= 1L),
    all(l1_document_frequency <= n_l1_documents),
    all(is.finite(l1_idf)),
    all(l1_idf >= 0),
    identical(
      names(l1_idf),
      l1_vocabulary
    )
  )

  l1_idf_bundle <- list(
    vocabulary = l1_vocabulary,
    document_frequency = l1_document_frequency,
    idf = l1_idf,
    n_documents = n_l1_documents,
    formula = "log(N / df)"
  )

  saveRDS(
    l1_idf_bundle,
    l1_idf_bundle_path
  )

  message(
    "L1 vocabulary and IDF calculated and frozen."
  )
}

l1_idf_audit <- tibble::tibble(
  n_documents = l1_idf_bundle$n_documents,
  n_features = length(l1_vocabulary),
  min_df = min(l1_document_frequency),
  max_df = max(l1_document_frequency),
  min_idf = min(l1_idf),
  median_idf = stats::median(l1_idf),
  max_idf = max(l1_idf)
)

print(
  l1_idf_audit,
  n = Inf
)
n_single_document_features = sum(
  l1_document_frequency == 1L
)
```


```{r}
tibble::tibble(n_single_document_features = sum(l1_document_frequency == 1L), proportion_single_document_features = mean(l1_document_frequency == 1L))
```

# MODUL 6.31 — Apply and freeze log-TF–IDF weighting

```{r}

# MODUL 6.31 — Apply and freeze log-TF–IDF weighting

l1_tfidf_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_l1_tfidf.rds"
)

if (file.exists(l1_tfidf_path)) {

  l1_tfidf_matrix <- readRDS(
    l1_tfidf_path
  )

  message(
    "Saved L1 log-TF–IDF matrix loaded."
  )

} else {

  # Convert the DFM to a sparse numeric matrix

  l1_count_matrix <- methods::as(
    l1_dfm,
    "dgCMatrix"
  )

  # Validate vocabulary alignment

  stopifnot(
    identical(
      colnames(l1_count_matrix),
      l1_vocabulary
    ),
    identical(
      names(l1_idf),
      l1_vocabulary
    )
  )

  # Apply TF = log(1 + tf) to nonzero entries

  l1_log_tf_matrix <- l1_count_matrix
  l1_log_tf_matrix@x <- log1p(
    l1_log_tf_matrix@x
  )

  # Apply IDF = log(N / df)

  l1_tfidf_matrix <- l1_log_tf_matrix %*%
    Matrix::Diagonal(
      x = as.numeric(l1_idf)
    )

  colnames(l1_tfidf_matrix) <- l1_vocabulary
  rownames(l1_tfidf_matrix) <- quanteda::docnames(
    l1_dfm
  )

  # Validate the weighted matrix

  l1_tfidf_row_norms <- sqrt(
    Matrix::rowSums(
      l1_tfidf_matrix ^ 2
    )
  )

  stopifnot(
    all(
      dim(l1_tfidf_matrix) ==
        c(2004L, 9109L)
    ),
    all(is.finite(l1_tfidf_matrix@x)),
    all(l1_tfidf_matrix@x > 0),
    all(l1_tfidf_row_norms > 0)
  )

  saveRDS(
    l1_tfidf_matrix,
    l1_tfidf_path
  )

  message(
    "L1 log-TF–IDF matrix calculated and frozen."
  )
}

# Audit the weighted matrix

l1_tfidf_row_norms <- sqrt(
  Matrix::rowSums(
    l1_tfidf_matrix ^ 2
  )
)

l1_tfidf_audit <- tibble::tibble(
  n_documents = nrow(l1_tfidf_matrix),
  n_features = ncol(l1_tfidf_matrix),
  n_nonzero = length(l1_tfidf_matrix@x),
  min_row_norm = min(l1_tfidf_row_norms),
  median_row_norm = stats::median(
    l1_tfidf_row_norms
  ),
  max_row_norm = max(l1_tfidf_row_norms),
  n_zero_rows = sum(
    l1_tfidf_row_norms == 0
  )
)

print(
  l1_tfidf_audit,
  n = Inf
)
```

# MODUL 6.32 — Fit and freeze the fixed 50-dimensional SVD

```{r}
# MODUL 6.32 — Fit and freeze the fixed 50-dimensional SVD

l1_svd_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_svd.rds"
)

if (file.exists(l1_svd_path)) {

  l1_svd <- readRDS(
    l1_svd_path
  )

  message(
    "Saved 50-dimensional LSA decomposition loaded."
  )

} else {

  stopifnot(
    lsa_dimensions == 50L,
    lsa_dimensions < min(
      dim(l1_tfidf_matrix)
    )
  )

  set.seed(
    lsa_random_seed
  )

  l1_svd <- irlba::irlba(
    l1_tfidf_matrix,
    nv = lsa_dimensions,
    nu = lsa_dimensions
  )

  stopifnot(
    ncol(l1_svd$u) == 50L,
    length(l1_svd$d) == 50L,
    ncol(l1_svd$v) == 50L,
    all(is.finite(l1_svd$d)),
    all(l1_svd$d > 0)
  )

  saveRDS(
    l1_svd,
    l1_svd_path
  )

  message(
    "Fixed 50-dimensional LSA decomposition fitted and frozen."
  )
}

lsa_retained_energy <- sum(
  l1_svd$d ^ 2
) / sum(
  l1_tfidf_matrix@x ^ 2
)

l1_svd_audit <- tibble::tibble(
  n_documents = nrow(l1_svd$u),
  n_features = nrow(l1_svd$v),
  n_dimensions = length(l1_svd$d),
  largest_singular_value = max(l1_svd$d),
  smallest_retained_singular_value = min(l1_svd$d),
  retained_matrix_energy = lsa_retained_energy
)

print(
  l1_svd_audit,
  n = Inf
)
```
# MODUL 6.33  loads the frozen LSA reference space. It does not reconstruct these target segments.

```{r}
# MODUL 6.33 — Freeze and validate the LSA reference space

lsa_reference_space_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_reference_space.rds"
)

if (file.exists(lsa_reference_space_path)) {

  lsa_reference_space <- readRDS(
    lsa_reference_space_path
  )

  message(
    "Saved actual-LSA reference space loaded."
  )

} else {

  dimension_names <- sprintf(
    "dim_%02d",
    seq_len(lsa_dimensions)
  )

  right_singular_vectors <- l1_svd$v
  rownames(right_singular_vectors) <- l1_vocabulary
  colnames(right_singular_vectors) <- dimension_names

  reference_document_vectors <- l1_svd$u
  rownames(reference_document_vectors) <- rownames(
    l1_tfidf_matrix
  )
  colnames(reference_document_vectors) <- dimension_names

  singular_values <- l1_svd$d
  names(singular_values) <- dimension_names

  stopifnot(
    identical(
      rownames(right_singular_vectors),
      l1_vocabulary
    ),
    identical(
      names(l1_idf),
      l1_vocabulary
    ),
    nrow(right_singular_vectors) == 9109L,
    ncol(right_singular_vectors) == 50L,
    nrow(reference_document_vectors) == 2004L,
    ncol(reference_document_vectors) == 50L,
    all(singular_values > 0)
  )

  lsa_reference_space <- list(
    vocabulary = l1_vocabulary,
    document_frequency = l1_document_frequency,
    idf = l1_idf,
    singular_values = singular_values,
    right_singular_vectors = right_singular_vectors,
    reference_document_vectors = reference_document_vectors,

    configuration = list(
      dimensions = lsa_dimensions,
      tf_weighting = "log(1 + tf)",
      idf_weighting = "log(N / df)",
      fold_in = "weighted_segment %*% V %*% D^-1",
      lsa_pos_keep = lsa_pos_keep,
      random_seed = lsa_random_seed
    ),

    audit = l1_svd_audit
  )

  saveRDS(
    lsa_reference_space,
    lsa_reference_space_path
  )

  message(
    "Actual-LSA reference space frozen successfully."
  )
}

stopifnot(
  length(lsa_reference_space$vocabulary) == 9109L,
  length(lsa_reference_space$idf) == 9109L,
  length(lsa_reference_space$singular_values) == 50L,
  all(
    dim(
      lsa_reference_space$right_singular_vectors
    ) == c(9109L, 50L)
  )
)

tibble::tibble(
  n_documents = nrow(
    lsa_reference_space$reference_document_vectors
  ),
  n_features = length(
    lsa_reference_space$vocabulary
  ),
  n_dimensions = length(
    lsa_reference_space$singular_values
  ),
  retained_matrix_energy =
    lsa_reference_space$audit$retained_matrix_energy
)
```

# MODUL 6.34 — Assemble and freeze all target and heading segments

```{r}

# MODUL 6.34 — Assemble and freeze all target and heading segments

target_segments_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_target_segments.rds"
)

segment_bundle_version <-
  "M6.34_heading_lemma_fix_2026-07-14_v3"

# Define the selected lemma representation by stable heading ID

expected_heading_lsa <- tibble::tibble(
  heading_id = c(
    "main_title",
    "subheading_1",
    "subheading_2",
    "subheading_3"
  ),
  section_id = c(
    NA_integer_,
    1L,
    2L,
    3L
  ),
  lsa_input_text = c(
    "digitalisierung arbeitswelt",
    "digitalisierung sorge arbeitsplatz",
    "fortschritt erfordern anpassung",
    "menschlich qualität erfolgreich zukunft"
  )
)

downstream_segment_paths <- c(
  file.path(
    actual_lsa_output_dir,
    "modul_6_actual_lsa_target_projections.rds"
  ),
  file.path(
    actual_lsa_output_dir,
    "modul_6_actual_lsa_target_coverage.rds"
  )
)

# Check whether the saved segment checkpoint is current

load_saved_segments <- FALSE
saved_segment_bundle <- NULL

if (file.exists(target_segments_path)) {

  saved_segment_bundle <- tryCatch(
    readRDS(target_segments_path),
    error = function(e) NULL
  )

  if (
    is.list(saved_segment_bundle) &&
    all(
      c(
        "version",
        "body_segments",
        "heading_segments",
        "all_target_segments",
        "target_segment_audit"
      ) %in% names(saved_segment_bundle)
    )
  ) {

    saved_heading_check <-
      saved_segment_bundle$heading_segments |>
      dplyr::arrange(segment_order) |>
      dplyr::transmute(
        heading_id = governing_heading_id,
        section_id,
        lsa_input_text
      )

    load_saved_segments <-
      identical(
        saved_segment_bundle$version,
        segment_bundle_version
      ) &&
      identical(
        saved_heading_check,
        expected_heading_lsa
      ) &&
      nrow(
        saved_segment_bundle$all_target_segments
      ) == 129L
  }
}

if (load_saved_segments) {

  target_segment_bundle <- saved_segment_bundle

  body_segments <-
    target_segment_bundle$body_segments

  heading_segments <-
    target_segment_bundle$heading_segments

  all_target_segments <-
    target_segment_bundle$all_target_segments

  target_segment_audit <-
    target_segment_bundle$target_segment_audit

  message(
    "Current target and heading segment checkpoint loaded."
  )

} else {

  # Assemble the 125 heading-free body segments

  body_segments <- dplyr::bind_rows(

    sentence_segment_table |>
      dplyr::transmute(
        text_id,
        model,
        segment_id,
        segment_level = "sentence",
        segment_order = sentence_number,
        paragraph_number,
        section_id,
        governing_heading_id = subheading_id,
        segment_text = body_token_text,
        lsa_input_text = lsa_lemma_text,
        n_input_terms = n_lsa_tokens
      ),

    paragraph_segment_table |>
      dplyr::transmute(
        text_id,
        model,
        segment_id,
        segment_level = "paragraph",
        segment_order = paragraph_number,
        paragraph_number,
        section_id,
        governing_heading_id = subheading_id,
        segment_text = body_token_text,
        lsa_input_text = lsa_lemma_text,
        n_input_terms = n_lsa_tokens
      ),

    whole_body_text_segment_table |>
      dplyr::transmute(
        text_id,
        model,
        segment_id,
        segment_level = "whole_body_text",
        segment_order = 1L,
        paragraph_number = NA_integer_,
        section_id = NA_integer_,
        governing_heading_id = NA_character_,
        segment_text = body_token_text,
        lsa_input_text = lsa_lemma_text,
        n_input_terms = n_lsa_tokens
      )
  )

  # Assemble the four headings by stable heading ID

  heading_segments <- shared_prompt_headings |>
    dplyr::select(
      heading_order,
      heading_id,
      heading_type,
      heading_text
    ) |>
    dplyr::left_join(
      expected_heading_lsa,
      by = "heading_id"
    ) |>
    dplyr::arrange(heading_order) |>
    dplyr::transmute(
      text_id = "SHARED_PROMPT",
      model = "shared_prompt",
      segment_id = paste0(
        "PROMPT_",
        heading_id
      ),
      segment_level = dplyr::if_else(
        heading_type == "main_title",
        "main_title",
        "subheading"
      ),
      segment_order = as.integer(
        heading_order
      ),
      paragraph_number = NA_integer_,
      section_id,
      governing_heading_id = heading_id,
      segment_text = heading_text,
      lsa_input_text,
      n_input_terms = stringr::str_count(
        lsa_input_text,
        "\\S+"
      )
    )

  # Combine all 129 directly projected segments

  all_target_segments <- dplyr::bind_rows(
    body_segments,
    heading_segments
  )

  target_segment_audit <- all_target_segments |>
    dplyr::count(
      segment_level,
      name = "n_segments"
    ) |>
    dplyr::arrange(segment_level)

  # Validate the corrected segment bundle

  heading_check <- heading_segments |>
    dplyr::transmute(
      heading_id = governing_heading_id,
      section_id,
      lsa_input_text
    )

  stopifnot(
    nrow(body_segments) == 125L,
    nrow(heading_segments) == 4L,
    nrow(all_target_segments) == 129L,
    !anyDuplicated(
      all_target_segments$segment_id
    ),
    !anyNA(
      all_target_segments$lsa_input_text
    ),
    all(
      all_target_segments$lsa_input_text != ""
    ),
    identical(
      heading_check,
      expected_heading_lsa
    )
  )

  # Freeze the corrected segment bundle

  target_segment_bundle <- list(
    version = segment_bundle_version,
    body_segments = body_segments,
    heading_segments = heading_segments,
    all_target_segments = all_target_segments,
    target_segment_audit = target_segment_audit
  )

  saveRDS(
    target_segment_bundle,
    target_segments_path
  )

  # Remove downstream checkpoints based on older segment data

  unlink(
    downstream_segment_paths[
      file.exists(downstream_segment_paths)
    ]
  )

  message(
    paste0(
      "Corrected target and heading segments frozen; ",
      "dependent projection and coverage checkpoints removed."
    )
  )
}

print(
  target_segment_audit,
  n = Inf
)

print(
  heading_segments |>
    dplyr::select(
      segment_level,
      segment_order,
      governing_heading_id,
      segment_text,
      lsa_input_text,
      n_input_terms
    ),
  n = Inf,
  width = Inf
)
        
```

```{r}
heading_segments |>
  dplyr::select(
    governing_heading_id,
    lsa_input_text,
    n_input_terms
  )
```
#

```{r}
shared_prompt_headings |>
  dplyr::select(
    heading_order,
    heading_id,
    heading_type,
    heading_text
  ) |>
  print(n = Inf, width = Inf)
```
# ODUL 6.35 — Fold target segments into the fixed LSA space - uses the frozen 129 segments, the frozen vocabulary, IDF, right-singular vectors, and singular values. It does not refit the LSA space.

```{r}

# MODUL 6.35 — Fold target segments into the fixed LSA space
# MODUL 6.35 — Fold target segments into the fixed LSA space

target_projection_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_target_projections.rds"
)

if (file.exists(target_projection_path)) {

  target_projection_bundle <- readRDS(
    target_projection_path
  )

  target_dfm_full <-
    target_projection_bundle$target_dfm_full

  target_dfm <-
    target_projection_bundle$target_dfm

  target_tfidf_matrix <-
    target_projection_bundle$target_tfidf_matrix

  target_segment_vectors_raw <-
    target_projection_bundle$segment_vectors_raw

  target_segment_vectors <-
    target_projection_bundle$segment_vectors

  target_vector_norms <-
    target_projection_bundle$vector_norms

  target_vector_valid <-
    target_projection_bundle$valid_vector

  message(
    "Saved target-segment projections loaded."
  )

} else {

  # Build one DFM for all frozen target and heading segments

  target_tokens <- quanteda::tokens(
    stats::setNames(
      all_target_segments$lsa_input_text,
      all_target_segments$segment_id
    ),
    remove_punct = TRUE,
    remove_symbols = TRUE
  )

  target_dfm_full <- quanteda::dfm(
    target_tokens,
    tolower = FALSE
  )

  # Match the target DFM to the frozen L1 vocabulary

  target_dfm <- quanteda::dfm_match(
    target_dfm_full,
    features = lsa_reference_space$vocabulary
  )

  target_count_matrix <- methods::as(
    target_dfm,
    "dgCMatrix"
  )

  stopifnot(
    identical(
      colnames(target_count_matrix),
      lsa_reference_space$vocabulary
    ),
    identical(
      names(lsa_reference_space$idf),
      lsa_reference_space$vocabulary
    ),
    nrow(target_count_matrix) == 129L
  )

  # Apply the frozen log-TF weighting

  target_log_tf_matrix <- target_count_matrix

  target_log_tf_matrix@x <- log1p(
    target_log_tf_matrix@x
  )

  # Apply the frozen L1-derived IDF weights

  target_tfidf_matrix <-
    target_log_tf_matrix %*%
    Matrix::Diagonal(
      x = as.numeric(
        lsa_reference_space$idf
      )
    )

  # Construct the fixed fold-in basis

  fold_in_basis <-
    lsa_reference_space$right_singular_vectors %*%
    Matrix::Diagonal(
      x = 1 /
        lsa_reference_space$singular_values
    )

  # Fold all segments into the fixed 50-dimensional space

  target_segment_vectors_raw <- as.matrix(
    target_tfidf_matrix %*%
      fold_in_basis
  )

  rownames(target_segment_vectors_raw) <-
    all_target_segments$segment_id

  colnames(target_segment_vectors_raw) <-
    colnames(
      lsa_reference_space$right_singular_vectors
    )

  # Calculate vector norms and identify valid vectors

  target_vector_norms <- sqrt(
    rowSums(
      target_segment_vectors_raw ^ 2
    )
  )

  vector_tolerance <- sqrt(
    .Machine$double.eps
  )

  target_vector_valid <-
    is.finite(target_vector_norms) &
    target_vector_norms > vector_tolerance

  stopifnot(
    any(target_vector_valid)
  )

  # L2-normalize valid projected vectors

  target_segment_vectors <- matrix(
    NA_real_,
    nrow = nrow(target_segment_vectors_raw),
    ncol = ncol(target_segment_vectors_raw),
    dimnames = dimnames(
      target_segment_vectors_raw
    )
  )

  target_segment_vectors[
    target_vector_valid,
  ] <- sweep(
    target_segment_vectors_raw[
      target_vector_valid,
      ,
      drop = FALSE
    ],
    MARGIN = 1,
    STATS = target_vector_norms[
      target_vector_valid
    ],
    FUN = "/"
  )

  # Validate the projected space

  stopifnot(
    nrow(target_segment_vectors) == 129L,
    ncol(target_segment_vectors) == 50L,
    all(
      is.finite(
        target_segment_vectors[
          target_vector_valid,
          ,
          drop = FALSE
        ]
      )
    ),
    all(
      abs(
        sqrt(
          rowSums(
            target_segment_vectors[
              target_vector_valid,
              ,
              drop = FALSE
            ] ^ 2
          )
        ) - 1
      ) < 1e-8
    )
  )

  # Freeze the complete projection bundle

  target_projection_bundle <- list(
    segment_metadata = all_target_segments,
    target_dfm_full = target_dfm_full,
    target_dfm = target_dfm,
    target_tfidf_matrix = target_tfidf_matrix,
    segment_vectors_raw = target_segment_vectors_raw,
    segment_vectors = target_segment_vectors,
    vector_norms = target_vector_norms,
    valid_vector = target_vector_valid
  )

  saveRDS(
    target_projection_bundle,
    target_projection_path
  )

  message(
    paste0(
      "Target segments folded into the fixed actual-LSA ",
      "space and frozen."
    )
  )
}

# Summarize the projection result

target_projection_audit <- tibble::tibble(
  n_segments = nrow(
    target_segment_vectors
  ),
  n_dimensions = ncol(
    target_segment_vectors
  ),
  n_valid_vectors = sum(
    target_vector_valid
  ),
  n_invalid_vectors = sum(
    !target_vector_valid
  )
)

print(
  target_projection_audit,
  n = Inf
)

```
#  MODUL 6.36 — Audit vocabulary coverage and vector validity

```{r}

# MODUL 6.36 — Audit vocabulary coverage and vector validity

target_coverage_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_target_coverage.rds"
)

if (file.exists(target_coverage_path)) {

  target_coverage_bundle <- readRDS(
    target_coverage_path
  )

  target_segment_audit <-
    target_coverage_bundle$segment_audit

  target_coverage_summary <-
    target_coverage_bundle$coverage_summary

  message(
    "Saved target vocabulary-coverage audit loaded."
  )

} else {

  # Calculate token and type counts before and after vocabulary matching

  n_input_lemmas <- as.numeric(
    Matrix::rowSums(target_dfm_full)
  )

  n_in_vocabulary <- as.numeric(
    Matrix::rowSums(target_dfm)
  )

  n_input_types <- as.numeric(
    Matrix::rowSums(target_dfm_full > 0)
  )

  n_in_vocabulary_types <- as.numeric(
    Matrix::rowSums(target_dfm > 0)
  )

  # Assemble the complete segment-level audit

  target_segment_audit <- all_target_segments |>
    dplyr::mutate(
      n_input_lemmas = n_input_lemmas,
      n_in_vocabulary = n_in_vocabulary,
      n_oov = n_input_lemmas - n_in_vocabulary,
      token_coverage = dplyr::if_else(
        n_input_lemmas > 0,
        n_in_vocabulary / n_input_lemmas,
        NA_real_
      ),
      n_input_types = n_input_types,
      n_in_vocabulary_types = n_in_vocabulary_types,
      type_coverage = dplyr::if_else(
        n_input_types > 0,
        n_in_vocabulary_types / n_input_types,
        NA_real_
      ),
      vector_norm = target_vector_norms,
      valid_vector = target_vector_valid
    )

  # Create a compact audit by segment level

  target_coverage_summary <- target_segment_audit |>
    dplyr::group_by(segment_level) |>
    dplyr::summarise(
      n_segments = dplyr::n(),
      lsa_coverage_mean = mean(
        token_coverage,
        na.rm = TRUE
      ),
      lsa_coverage_min = min(
        token_coverage,
        na.rm = TRUE
      ),
      n_invalid_segments = sum(
        !valid_vector
      ),
      .groups = "drop"
    )

  stopifnot(
    nrow(target_segment_audit) == 129L,
    all(target_segment_audit$n_oov >= 0),
    all(
      target_segment_audit$token_coverage >= 0 &
        target_segment_audit$token_coverage <= 1
    ),
    all(
      target_segment_audit$type_coverage >= 0 &
        target_segment_audit$type_coverage <= 1
    ),
    sum(!target_segment_audit$valid_vector) == 0L
  )

  target_coverage_bundle <- list(
    segment_audit = target_segment_audit,
    coverage_summary = target_coverage_summary
  )

  saveRDS(
    target_coverage_bundle,
    target_coverage_path
  )

  message(
    "Target vocabulary coverage and vector validity audited and frozen."
  )
}

print(
  target_coverage_summary,
  n = Inf
)
```

```{r}
# MODUL 6.37 — Calculate and freeze the x3 similarities

x3_output_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_x3.rds"
)

if (file.exists(x3_output_path)) {

  x3_bundle <- readRDS(x3_output_path)

  x3_sentence_pairs <- x3_bundle$sentence_pairs
  x3_paragraph_pairs <- x3_bundle$paragraph_pairs
  x3_profiles <- x3_bundle$profiles

  message("Saved x3 similarities loaded.")

} else {

  # Calculate cosine similarity from the normalized segment vectors

  segment_cosine <- function(segment_id_1, segment_id_2) {

    vector_1 <- target_segment_vectors[
      segment_id_1,
      ,
      drop = TRUE
    ]

    vector_2 <- target_segment_vectors[
      segment_id_2,
      ,
      drop = TRUE
    ]

    if (
      anyNA(vector_1) ||
      anyNA(vector_2)
    ) {
      return(NA_real_)
    }

    sum(vector_1 * vector_2)
  }

  # Calculate adjacent-sentence similarities

  x3_sentence_pairs <- all_target_segments |>
    dplyr::filter(
      segment_level == "sentence"
    ) |>
    dplyr::arrange(
      text_id,
      segment_order
    ) |>
    dplyr::group_by(
      text_id,
      model
    ) |>
    dplyr::mutate(
      next_segment_id = dplyr::lead(segment_id),
      next_segment_order = dplyr::lead(segment_order)
    ) |>
    dplyr::filter(
      !is.na(next_segment_id)
    ) |>
    dplyr::ungroup() |>
    dplyr::mutate(
      cosine_similarity = purrr::map2_dbl(
        segment_id,
        next_segment_id,
        segment_cosine
      )
    ) |>
    dplyr::select(
      text_id,
      model,
      segment_id_1 = segment_id,
      segment_id_2 = next_segment_id,
      sentence_order_1 = segment_order,
      sentence_order_2 = next_segment_order,
      cosine_similarity
    )

  # Calculate adjacent-paragraph similarities

  x3_paragraph_pairs <- all_target_segments |>
    dplyr::filter(
      segment_level == "paragraph"
    ) |>
    dplyr::arrange(
      text_id,
      segment_order
    ) |>
    dplyr::group_by(
      text_id,
      model
    ) |>
    dplyr::mutate(
      next_segment_id = dplyr::lead(segment_id),
      next_segment_order = dplyr::lead(segment_order)
    ) |>
    dplyr::filter(
      !is.na(next_segment_id)
    ) |>
    dplyr::ungroup() |>
    dplyr::mutate(
      cosine_similarity = purrr::map2_dbl(
        segment_id,
        next_segment_id,
        segment_cosine
      )
    ) |>
    dplyr::select(
      text_id,
      model,
      segment_id_1 = segment_id,
      segment_id_2 = next_segment_id,
      paragraph_order_1 = segment_order,
      paragraph_order_2 = next_segment_order,
      cosine_similarity
    )

  # Assemble the text-level x3 profile

  x3_sentence_summary <- x3_sentence_pairs |>
    dplyr::group_by(
      text_id,
      model
    ) |>
    dplyr::summarise(
      x3_adjacent_sentence_similarity_mean = mean(
        cosine_similarity,
        na.rm = TRUE
      ),
      x3_adjacent_sentence_similarity_sd = stats::sd(
        cosine_similarity,
        na.rm = TRUE
      ),
      n_adjacent_sentence_pairs = dplyr::n(),
      .groups = "drop"
    )

  x3_paragraph_summary <- x3_paragraph_pairs |>
    dplyr::group_by(
      text_id,
      model
    ) |>
    dplyr::summarise(
      x3_adjacent_paragraph_similarity_mean = mean(
        cosine_similarity,
        na.rm = TRUE
      ),
      n_adjacent_paragraph_pairs = dplyr::n(),
      .groups = "drop"
    )

  x3_profiles <- x3_sentence_summary |>
    dplyr::left_join(
      x3_paragraph_summary,
      by = c(
        "text_id",
        "model"
      )
    )

  stopifnot(
    nrow(x3_profiles) == 4L,
    all(
      is.finite(
        x3_profiles$x3_adjacent_sentence_similarity_mean
      )
    ),
    all(
      is.finite(
        x3_profiles$x3_adjacent_paragraph_similarity_mean
      )
    ),
    all(
      x3_sentence_pairs$cosine_similarity >= -1 &
        x3_sentence_pairs$cosine_similarity <= 1
    ),
    all(
      x3_paragraph_pairs$cosine_similarity >= -1 &
        x3_paragraph_pairs$cosine_similarity <= 1
    )
  )

  x3_bundle <- list(
    sentence_pairs = x3_sentence_pairs,
    paragraph_pairs = x3_paragraph_pairs,
    profiles = x3_profiles
  )

  saveRDS(
    x3_bundle,
    x3_output_path
  )

  message("x3 similarities calculated and frozen.")
}

print(
  x3_profiles,
  n = Inf,
  width = Inf
)
```
# Calculates the within-text gradient - 

```{r}
# Calculate within-text semantic-continuity gradients

x3_sentence_gradient <- x3_sentence_pairs |>
  dplyr::group_by(
    text_id,
    model
  ) |>
  dplyr::arrange(
    sentence_order_1,
    .by_group = TRUE
  ) |>
  dplyr::mutate(
    normalized_text_position =
      (dplyr::row_number() - 1) /
      pmax(dplyr::n() - 1, 1)
  ) |>
  dplyr::summarise(
    x3_adjacent_sentence_similarity_gradient =
      if (dplyr::n() > 1L) {
        unname(
          stats::coef(
            stats::lm(
              cosine_similarity ~ normalized_text_position
            )
          )[["normalized_text_position"]]
        )
      } else {
        NA_real_
      },

    x3_adjacent_sentence_similarity_first_third =
      mean(
        cosine_similarity[
          normalized_text_position <= 1 / 3
        ],
        na.rm = TRUE
      ),

    x3_adjacent_sentence_similarity_final_third =
      mean(
        cosine_similarity[
          normalized_text_position >= 2 / 3
        ],
        na.rm = TRUE
      ),

    x3_adjacent_sentence_similarity_retention =
      x3_adjacent_sentence_similarity_final_third -
      x3_adjacent_sentence_similarity_first_third,

    .groups = "drop"
  )

print(
  x3_sentence_gradient,
  n = Inf,
  width = Inf
)
```
#

```{r}
# MODUL 6.37a — Calculate and freeze x3 sentence-gradient diagnostics

x3_gradient_diagnostics <- x3_sentence_pairs |>
  dplyr::group_by(text_id, model) |>
  dplyr::mutate(normalized_position = (dplyr::row_number() - 1) / (dplyr::n() - 1)) |>
  dplyr::group_modify(~ {
    fit <- stats::lm(cosine_similarity ~ normalized_position, data = .x)
    thirds <- cut(.x$normalized_position, c(-Inf, 1/3, 2/3, Inf), labels = c("first", "middle", "final"))
    means <- tapply(.x$cosine_similarity, thirds, mean)
    tibble::tibble(
      x3_adjacent_sentence_similarity_gradient = unname(stats::coef(fit)[2]),
      x3_adjacent_sentence_gradient_se = summary(fit)$coefficients[2, 2],
      x3_adjacent_sentence_gradient_r2 = summary(fit)$r.squared,
      x3_adjacent_sentence_similarity_first_third = unname(means["first"]),
      x3_adjacent_sentence_similarity_middle_third = unname(means["middle"]),
      x3_adjacent_sentence_similarity_final_third = unname(means["final"]),
      x3_adjacent_sentence_similarity_retention = unname(means["final"] - means["first"])
    )
  }) |>
  dplyr::ungroup()

x3_profiles_extended <- x3_profiles |>
  dplyr::left_join(x3_gradient_diagnostics, by = c("text_id", "model"))

saveRDS(list(profiles = x3_profiles_extended, diagnostics = x3_gradient_diagnostics),
        file.path(actual_lsa_output_dir, "modul_6_actual_lsa_x3_gradient.rds"))

print(x3_profiles_extended, n = Inf, width = Inf)
print(x3_gradient_diagnostics, n = Inf, width = Inf)
View(x3_gradient_diagnostics)

readr::write_csv(x3_gradient_diagnostics,
                 file.path(actual_lsa_output_dir, "modul_6_actual_lsa_x3_gradient_diagnostics.csv"))
```

#

```{r}
print(x3_profiles_extended, n = Inf, width = Inf)
View(x3_profiles_extended)

readr::write_csv(x3_profiles_extended,
                 file.path(actual_lsa_output_dir, "modul_6_actual_lsa_x3_profiles_extended.csv"))
```


# Plot save: modul_6_x3_sentence_gradient

```{r}
# MODUL 6.37b — Plot x3 sentence-gradient diagnostics

x3_plot_data <- x3_sentence_pairs |>
  dplyr::group_by(text_id, model) |> dplyr::mutate(normalized_position = (dplyr::row_number() - 1) / (dplyr::n() - 1)) |> dplyr::ungroup()

x3_plot_labels <- x3_gradient_diagnostics |>
  dplyr::mutate(label = sprintf("Slope = %.3f\nSE = %.3f | R² = %.3f\nFirst = %.3f | Middle = %.3f | Final = %.3f\nRetention = %.3f",
                                x3_adjacent_sentence_similarity_gradient, x3_adjacent_sentence_gradient_se,
                                x3_adjacent_sentence_gradient_r2, x3_adjacent_sentence_similarity_first_third,
                                x3_adjacent_sentence_similarity_middle_third, x3_adjacent_sentence_similarity_final_third,
                                x3_adjacent_sentence_similarity_retention))

modul_6_x3_sentence_gradient_diagnostics <- ggplot2::ggplot(x3_plot_data, ggplot2::aes(normalized_position, cosine_similarity)) +
  ggplot2::geom_point(size = 0.5) + ggplot2::geom_smooth(method = "lm", se = FALSE, color = "black", linewidth = 0.4) +
  ggplot2::geom_text(data = x3_plot_labels, ggplot2::aes(x = 0, y = Inf, label = label), hjust = 0, vjust = 1.1, size = 2.5, inherit.aes = FALSE) +
  ggplot2::facet_wrap(~ model) + ggplot2::labs(title = "MODUL 6 x3 Sentence-Gradient Diagnostics", x = "Normalized sentence-pair position", y = "Adjacent-sentence similarity")

modul_6_x3_sentence_gradient_diagnostics
ggplot2::ggsave(file.path(actual_lsa_output_dir, "modul_6_x3_sentence_gradient_diagnostics.png"), modul_6_x3_sentence_gradient_diagnostics, width = 10, height = 7, dpi = 300)
```

```{r}
# MODUL 6.37d — Human-referenced x3 trajectory comparison

human_level <- grep("human", unique(x3_plot_data$model), ignore.case = TRUE, value = TRUE)[1]
stopifnot(!is.na(human_level))
x3_plot_data$model <- stats::relevel(factor(x3_plot_data$model), ref = human_level)
x3_human_model <- stats::lm(cosine_similarity ~ normalized_position * model, data = x3_plot_data)

x3_human_predictions <- tidyr::expand_grid(normalized_position = seq(0, 1, length.out = 101),
                                           model = levels(x3_plot_data$model))
x3_human_predictions$predicted_similarity <- predict(x3_human_model, newdata = x3_human_predictions)

x3_human_predictions <- x3_human_predictions |>
  dplyr::group_by(normalized_position) |>
  dplyr::mutate(human_prediction = predicted_similarity[as.character(model) == human_level],
                human_difference = predicted_similarity - human_prediction) |>
  dplyr::ungroup()

x3_human_trajectory_summary <- x3_human_predictions |>
  dplyr::filter(as.character(model) != human_level) |>
  dplyr::group_by(model) |>
  dplyr::summarise(x3_human_curve_distance_mean = mean(abs(human_difference)),
                   difference_beginning = dplyr::first(human_difference),
                   difference_middle = human_difference[normalized_position == 0.5],
                   difference_end = dplyr::last(human_difference), .groups = "drop")

model_labels <- c(chatgpt_native = "ChatGPT", claude_native = "Claude",
                  gemini_native = "Gemini", human = "Human")

modul_6_x3_human_reference <- ggplot2::ggplot(x3_human_predictions, ggplot2::aes(normalized_position, predicted_similarity, color = model)) +
  ggplot2::geom_line(linewidth = 0.5) +
  ggplot2::scale_color_discrete(name = "Model", labels = model_labels) +
  ggplot2::labs(title = "Human-Referenced x3 Trajectories", x = "Normalized sentence-pair position", y = "Predicted adjacent-sentence similarity")

saveRDS(list(human_reference = human_level, model = x3_human_model, predictions = x3_human_predictions,
             summary = x3_human_trajectory_summary),
        file.path(actual_lsa_output_dir, "modul_6_actual_lsa_x3_human_reference.rds"))
ggplot2::ggsave(file.path(actual_lsa_output_dir, "modul_6_x3_human_reference.png"),
                modul_6_x3_human_reference, width = 10, height = 7, dpi = 300)

print(x3_human_trajectory_summary, n = Inf)
modul_6_x3_human_reference
```
# Note: This cell fits a joint regression model of adjacent-sentence similarity over normalized text position, using the Human baseline as the reference category. It predicts the Human and LM trajectories on a common 0–1 scale and calculates each LM’s difference from Human at the beginning, middle, and end, as well as the mean absolute curve distance. The fitted trajectories, Human-relative summary, regression model, and combined comparison plot are saved for diagnostic analysis.


# MODUL 6.37e — Build and project connector-only x3 segments

```{r}

# MODUL 6.37e — Build and project connector-only x3 segments

x3co_source <- readRDS(file.path(project_root, "data_prep_CCP_output", "current_texts_parsed.rds")) |>
  dplyr::mutate(pos_clean = stringr::str_to_upper(stringr::str_trim(as.character(pos))))

x3co_sentence_map <- sentence_segment_table |>
  dplyr::transmute(text_id = as.character(text_id), sentence_id = as.integer(sentence_id),
                   paragraph_id = as.character(paragraph_id), model) |>
  dplyr::distinct()

stopifnot(!anyDuplicated(x3co_sentence_map[c("text_id", "sentence_id")]))

x3co_tokens <- x3co_source |>
  dplyr::filter(pos_clean %in% c("CCONJ", "SCONJ"), !is.na(lemma), stringr::str_trim(lemma) != "") |>
  dplyr::transmute(text_id = as.character(text_id), sentence_id = as.integer(sentence_id),
                   lemma = stringr::str_to_lower(stringr::str_trim(lemma)), pos_clean) |>
  dplyr::inner_join(x3co_sentence_map, by = c("text_id", "sentence_id"))

stopifnot(nrow(x3co_tokens) > 0L, !anyNA(x3co_tokens$paragraph_id))
print(dplyr::count(x3co_tokens, pos_clean, sort = TRUE), n = Inf)

x3co_sentence_segments <- x3co_sentence_map |>
  dplyr::arrange(text_id, sentence_id) |> dplyr::group_by(text_id) |>
  dplyr::mutate(segment_order = dplyr::row_number()) |> dplyr::ungroup() |>
  dplyr::left_join(x3co_tokens |> dplyr::group_by(text_id, sentence_id) |>
                     dplyr::summarise(terms = paste(lemma, collapse = " "), n_connectors = dplyr::n(), .groups = "drop"),
                   by = c("text_id", "sentence_id")) |>
  dplyr::mutate(segment_id = paste0("x3co_s_", dplyr::row_number()),
                terms = tidyr::replace_na(terms, ""), n_connectors = tidyr::replace_na(n_connectors, 0L))

x3co_paragraph_segments <- paragraph_segment_table |>
  dplyr::transmute(text_id = as.character(text_id), paragraph_id = as.character(paragraph_id), model) |>
  dplyr::distinct() |> dplyr::arrange(text_id, paragraph_id) |> dplyr::group_by(text_id) |>
  dplyr::mutate(segment_order = dplyr::row_number()) |> dplyr::ungroup() |>
  dplyr::left_join(x3co_tokens |> dplyr::group_by(text_id, paragraph_id) |>
                     dplyr::summarise(terms = paste(lemma, collapse = " "), n_connectors = dplyr::n(), .groups = "drop"),
                   by = c("text_id", "paragraph_id")) |>
  dplyr::mutate(segment_id = paste0("x3co_p_", dplyr::row_number()),
                terms = tidyr::replace_na(terms, ""), n_connectors = tidyr::replace_na(n_connectors, 0L))

project_x3co <- function(x) {
  m <- quanteda::dfm_match(quanteda::dfm(quanteda::tokens(x$terms)), features = lsa_reference_space$vocabulary)
  m <- methods::as(m, "dgCMatrix"); m@x <- log1p(m@x)
  m <- m %*% Matrix::Diagonal(x = lsa_reference_space$idf)
  v <- as.matrix(m %*% (lsa_reference_space$right_singular_vectors %*%
                          diag(1 / lsa_reference_space$singular_values)))
  norms <- sqrt(rowSums(v^2)); v <- sweep(v, 1, norms, "/"); v[!is.finite(v)] <- NA_real_
  rownames(v) <- x$segment_id
  list(vectors = v, valid = is.finite(norms) & norms > 0)
}

x3co_sentence_projection <- project_x3co(x3co_sentence_segments)
x3co_paragraph_projection <- project_x3co(x3co_paragraph_segments)

saveRDS(list(sentence_segments = x3co_sentence_segments, paragraph_segments = x3co_paragraph_segments,
             sentence_projection = x3co_sentence_projection, paragraph_projection = x3co_paragraph_projection),
        file.path(actual_lsa_output_dir, "modul_6_actual_lsa_x3co_segments.rds"))

dplyr::bind_rows(
  x3co_sentence_segments |> dplyr::summarise(level = "sentence", segments = dplyr::n(),
                                              with_connectors = sum(n_connectors > 0), connectors = sum(n_connectors)),
  x3co_paragraph_segments |> dplyr::summarise(level = "paragraph", segments = dplyr::n(),
                                               with_connectors = sum(n_connectors > 0), connectors = sum(n_connectors))
)
```

# 

```{r}
# MODUL 6.37g — Calculate and freeze text-internal connector continuity

x3cc_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_x3co_connector_continuity.rds"
)

if (file.exists(x3cc_path)) {

  x3cc_bundle <- readRDS(x3cc_path)
  x3cc_sentence_segments <- x3cc_bundle$sentence_segments
  x3cc_projection <- x3cc_bundle$projection
  x3cc_continuity <- x3cc_bundle$continuity
  x3cc_summary <- x3cc_bundle$summary

  message("Saved connector continuity loaded.")

} else {

  x3cc_sentence_segments <- x3co_sentence_segments |>
    dplyr::select(
      text_id, model, paragraph_id,
      sentence_id, segment_order
    ) |>
    dplyr::distinct() |>
    tidyr::crossing(
      connector_type = c("CCONJ", "SCONJ")
    ) |>
    dplyr::left_join(
      x3co_tokens |>
        dplyr::group_by(
          text_id, model, sentence_id,
          connector_type = pos_clean
        ) |>
        dplyr::summarise(
          terms = paste(lemma, collapse = " "),
          n_connectors = dplyr::n(),
          .groups = "drop"
        ),
      by = c(
        "text_id", "model",
        "sentence_id", "connector_type"
      )
    ) |>
    dplyr::mutate(
      terms = tidyr::replace_na(terms, ""),
      n_connectors = tidyr::replace_na(n_connectors, 0L),
      segment_id = paste(
        text_id, sentence_id, connector_type,
        sep = "_"
      )
    )

  x3cc_projection <- project_x3co(
    x3cc_sentence_segments
  )

  x3cc_sentence_segments$valid_vector <-
    x3cc_projection$valid

  x3cc_cosine <- function(a, b) {

    vectors <- x3cc_projection$vectors
    pair <- vectors[c(a, b), , drop = FALSE]

    if (any(!is.finite(pair))) {
      return(NA_real_)
    }

    sum(pair[1, ] * pair[2, ])
  }

  x3cc_continuity <- x3cc_sentence_segments |>
    dplyr::arrange(
      text_id, connector_type, segment_order
    ) |>
    dplyr::group_by(
      text_id, model, connector_type
    ) |>
    dplyr::mutate(
      next_sentence_id = dplyr::lead(sentence_id),
      next_segment_id = dplyr::lead(segment_id)
    ) |>
    dplyr::filter(
      !is.na(next_segment_id)
    ) |>
    dplyr::mutate(
      normalized_position =
        (dplyr::row_number() - 1) /
        pmax(dplyr::n() - 1, 1),
      transition_id = paste(
        text_id, sentence_id,
        next_sentence_id, connector_type,
        sep = "_"
      )
    ) |>
    dplyr::ungroup() |>
    dplyr::mutate(
      cc_score = purrr::map2_dbl(
        segment_id,
        next_segment_id,
        x3cc_cosine
      )
    )

  x3cc_summary <- x3cc_continuity |>
    dplyr::group_by(
      text_id, model, connector_type
    ) |>
    dplyr::summarise(
      n_pairs = dplyr::n(),
      n_valid_pairs = sum(is.finite(cc_score)),
      x3co_similarity_mean =
        if (n_valid_pairs > 0L) {
          mean(cc_score[is.finite(cc_score)])
        } else {
          NA_real_
        },
      x3co_similarity_sd =
        if (n_valid_pairs > 1L) {
          stats::sd(cc_score[is.finite(cc_score)])
        } else {
          NA_real_
        },
      pair_coverage = n_valid_pairs / n_pairs,
      .groups = "drop"
    )

  x3cc_bundle <- list(
    sentence_segments = x3cc_sentence_segments,
    projection = x3cc_projection,
    continuity = x3cc_continuity,
    summary = x3cc_summary,
    plot_data = x3cc_continuity
  )

  saveRDS(
    x3cc_bundle,
    x3cc_path
  )

  readr::write_csv(
    x3cc_continuity,
    file.path(
      actual_lsa_output_dir,
      "modul_6_actual_lsa_x3co_connector_pairs.csv"
    )
  )

  readr::write_csv(
    x3cc_summary,
    file.path(
      actual_lsa_output_dir,
      "modul_6_actual_lsa_x3co_connector_summary.csv"
    )
  )

  message("Connector continuity calculated and frozen.")
}

stopifnot(
  all(
    c("CCONJ", "SCONJ") %in%
      unique(x3cc_continuity$connector_type)
  )
)

print(
  x3cc_summary,
  n = Inf,
  width = Inf
)
```

# # MODUL 6.37h — Plot Human-centered connector-transition similarity

```{r}

# MODUL 6.37h — Plot Human-centered connector continuity

# Load frozen results

x3cc_path <- file.path(
  project_root,
  "measure_outputs",
  "modul_6_actual_lsa_connector_continuity_human_centered.rds"
)

stopifnot(file.exists(x3cc_path))
x3cc_frozen <- readRDS(x3cc_path)

# Prepare plot data

x3cc_plot <- x3cc_frozen$human_centered |>
  dplyr::mutate(
    model = dplyr::recode(
      as.character(model),
      chatgpt_native = "ChatGPT",
      claude_native = "Claude",
      gemini_native = "Gemini"
    )
  ) |>
  dplyr::select(
    model, connector_type, normalized_position,
    LM = cc_score, Human = human_score
  ) |>
  tidyr::pivot_longer(
    c(LM, Human),
    names_to = "source",
    values_to = "connector_continuity"
  )

# Build plot

modul_6_connector_continuity <- ggplot2::ggplot(
  x3cc_plot,
  ggplot2::aes(
    normalized_position,
    connector_continuity,
    color = source,
    group = source
  )
) +
  ggplot2::geom_hline(yintercept = 0, linetype = "dashed", linewidth = 0.3) +
  ggplot2::geom_line(linewidth = 0.5) +
  ggplot2::geom_point(size = 1.2) +
  ggplot2::facet_grid(connector_type ~ model) +
  ggplot2::coord_cartesian(ylim = c(-1, 1)) +
  ggplot2::theme_minimal(base_size = 11) +
  ggplot2::theme(
    legend.position = "bottom",
    panel.grid.minor = ggplot2::element_blank()
  ) +
  ggplot2::labs(
    title = "Connector Continuity: Human versus Language Models",
    x = "Normalized sentence-transition position",
    y = "Adjacent-sentence connector cosine",
    color = NULL
  )

# Freeze and export plot

saveRDS(
  list(data = x3cc_plot, plot = modul_6_connector_continuity),
  file.path(
    project_root,
    "measure_outputs",
    "modul_6_connector_continuity_plot.rds"
  )
)

ggplot2::ggsave(
  file.path(
    project_root,
    "measure_outputs",
    "modul_6_connector_continuity.png"
  ),
  modul_6_connector_continuity,
  width = 11,
  height = 7,
  dpi = 300
)

# Display plot

modul_6_connector_continuity
```
# Note: Decoupling of Semantics and Propositional Structure  (NEEDS REVISION)

The indices’ x3co plotting (Plot x) suggested that, within a very similar topic and semantic space, the logico‑semantic structure of the discourse is organised relatively differently. As realised through the patterning of connectors, this appears as a connector‑only latent structural skeleton versus a “content mass”.

If content‑word similarity shows monotonic decay (e.g., thematic shifts as sentences progress), but connector continuity does not, it means that the logical architecture of the text is decoupled from the topical progression we can observe within the human baseline. So even when the human writer and the models are writing or providing output about basically the same thing, the way they thread that content with coordinating and subordinating relations is not coupled to the semantic content in the same way as in the human texts.

# MODUL 6.38 -  Build and project x3cw paragraph

```{r}

# MODUL 6.38 — Build and project x3cw paragraphs

# Define output path

x3cw_segments_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_x3cw_segments.rds"
)

# Load frozen output when available

if (file.exists(x3cw_segments_path)) {

  x3cw_segments_bundle <- readRDS(
    x3cw_segments_path
  )

  x3cw_paragraph_segments <-
    x3cw_segments_bundle$paragraph_segments

  x3cw_projection <-
    x3cw_segments_bundle$projection

  message("Saved x3cw paragraph segments loaded.")

} else {

  # Build the validated sentence-to-paragraph map

  x3cw_sentence_map <- sentence_segment_table |>
    dplyr::transmute(
      text_id = as.character(text_id),
      sentence_id = as.integer(sentence_id),
      paragraph_id = as.character(paragraph_id),
      paragraph_number = as.integer(paragraph_number),
      model
    ) |>
    dplyr::distinct()

  # Prepare content-word tokens

  x3cw_tokens <- readRDS(
    file.path(
      project_root,
      "data_prep_CCP_output",
      "current_texts_parsed.rds"
    )
  ) |>
    dplyr::rename(
      parsed_paragraph_id = paragraph_id
    ) |>
    dplyr::mutate(
      text_id = as.character(text_id),
      sentence_id = as.integer(sentence_id),
      pos_clean = stringr::str_to_upper(
        stringr::str_trim(
          as.character(pos)
        )
      ),
      lemma_clean = stringr::str_to_lower(
        stringr::str_trim(
          as.character(lemma)
        )
      )
    ) |>
    dplyr::filter(
      pos_clean %in% c(
        "NOUN",
        "VERB",
        "ADJ",
        "ADV"
      ),
      !is.na(lemma_clean),
      lemma_clean != "",
      lemma_clean != "null"
    ) |>
    dplyr::inner_join(
      x3cw_sentence_map,
      by = c(
        "text_id",
        "sentence_id"
      ),
      relationship = "many-to-one"
    )

  # Build content-word paragraph segments

  x3cw_paragraph_segments <- paragraph_segment_table |>
    dplyr::transmute(
      text_id = as.character(text_id),
      model,
      paragraph_id = as.character(paragraph_id),
      paragraph_number = as.integer(paragraph_number),
      source_segment_id = segment_id
    ) |>
    dplyr::left_join(
      x3cw_tokens |>
        dplyr::group_by(
          text_id,
          model,
          paragraph_id,
          paragraph_number
        ) |>
        dplyr::summarise(
          terms = paste(
            lemma_clean,
            collapse = " "
          ),
          n_content_tokens = dplyr::n(),
          .groups = "drop"
        ),
      by = c(
        "text_id",
        "model",
        "paragraph_id",
        "paragraph_number"
      ),
      relationship = "one-to-one"
    ) |>
    dplyr::mutate(
      segment_id = paste0(
        source_segment_id,
        "_x3cw"
      ),
      terms = tidyr::replace_na(
        terms,
        ""
      ),
      n_content_tokens = tidyr::replace_na(
        n_content_tokens,
        0L
      )
    ) |>
    dplyr::arrange(
      text_id,
      paragraph_number
    ) |>
    dplyr::group_by(
      text_id,
      model
    ) |>
    dplyr::mutate(
      segment_order = dplyr::row_number()
    ) |>
    dplyr::ungroup()

  # Construct the matched DFM

  x3cw_dfm <- quanteda::dfm_match(
    quanteda::dfm(
      quanteda::tokens(
        x3cw_paragraph_segments$terms
      )
    ),
    features = lsa_reference_space$vocabulary
  )

  x3cw_matrix <- methods::as(
    x3cw_dfm,
    "dgCMatrix"
  )

  x3cw_matrix@x <- log1p(
    x3cw_matrix@x
  )

  x3cw_matrix <- x3cw_matrix %*%
    Matrix::Diagonal(
      x = lsa_reference_space$idf
    )

  # Fold segments into the frozen LSA space

  x3cw_vectors_raw <- as.matrix(
    x3cw_matrix %*%
      (
        lsa_reference_space$
          right_singular_vectors %*%
          diag(
            1 /
              lsa_reference_space$
                singular_values
          )
      )
  )

  x3cw_norms <- sqrt(
    rowSums(
      x3cw_vectors_raw^2
    )
  )

  x3cw_valid <- is.finite(
    x3cw_norms
  ) & x3cw_norms > 0

  x3cw_vectors <- matrix(
    NA_real_,
    nrow = nrow(
      x3cw_vectors_raw
    ),
    ncol = ncol(
      x3cw_vectors_raw
    )
  )

  x3cw_vectors[x3cw_valid, ] <- sweep(
    x3cw_vectors_raw[
      x3cw_valid,
      ,
      drop = FALSE
    ],
    1,
    x3cw_norms[x3cw_valid],
    "/"
  )

  rownames(x3cw_vectors) <-
    x3cw_paragraph_segments$segment_id

  x3cw_projection <- list(
    vectors = x3cw_vectors,
    norms = x3cw_norms,
    valid = x3cw_valid
  )

  # Validate and freeze

  stopifnot(
    nrow(
      x3cw_paragraph_segments
    ) == 15L,
    all(
      x3cw_paragraph_segments$
        n_content_tokens > 0L
    ),
    !anyDuplicated(
      x3cw_paragraph_segments$
        segment_id
    ),
    identical(
      rownames(
        x3cw_projection$vectors
      ),
      x3cw_paragraph_segments$
        segment_id
    )
  )

  x3cw_segments_bundle <- list(
    pos_keep = c(
      "NOUN",
      "VERB",
      "ADJ",
      "ADV"
    ),
    paragraph_segments =
      x3cw_paragraph_segments,
    projection =
      x3cw_projection
  )

  saveRDS(
    x3cw_segments_bundle,
    x3cw_segments_path
  )

  message(
    "x3cw paragraph segments projected and frozen."
  )
}

# Display audit

x3cw_paragraph_segments |>
  dplyr::summarise(
    n_paragraphs = dplyr::n(),
    n_content_tokens = sum(
      n_content_tokens
    ),
    n_valid_vectors = sum(
      x3cw_projection$valid
    )
  ) |>
  print(
    width = Inf
  )

```
# # MODUL 6.37j — Calculate and freeze x3cw continuity

```{r}

# MODUL 6.37j — Calculate and freeze x3cw continuity

x3cw_output_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_x3cw_paragraph_continuity.rds"
)

if (file.exists(x3cw_output_path)) {

  x3cw_bundle <- readRDS(x3cw_output_path)
  x3cw_paragraph_pairs <- x3cw_bundle$paragraph_pairs
  x3cw_profiles <- x3cw_bundle$profiles

  message("Saved x3cw continuity loaded.")

} else {

  x3cw_cosine <- function(segment_1, segment_2) {

    vector_1 <- x3cw_projection$vectors[
      segment_1,
      ,
      drop = TRUE
    ]

    vector_2 <- x3cw_projection$vectors[
      segment_2,
      ,
      drop = TRUE
    ]

    if (
      anyNA(vector_1) ||
      anyNA(vector_2)
    ) {
      return(NA_real_)
    }

    sum(vector_1 * vector_2)
  }

  x3cw_paragraph_pairs <- x3cw_paragraph_segments |>
    dplyr::arrange(
      text_id,
      segment_order
    ) |>
    dplyr::group_by(
      text_id,
      model
    ) |>
    dplyr::mutate(
      next_segment_id = dplyr::lead(segment_id),
      next_paragraph_id = dplyr::lead(paragraph_id),
      next_paragraph_number =
        dplyr::lead(paragraph_number)
    ) |>
    dplyr::filter(
      !is.na(next_segment_id)
    ) |>
    dplyr::mutate(
      normalized_position =
        (dplyr::row_number() - 1) /
        pmax(dplyr::n() - 1, 1)
    ) |>
    dplyr::ungroup() |>
    dplyr::mutate(
      x3cw_adjacent_paragraph_similarity =
        purrr::map2_dbl(
          segment_id,
          next_segment_id,
          x3cw_cosine
        )
    ) |>
    dplyr::select(
      text_id,
      model,
      paragraph_id,
      paragraph_number,
      next_paragraph_id,
      next_paragraph_number,
      segment_id_1 = segment_id,
      segment_id_2 = next_segment_id,
      normalized_position,
      x3cw_adjacent_paragraph_similarity
    )

  x3cw_profiles <- x3cw_paragraph_pairs |>
    dplyr::group_by(
      text_id,
      model
    ) |>
    dplyr::summarise(
      x3cw_adjacent_paragraph_similarity_mean =
        mean(
          x3cw_adjacent_paragraph_similarity,
          na.rm = TRUE
        ),
      n_adjacent_paragraph_pairs = dplyr::n(),
      n_valid_x3cw_pairs = sum(
        is.finite(
          x3cw_adjacent_paragraph_similarity
        )
      ),
      .groups = "drop"
    )

  valid_scores <-
    x3cw_paragraph_pairs$
      x3cw_adjacent_paragraph_similarity

  valid_scores <- valid_scores[
    is.finite(valid_scores)
  ]

  stopifnot(
    nrow(x3cw_paragraph_pairs) == 11L,
    nrow(x3cw_profiles) == 4L,
    all(valid_scores >= -1 & valid_scores <= 1)
  )

  x3cw_bundle <- list(
    definition =
      "Content-word-only paragraph fold-in using NOUN, VERB, ADJ, and ADV within the frozen seven-POS LSA reference space.",
    paragraph_segments = x3cw_paragraph_segments,
    projection = x3cw_projection,
    paragraph_pairs = x3cw_paragraph_pairs,
    profiles = x3cw_profiles
  )

  saveRDS(
    x3cw_bundle,
    x3cw_output_path
  )

  readr::write_csv(
    x3cw_paragraph_pairs,
    file.path(
      actual_lsa_output_dir,
      "modul_6_actual_lsa_x3cw_paragraph_pairs.csv"
    )
  )

  readr::write_csv(
    x3cw_profiles,
    file.path(
      actual_lsa_output_dir,
      "modul_6_actual_lsa_x3cw_profiles.csv"
    )
  )

  message("x3cw paragraph continuity calculated and frozen.")
}

print(
  x3cw_profiles,
  n = Inf,
  width = Inf
)

print(
  x3cw_paragraph_pairs,
  n = Inf,
  width = Inf
)
```


# # MODUL 6.38 — Canonicalize and freeze x3s, x3p, and x3pcw

```{r}

# MODUL 6.38 — Canonicalize and freeze x3s, x3p, and x3pcw

# 1. Define input and output paths

x3_legacy_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_x3.rds"
)

x3s_gradient_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_x3_gradient.rds"
)

x3pcw_legacy_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_x3cw_paragraph_continuity.rds"
)

x3s_output_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_x3s_sentence_continuity.rds"
)

x3p_output_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_x3p_paragraph_continuity.rds"
)

x3pcw_output_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_x3pcw_paragraph_continuity.rds"
)

# 2. Load the frozen legacy objects

stopifnot(
  file.exists(x3_legacy_path),
  file.exists(x3s_gradient_path),
  file.exists(x3pcw_legacy_path)
)

x3_legacy <- readRDS(x3_legacy_path)
x3s_gradient_legacy <- readRDS(x3s_gradient_path)
x3pcw_legacy <- readRDS(x3pcw_legacy_path)

# 3. Build the canonical x3s sentence layer

x3s_sentence_pairs <- x3_legacy$sentence_pairs |>
  dplyr::arrange(
    text_id,
    sentence_order_1
  ) |>
  dplyr::group_by(
    text_id,
    model
  ) |>
  dplyr::mutate(
    normalized_position =
      (dplyr::row_number() - 1) /
      pmax(dplyr::n() - 1, 1)
  ) |>
  dplyr::ungroup() |>
  dplyr::rename(
    x3s_adjacent_sentence_similarity =
      cosine_similarity
  )

x3s_profiles <- x3s_gradient_legacy$profiles |>
  dplyr::transmute(
    text_id,
    model,
    x3s_adjacent_sentence_similarity_mean =
      x3_adjacent_sentence_similarity_mean,
    x3s_adjacent_sentence_similarity_sd =
      x3_adjacent_sentence_similarity_sd,
    x3s_adjacent_sentence_similarity_gradient =
      x3_adjacent_sentence_similarity_gradient,
    x3s_adjacent_sentence_gradient_se =
      x3_adjacent_sentence_gradient_se,
    x3s_adjacent_sentence_gradient_r2 =
      x3_adjacent_sentence_gradient_r2,
    x3s_adjacent_sentence_similarity_first_third =
      x3_adjacent_sentence_similarity_first_third,
    x3s_adjacent_sentence_similarity_middle_third =
      x3_adjacent_sentence_similarity_middle_third,
    x3s_adjacent_sentence_similarity_final_third =
      x3_adjacent_sentence_similarity_final_third,
    x3s_adjacent_sentence_similarity_retention =
      x3_adjacent_sentence_similarity_retention,
    n_x3s_adjacent_sentence_pairs =
      n_adjacent_sentence_pairs
  )

# 4. Build the canonical x3p paragraph layer

x3p_paragraph_pairs <- x3_legacy$paragraph_pairs |>
  dplyr::arrange(
    text_id,
    paragraph_order_1
  ) |>
  dplyr::group_by(
    text_id,
    model
  ) |>
  dplyr::mutate(
    normalized_position =
      (dplyr::row_number() - 1) /
      pmax(dplyr::n() - 1, 1)
  ) |>
  dplyr::ungroup() |>
  dplyr::rename(
    x3p_adjacent_paragraph_similarity =
      cosine_similarity
  )

x3p_profiles <- x3_legacy$profiles |>
  dplyr::transmute(
    text_id,
    model,
    x3p_adjacent_paragraph_similarity_mean =
      x3_adjacent_paragraph_similarity_mean,
    n_x3p_adjacent_paragraph_pairs =
      n_adjacent_paragraph_pairs
  )

# 5. Build the canonical x3pcw paragraph layer

x3pcw_paragraph_pairs <-
  x3pcw_legacy$paragraph_pairs |>
  dplyr::rename(
    x3pcw_adjacent_paragraph_similarity =
      x3cw_adjacent_paragraph_similarity
  )

x3pcw_profiles <- x3pcw_legacy$profiles |>
  dplyr::transmute(
    text_id,
    model,
    x3pcw_adjacent_paragraph_similarity_mean =
      x3cw_adjacent_paragraph_similarity_mean,
    n_x3pcw_adjacent_paragraph_pairs =
      n_adjacent_paragraph_pairs,
    n_valid_x3pcw_pairs =
      n_valid_x3cw_pairs
  )

# 6. Validate the canonical layers

stopifnot(
  nrow(x3s_profiles) == 4L,
  nrow(x3p_paragraph_pairs) == 11L,
  nrow(x3p_profiles) == 4L,
  nrow(x3pcw_paragraph_pairs) == 11L,
  nrow(x3pcw_profiles) == 4L,
  sum(x3s_profiles$n_x3s_adjacent_sentence_pairs) ==
    nrow(x3s_sentence_pairs),
  sum(x3p_profiles$n_x3p_adjacent_paragraph_pairs) ==
    nrow(x3p_paragraph_pairs),
  sum(x3pcw_profiles$n_x3pcw_adjacent_paragraph_pairs) ==
    nrow(x3pcw_paragraph_pairs),
  all(is.finite(
    x3s_sentence_pairs$x3s_adjacent_sentence_similarity
  )),
  all(is.finite(
    x3p_paragraph_pairs$x3p_adjacent_paragraph_similarity
  )),
  all(is.finite(
    x3pcw_paragraph_pairs$
      x3pcw_adjacent_paragraph_similarity
  ))
)

# 7. Freeze the canonical layers

x3s_bundle <- list(
  level = "sentence",
  definition =
    "Sentence-level adjacent semantic continuity.",
  sentence_pairs = x3s_sentence_pairs,
  profiles = x3s_profiles
)

x3p_bundle <- list(
  level = "paragraph",
  definition =
    "Paragraph-level adjacent semantic continuity using the full seven-POS target layer.",
  paragraph_pairs = x3p_paragraph_pairs,
  profiles = x3p_profiles
)

x3pcw_bundle <- list(
  level = "paragraph_content_words",
  definition =
    "Paragraph-level adjacent topic continuity using NOUN, VERB, ADJ, and ADV.",
  pos_keep = c(
    "NOUN",
    "VERB",
    "ADJ",
    "ADV"
  ),
  paragraph_segments =
    x3pcw_legacy$paragraph_segments,
  projection =
    x3pcw_legacy$projection,
  paragraph_pairs =
    x3pcw_paragraph_pairs,
  profiles =
    x3pcw_profiles
)

saveRDS(x3s_bundle, x3s_output_path)
saveRDS(x3p_bundle, x3p_output_path)
saveRDS(x3pcw_bundle, x3pcw_output_path)

readr::write_csv(
  x3s_sentence_pairs,
  file.path(
    actual_lsa_output_dir,
    "modul_6_actual_lsa_x3s_sentence_pairs.csv"
  )
)

readr::write_csv(
  x3p_paragraph_pairs,
  file.path(
    actual_lsa_output_dir,
    "modul_6_actual_lsa_x3p_paragraph_pairs.csv"
  )
)

readr::write_csv(
  x3pcw_paragraph_pairs,
  file.path(
    actual_lsa_output_dir,
    "modul_6_actual_lsa_x3pcw_paragraph_pairs.csv"
  )
)

readr::write_csv(
  x3s_profiles,
  file.path(
    actual_lsa_output_dir,
    "modul_6_actual_lsa_x3s_profiles.csv"
  )
)

readr::write_csv(
  x3p_profiles,
  file.path(
    actual_lsa_output_dir,
    "modul_6_actual_lsa_x3p_profiles.csv"
  )
)

readr::write_csv(
  x3pcw_profiles,
  file.path(
    actual_lsa_output_dir,
    "modul_6_actual_lsa_x3pcw_profiles.csv"
  )
)

# 8. Display the canonicalization audit

tibble::tibble(
  layer = c(
    "x3s sentence pairs",
    "x3s profiles",
    "x3p paragraph pairs",
    "x3p profiles",
    "x3pcw paragraph pairs",
    "x3pcw profiles"
  ),
  rows = c(
    nrow(x3s_sentence_pairs),
    nrow(x3s_profiles),
    nrow(x3p_paragraph_pairs),
    nrow(x3p_profiles),
    nrow(x3pcw_paragraph_pairs),
    nrow(x3pcw_profiles)
  )
) |>
  print(
    n = Inf,
    width = Inf
  )
```


# 8. Reload and verify the canonical frozen objects

```{r}
# 8. Reload and verify the canonical frozen objects

x3s_reload <- readRDS(x3s_output_path)
x3p_reload <- readRDS(x3p_output_path)
x3pcw_reload <- readRDS(x3pcw_output_path)

stopifnot(
  "x3s_adjacent_sentence_similarity" %in%
    names(x3s_reload$sentence_pairs),

  "x3s_adjacent_sentence_similarity_mean" %in%
    names(x3s_reload$profiles),

  "x3p_adjacent_paragraph_similarity" %in%
    names(x3p_reload$paragraph_pairs),

  "x3p_adjacent_paragraph_similarity_mean" %in%
    names(x3p_reload$profiles),

  "x3pcw_adjacent_paragraph_similarity" %in%
    names(x3pcw_reload$paragraph_pairs),

  "x3pcw_adjacent_paragraph_similarity_mean" %in%
    names(x3pcw_reload$profiles)
)

message(
  "Canonical x3s, x3p, and x3pcw names verified in frozen outputs."
)
```


# MODUL 6.38a — Calculate and freeze x3pcw topic-continuity (alias semantic similarity continuity) gradients

```{r}

# MODUL 6.38a — Calculate and freeze x3pcw topic-continuity gradients

# 1. Define input and output paths

x3pcw_input_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_x3pcw_paragraph_continuity.rds"
)

x3pcw_gradient_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_x3pcw_topic_continuity_gradient.rds"
)

x3pcw_gradient_csv_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_x3pcw_topic_continuity_gradient.csv"
)

# 2. Load the canonical x3pcw layer

stopifnot(
  file.exists(x3pcw_input_path)
)

x3pcw_bundle <- readRDS(
  x3pcw_input_path
)

x3pcw_paragraph_pairs <-
  x3pcw_bundle$paragraph_pairs

x3pcw_profiles <-
  x3pcw_bundle$profiles

stopifnot(
  nrow(x3pcw_paragraph_pairs) == 11L,
  nrow(x3pcw_profiles) == 4L,
  all(
    c(
      "text_id",
      "model",
      "normalized_position",
      "x3pcw_adjacent_paragraph_similarity"
    ) %in%
      names(x3pcw_paragraph_pairs)
  )
)

# 3. Define the gradient function

calculate_x3pcw_gradient <- function(data) {

  valid_data <- data |>
    dplyr::filter(
      is.finite(normalized_position),
      is.finite(
        x3pcw_adjacent_paragraph_similarity
      )
    ) |>
    dplyr::arrange(
      normalized_position
    )

  n_valid <- nrow(
    valid_data
  )

  if (n_valid < 2L) {
    return(
      tibble::tibble(
        n_valid_x3pcw_gradient_pairs = n_valid,
        x3pcw_topic_continuity_first = NA_real_,
        x3pcw_topic_continuity_final = NA_real_,
        x3pcw_topic_continuity_change = NA_real_,
        x3pcw_adjacent_paragraph_similarity_gradient = NA_real_,
        x3pcw_adjacent_paragraph_gradient_se = NA_real_,
        x3pcw_adjacent_paragraph_gradient_r2 = NA_real_
      )
    )
  }

  gradient_model <- stats::lm(
    x3pcw_adjacent_paragraph_similarity ~
      normalized_position,
    data = valid_data
  )

  gradient_summary <- summary(
    gradient_model
  )

  tibble::tibble(
    n_valid_x3pcw_gradient_pairs = n_valid,
    x3pcw_topic_continuity_first =
      dplyr::first(
        valid_data$
          x3pcw_adjacent_paragraph_similarity
      ),
    x3pcw_topic_continuity_final =
      dplyr::last(
        valid_data$
          x3pcw_adjacent_paragraph_similarity
      ),
    x3pcw_topic_continuity_change =
      x3pcw_topic_continuity_final -
      x3pcw_topic_continuity_first,
    x3pcw_adjacent_paragraph_similarity_gradient =
      unname(
        stats::coef(
          gradient_model
        )[["normalized_position"]]
      ),
    x3pcw_adjacent_paragraph_gradient_se =
      dplyr::if_else(
        n_valid >= 3L,
        unname(
          gradient_summary$coefficients[
            "normalized_position",
            "Std. Error"
          ]
        ),
        NA_real_
      ),
    x3pcw_adjacent_paragraph_gradient_r2 =
      dplyr::if_else(
        n_valid >= 3L,
        unname(
          gradient_summary$r.squared
        ),
        NA_real_
      )
  )
}

# 4. Calculate topic-continuity gradients

x3pcw_gradient_profiles <-
  x3pcw_paragraph_pairs |>
  dplyr::group_by(
    text_id,
    model
  ) |>
  dplyr::group_modify(
    ~ calculate_x3pcw_gradient(.x)
  ) |>
  dplyr::ungroup()

# 5. Extend the x3pcw text profiles

x3pcw_profiles_extended <-
  x3pcw_profiles |>
  dplyr::left_join(
    x3pcw_gradient_profiles,
    by = c(
      "text_id",
      "model"
    )
  ) |>
  dplyr::arrange(
    text_id
  )

# 6. Validate the gradient results

stopifnot(
  nrow(x3pcw_profiles_extended) == 4L,
  all(
    is.finite(
      x3pcw_profiles_extended$
        x3pcw_adjacent_paragraph_similarity_gradient
    )
  ),
  all(
    x3pcw_profiles_extended$
      n_valid_x3pcw_gradient_pairs >= 2L
  ),
  all(
    x3pcw_profiles_extended$
      x3pcw_adjacent_paragraph_similarity_gradient >= -2 &
      x3pcw_profiles_extended$
      x3pcw_adjacent_paragraph_similarity_gradient <= 2
  )
)

# 7. Freeze the gradient bundle

x3pcw_gradient_bundle <- list(
  definition =
    paste(
      "Development of paragraph-level topic continuity",
      "over normalized text position using content-word-only",
      "adjacent-paragraph LSA similarity."
    ),
  paragraph_pairs =
    x3pcw_paragraph_pairs,
  gradient_profiles =
    x3pcw_gradient_profiles,
  profiles_extended =
    x3pcw_profiles_extended
)

saveRDS(
  x3pcw_gradient_bundle,
  x3pcw_gradient_path
)

readr::write_csv(
  x3pcw_profiles_extended,
  x3pcw_gradient_csv_path
)

# 8. Reload and verify the frozen output

x3pcw_gradient_reload <- readRDS(
  x3pcw_gradient_path
)

stopifnot(
  nrow(
    x3pcw_gradient_reload$
      paragraph_pairs
  ) == 11L,
  nrow(
    x3pcw_gradient_reload$
      profiles_extended
  ) == 4L,
  "x3pcw_adjacent_paragraph_similarity_gradient" %in%
    names(
      x3pcw_gradient_reload$
        profiles_extended
    )
)

message(
  "x3pcw topic-continuity gradients verified in frozen output."
)

# 9. Display the extended profiles

x3pcw_profiles_extended |>
  dplyr::select(
    text_id,
    model,
    x3pcw_adjacent_paragraph_similarity_mean,
    x3pcw_topic_continuity_first,
    x3pcw_topic_continuity_final,
    x3pcw_topic_continuity_change,
    x3pcw_adjacent_paragraph_similarity_gradient,
    x3pcw_adjacent_paragraph_gradient_se,
    x3pcw_adjacent_paragraph_gradient_r2,
    n_valid_x3pcw_gradient_pairs
  ) |>
  print(
    n = Inf,
    width = Inf
  )
```
# Note: The Semantic Coherence Matrix
In the yet to follow matrix, Topic Continuity (operationalized as adjacent semantic similarity) is a graded measure of semantic continuity that can be higher or lower. The Topic-continuity gradient represents the change in adjacent semantic similarity over normalized text position (here on paragraph level, a paragraph was vectorized for the LSA to be applied). The x3 (x3s [sentnce-level], x3p [paragraph-level], x3pcw [content words over paragraphs]) indices were calculated over 
x3s sentence pairs:       102
x3s profiles:               4
+
x3p paragraph pairs:       11
x3p profiles:               4
+
x3pcw paragraph pairs:     11
x3pcw profiles:             4


Current x3pcw results
Mean Topic-Continuity level
Human      0.3056374
Claude     0.2454105
Gemini     0.1643951
ChatGPT    0.1482608

Ranking:

Human > Claude > Gemini > ChatGPT
Topic-continuity gradients
ChatGPT   +0.11153652
Claude    −0.13371286
Gemini    −0.27977701
Human     +0.05262017

Interpretation:

Human
→ highest overall Topic Continuity;
  small positive trajectory

Claude
→ relatively high Topic Continuity;
  declining trajectory

ChatGPT
→ lowest overall Topic Continuity;
  improving trajectory from a very low starting point

Gemini
→ low Topic Continuity;
  strongest decline

For texts with only two paragraph transitions, gradient standard error and R 2 remain NA.

Observed R
2
 values:

ChatGPT   0.1062
Human     0.0367
Claude    NA
Gemini    NA

The low values for ChatGPT and Human show that their trajectories are not strongly linear.

Semantic Coherence Matrix
Topic-continuity level	Positive or near-stable gradient	Negative gradient
High	Human: high and comparatively stable continuity	Claude: high level, fading continuity
Low	ChatGPT: low but improving continuity	Gemini: low level, rapid continuity loss

Out of need for normalization, we hypothesize a idealized Point d’Or which configaration would be

high Topic-Continuity level + gradient near zero = high and stable semantic continuity

The human generated text stands out in that regard and is currently closest to this configuration.

Topic Continuity (Level): This is given by the mean of adjacent semantic similarity. It answers the question: How high or low is the topic continuity throughout the text over existent paragraphs?

Trajectory (Gradient): This is given by the change in adjacent semantic similarity over normalized text (continuiation). It answers the question: Does the topic continuity increase, decrease, or remain stable as the text progresses?

Hovewer it appears obvious to ask whether and how is paragraph-level LSA Topic Continuity associated with exact content-lemma overlap? Because current text-level rankings appear different: y4 content-lemma overlap → 1.ChatGPT > 2. Human > 3.Gemini > 4. Claude and x3pcw Topic Continuity → 1. Human > 2. Claude > 3. Gemini > 4. ChatGPT,  the analysis asks whether higher exact content-lemma overlap predicts higher LSA-based Topic Continuity, which motivates the the later pair-level correlation analysis. Therefore to grasp that central research question at this point the relevant measures become  y4 (content_jaccard) and x3pcw and its exploring their relation:

y4_content_jaccard ↔ x3pcw_adjacent_paragraph_similarity

Both measures use the same 11 adjacent paragraph transitions, i.e. they use NOUN, VERB, ADJ, ADV and operate on paragraph level.

y4_content_jaccard shows the exact recurrence of content lemmas, while x3pcw is intended to latent semantic Topic Continuity. 

# --------- end of Note -------------------------------------------

# # MODUL 6.38b — Join and freeze matched y4 and x3pcw paragraph transitions
This cell reads the two frozen pair-level sources, verifies identical paragraph boundaries, creates normalized within-text position, and freezes the 11-transition table. The y4 source already contains paragraph_number, next_paragraph_number, and y4_content_jaccard.
* We are doing this to test whether exact content-word repetition (y4) actually co-varies with latent paragraph-level Topic Continuity (x3pcw) across the same 11 paragraph transitions. The joined table is needed because the correlation, regression, and scatterplot require both measures aligned row by row for each identical transition.


#

```{r}
file.exists(
  file.path(
    actual_lsa_output_dir,
    "modul_6_actual_lsa_y4_x3pcw_matched_transitions.rds"
  )
)
```


# Modul 6.38b -

```{r}
# MODUL 6.38b — Join and freeze matched y4 and x3pcw paragraph transitions

matched_transition_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_y4_x3pcw_matched_transitions.rds"
)

if (file.exists(matched_transition_path)) {

  y4_x3pcw_matched_transitions <- readRDS(matched_transition_path)

} else {

  y4_pairs <- readRDS(
    file.path(actual_lsa_output_dir, "ccp_module_5c_y4_pairwise.rds")
  ) |>
    dplyr::transmute(
      text_id,
      model,
      paragraph_number_1 = paragraph_number,
      paragraph_number_2 = next_paragraph_number,
      y4_content_jaccard
    )

  x3pcw_pairs <- readRDS(
    file.path(
      actual_lsa_output_dir,
      "modul_6_actual_lsa_x3pcw_paragraph_continuity.rds"
    )
  )$paragraph_pairs |>
    dplyr::transmute(
      text_id,
      model,
      paragraph_number_1 = paragraph_number,
      paragraph_number_2 = next_paragraph_number,
      x3pcw_adjacent_paragraph_similarity
    ) |>
    dplyr::arrange(text_id, paragraph_number_1) |>
    dplyr::group_by(text_id, model) |>
    dplyr::mutate(
      paragraph_pair_id = paste0(
        "P", paragraph_number_1, "_P", paragraph_number_2
      ),
      normalized_position =
        (dplyr::row_number() - 1) / pmax(dplyr::n() - 1, 1)
    ) |>
    dplyr::ungroup()

  y4_x3pcw_matched_transitions <- x3pcw_pairs |>
    dplyr::inner_join(
      y4_pairs,
      by = c(
        "text_id",
        "model",
        "paragraph_number_1",
        "paragraph_number_2"
      ),
      relationship = "one-to-one"
    ) |>
    dplyr::mutate(
      source_group = dplyr::if_else(
        text_id == "HUMAN_BASELINE_001",
        "human",
        "lm"
      )
    ) |>
    dplyr::select(
      text_id,
      model,
      source_group,
      paragraph_number_1,
      paragraph_number_2,
      paragraph_pair_id,
      normalized_position,
      y4_content_jaccard,
      x3pcw_adjacent_paragraph_similarity
    )

  saveRDS(y4_x3pcw_matched_transitions, matched_transition_path)
}

stopifnot(
  nrow(y4_x3pcw_matched_transitions) == 11L,
  dplyr::n_distinct(y4_x3pcw_matched_transitions$text_id) == 4L,
  table(y4_x3pcw_matched_transitions$source_group)[["lm"]] == 7L,
  table(y4_x3pcw_matched_transitions$source_group)[["human"]] == 4L,
  !anyNA(
    y4_x3pcw_matched_transitions[
      c(
        "y4_content_jaccard",
        "x3pcw_adjacent_paragraph_similarity"
      )
    ]
  )
)

y4_x3pcw_matched_transitions
```


#

```{r}
stopifnot(
  nrow(y4_x3pcw_matched_transitions) == 11L,
  dplyr::n_distinct(y4_x3pcw_matched_transitions$text_id) == 4L,
  sum(y4_x3pcw_matched_transitions$source_group == "lm") == 7L,
  sum(y4_x3pcw_matched_transitions$source_group == "human") == 4L,
  all(
    y4_x3pcw_matched_transitions$paragraph_number_2 ==
      y4_x3pcw_matched_transitions$paragraph_number_1 + 1L
  ),
  sum(is.na(
    y4_x3pcw_matched_transitions$y4_content_jaccard
  )) == 0L,
  sum(is.na(
    y4_x3pcw_matched_transitions$
      x3pcw_adjacent_paragraph_similarity
  )) == 0L
)
message(
  "Matched y4–x3pcw transitions validated successfully: ",
  nrow(y4_x3pcw_matched_transitions),
  " transitions across ",
  dplyr::n_distinct(y4_x3pcw_matched_transitions$text_id),
  " texts."
)
```
# Note: stopifnot() produces no output when all conditions are satisfied.


# # MODUL 6.38c — Calculate and freeze y4–x3pcw relationship diagnostics

```{r}
# MODUL 6.38c — Calculate and freeze y4–x3pcw relationship diagnostics

y4_x3pcw_matched_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_y4_x3pcw_matched_transitions.rds"
)

y4_x3pcw_diagnostics_rds_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_y4_x3pcw_relationship_diagnostics.rds"
)

y4_x3pcw_diagnostics_csv_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_y4_x3pcw_relationship_diagnostics.csv"
)

y4_x3pcw_matched_transitions <- readRDS(
  y4_x3pcw_matched_path
)

calculate_y4_x3pcw_relationship <- function(data, scope) {
  analysis_data <- data |>
    dplyr::filter(
      !is.na(y4_content_jaccard),
      !is.na(x3pcw_adjacent_paragraph_similarity)
    )

  relationship_model <- stats::lm(
    x3pcw_adjacent_paragraph_similarity ~
      y4_content_jaccard,
    data = analysis_data
  )

  tibble::tibble(
    analysis_scope = scope,
    n_matched_transitions = nrow(analysis_data),
    pearson_correlation = stats::cor(
      analysis_data$y4_content_jaccard,
      analysis_data$x3pcw_adjacent_paragraph_similarity,
      method = "pearson"
    ),
    spearman_correlation = stats::cor(
      analysis_data$y4_content_jaccard,
      analysis_data$x3pcw_adjacent_paragraph_similarity,
      method = "spearman"
    ),
    regression_intercept = unname(
      stats::coef(relationship_model)[1]
    ),
    regression_slope = unname(
      stats::coef(relationship_model)[2]
    ),
    regression_r2 = summary(
      relationship_model
    )$r.squared
  )
}

y4_x3pcw_relationship_diagnostics <- dplyr::bind_rows(
  calculate_y4_x3pcw_relationship(
    y4_x3pcw_matched_transitions,
    "all_texts"
  ),
  calculate_y4_x3pcw_relationship(
    dplyr::filter(
      y4_x3pcw_matched_transitions,
      source_group == "lm"
    ),
    "lms_only"
  ),
  calculate_y4_x3pcw_relationship(
    dplyr::filter(
      y4_x3pcw_matched_transitions,
      source_group == "human"
    ),
    "human_only"
  )
)

saveRDS(
  y4_x3pcw_relationship_diagnostics,
  y4_x3pcw_diagnostics_rds_path
)

readr::write_csv(
  y4_x3pcw_relationship_diagnostics,
  y4_x3pcw_diagnostics_csv_path
)

y4_x3pcw_relationship_diagnostics <- readRDS(
  y4_x3pcw_diagnostics_rds_path
)

stopifnot(
  nrow(y4_x3pcw_relationship_diagnostics) == 3L,
  identical(
    y4_x3pcw_relationship_diagnostics$
      n_matched_transitions,
    c(11L, 7L, 4L)
  ),
  all(
    y4_x3pcw_relationship_diagnostics$analysis_scope ==
      c("all_texts", "lms_only", "human_only")
  )
)

y4_x3pcw_relationship_diagnostics
```
# Interpretative note:

The primary all-text analysis suggests 

greater exact content-lemma recurrence
tends to coincide with greater latent Topic Continuity, but only within the human baseline. The LM-only result is essentially zero:

Pearson r = −0.0224
Spearman ρ = −0.0714
R² = 0.0005

This means that exact content-word repetition does not meaningfully predict x3pcw Topic Continuity across the seven LM transitions!

The positive pooled relationship appears to be exclusivelly influenced by the four Human transitions, which show a strong positive pattern. Because the Human analysis has only four observations, it must remain exploratory.

Conclusive note: Across all eleven transitions, exact content-lemma recurrence showed a moderate positive descriptive association with latent Topic Continuity. This relationship disappeared when the analysis was restricted to the seven LM transitions, suggesting that the pooled association was substantially influenced by the Human paragraph transitions. Exact lexical recurrence and latent semantic continuity therefore cannot be treated as interchangeable properties in the present data.
The way the LMs use lexical material repetitively to maintain Topic Continuity does not lead to a corresponding Topic-Continuity resolution as measured with LSA. This is therefore one of the clearest decoupling indices the CCP has shown so far: exact lexical recurrence and latent semantic continuity are structurally distinct and do not necessarily co-vary.

More specifically, the LM-only analysis shows that the repetitive use of content lemmas does not contribute at all to explaining variation in x3pcw Topic Continuity:


r=-0.0224,\qquad
\rho=-0.0714,\qquad
R^2=0.0005.

Interpretation Note:  The LMs can therefore use lexical material repetitively to maintain the appearance of topic continuity without this repetition structurally leading to a corresponding latent Topic-Continuity resolution in the LSA space, something we know from Bajohr’s observations about “surface cohesion.” However, the CCP provides a more fine-grained hypothesis: in a deeper sense, as explained throughout the Theoretical Part, LMs are not even cohesive in a human-centered manner. Their cohesion is not motivated by meaning generation (which would converge on coherence), but by a task-driven notion of topicality. The result does not establish causality, but it provides strong descriptive evidence that lexical repetition cannot be treated as equivalent to semantic Topic Continuity.


#

```{r}

# Freeze plot-ready text-specific relationships

y4_x3pcw_scatterplot_data_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_y4_x3pcw_scatterplot_data.rds"
)

y4_x3pcw_plot_points <- y4_x3pcw_matched_transitions |>
  dplyr::mutate(
    text_label = dplyr::recode(
      model,
      human = "Human",
      chatgpt_native = "ChatGPT",
      claude_native = "Claude",
      gemini_native = "Gemini"
    ),
    text_label = factor(
      text_label,
      levels = c(
        "Human",
        "ChatGPT",
        "Claude",
        "Gemini"
      )
    )
  ) |>
  dplyr::arrange(
    text_label,
    y4_content_jaccard
  )

y4_x3pcw_fitted_lines <- y4_x3pcw_plot_points |>
  dplyr::filter(
    text_label %in% c(
      "Human",
      "ChatGPT"
    )
  ) |>
  dplyr::group_by(
    text_id,
    text_label
  ) |>
  dplyr::group_modify(
    ~ {
      relationship_model <- stats::lm(
        x3pcw_adjacent_paragraph_similarity ~
          y4_content_jaccard,
        data = .x
      )

      fitted_x <- seq(
        min(.x$y4_content_jaccard),
        max(.x$y4_content_jaccard),
        length.out = 100L
      )

      tibble::tibble(
        y4_content_jaccard = fitted_x,
        x3pcw_fitted = stats::predict(
          relationship_model,
          newdata = tibble::tibble(
            y4_content_jaccard = fitted_x
          )
        )
      )
    }
  ) |>
  dplyr::ungroup()

y4_x3pcw_scatterplot_data <- list(
  points = y4_x3pcw_plot_points,
  fitted_lines = y4_x3pcw_fitted_lines
)

saveRDS(
  y4_x3pcw_scatterplot_data,
  y4_x3pcw_scatterplot_data_path
)

y4_x3pcw_scatterplot_data <- readRDS(
  y4_x3pcw_scatterplot_data_path
)

stopifnot(
  nrow(y4_x3pcw_scatterplot_data$points) == 11L,
  nrow(y4_x3pcw_scatterplot_data$fitted_lines) == 200L,
  dplyr::n_distinct(
    y4_x3pcw_scatterplot_data$points$text_id
  ) == 4L
)

message(
  "Plot-ready y4–x3pcw relationships validated successfully."
)
```

# # MODUL 6.38d — Plot lexical overlap versus Topic Continuity

```{r}

```

#

```{r}
# MODUL 6.38e — Plot observed y4 content-word continuity waves

y4_x3pcw_matched_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_y4_x3pcw_matched_transitions.rds"
)

y4_content_wave_plot_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_y4_content_continuity_waves.png"
)

y4_content_wave_data <- readRDS(
  y4_x3pcw_matched_path
) |>
  dplyr::mutate(
    text_label = dplyr::recode(
      model,
      human = "Human",
      chatgpt_native = "ChatGPT",
      claude_native = "Claude",
      gemini_native = "Gemini"
    ),
    text_label = factor(
      text_label,
      levels = c(
        "Human",
        "ChatGPT",
        "Claude",
        "Gemini"
      )
    )
  ) |>
  dplyr::arrange(
    text_label,
    normalized_position
  )

modul_6_y4_content_continuity_waves <- ggplot2::ggplot(
  y4_content_wave_data,
  ggplot2::aes(
    x = normalized_position,
    y = y4_content_jaccard,
    group = text_label,
    colour = text_label
  )
) +
  ggplot2::geom_line(
    linewidth = 0.8
  ) +
  ggplot2::geom_point(
    size = 3,
    alpha = 0.9
  ) +
  ggplot2::facet_wrap(
    ~ text_label,
    ncol = 2,
    scales = "fixed"
  ) +
  ggplot2::scale_x_continuous(
    breaks = c(0, 0.5, 1),
    limits = c(0, 1)
  ) +
  ggplot2::scale_color_brewer(
    palette = "Set1"
  ) +
  ggplot2::labs(
    title = "Observed exact lexical continuity across paragraph transitions",
    subtitle = "Content-word Jaccard overlap for the four texts",
    x = "Normalized paragraph-transition position",
    y = "Content-word Jaccard overlap"
  ) +
  ggplot2::theme_minimal(
    base_size = 11
  ) +
  ggplot2::theme(
    legend.position = "none",
    panel.grid.minor = ggplot2::element_blank(),
    strip.text = ggplot2::element_text(
      face = "bold",
      size = 12
    )
  )

ggplot2::ggsave(
  filename = y4_content_wave_plot_path,
  plot = modul_6_y4_content_continuity_waves,
  width = 9,
  height = 7,
  dpi = 300
)

modul_6_y4_content_continuity_waves
```

# Caveat Note: Because the four texts contain different numbers of paragraphs, their adjacent-paragraph transitions are not directly equivalent on a one-to-one basis. Normalized transition position allows only a descriptive comparison of relative within-text development; it does not equalize the number or structural function of the paragraph boundaries. The x3pcw text means are therefore treated as the principal comparative measures, while the gradients and observed waves remain secondary descriptive diagnostics. For Claude and Gemini, which contain only two transitions, the gradients represent simple two-point differences rather than sustained developmental trajectories. + What remains strongly comparable is the shared measure "x3pcw_adjacent_paragraph_similarity_mean". It summarizes the overall meso-to-macro, near-global level of Topic Continuity despite different numbers of transitions. The gradient is secondary: for Human and ChatGPT it offers only limited directional diagnostics, and for Claude and Gemini it reduces to two-point differences only.


# 

```{r}

matched_object <- readRDS(
  y4_x3pcw_matched_path
)

gradient_object <- readRDS(
  x3pcw_gradient_path
)

class(matched_object)
names(matched_object)

class(gradient_object)
names(gradient_object)

```


#


```{r}
# MODUL 6.38f — Plot x3pcw Topic-Continuity trajectories

# MODUL 6.38e — Plot observed x3pcw Topic-Continuity trajectories

y4_x3pcw_matched_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_y4_x3pcw_matched_transitions.rds"
)

x3pcw_trajectory_plot_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_x3pcw_topic_continuity_trajectories.png"
)

x3pcw_trajectory_data <- readRDS(
  y4_x3pcw_matched_path
) |>
  dplyr::mutate(
    text_label = dplyr::recode(
      model,
      human = "Human",
      chatgpt_native = "ChatGPT",
      claude_native = "Claude",
      gemini_native = "Gemini"
    ),
    text_label = factor(
      text_label,
      levels = c(
        "Human",
        "ChatGPT",
        "Claude",
        "Gemini"
      )
    )
  ) |>
  dplyr::arrange(
    text_label,
    normalized_position
  )

stopifnot(
  nrow(x3pcw_trajectory_data) == 11L,
  dplyr::n_distinct(x3pcw_trajectory_data$text_id) == 4L,
  all(
    !is.na(
      x3pcw_trajectory_data$
        x3pcw_adjacent_paragraph_similarity
    )
  )
)

modul_6_x3pcw_topic_continuity_trajectories <- ggplot2::ggplot(
  x3pcw_trajectory_data,
  ggplot2::aes(
    x = normalized_position,
    y = x3pcw_adjacent_paragraph_similarity,
    colour = text_label,
    group = text_label
  )
) +
  ggplot2::geom_line(
    linewidth = 0.8
  ) +
  ggplot2::geom_point(
    size = 3,
    alpha = 0.9
  ) +
  ggplot2::facet_wrap(
    ~ text_label,
    ncol = 2,
    scales = "fixed"
  ) +
  ggplot2::scale_x_continuous(
    breaks = c(0, 0.5, 1),
    limits = c(0, 1)
  ) +
  ggplot2::scale_color_brewer(
    palette = "Set1"
  ) +
  ggplot2::labs(
    title = "Observed meso-level Topic Continuity across paragraph transitions",
    subtitle = "Content-word x3pcw similarity within each text",
    x = "Normalized paragraph-transition position",
    y = "Paragraph-level Topic Continuity",
    caption = paste(
      "The trajectories are descriptive.",
      "The texts contain different numbers of paragraph transitions."
    )
  ) +
  ggplot2::theme_minimal(
    base_size = 11
  ) +
  ggplot2::theme(
    legend.position = "none",
    panel.grid.minor = ggplot2::element_blank(),
    strip.text = ggplot2::element_text(
      face = "bold"
    )
  )

ggplot2::ggsave(
  filename = x3pcw_trajectory_plot_path,
  plot = modul_6_x3pcw_topic_continuity_trajectories,
  width = 9,
  height = 7,
  dpi = 300
)

modul_6_x3pcw_topic_continuity_trajectories
```


#

```{r}
# MODUL 6.38e — Overlapped Continuity Waves (Lexical y4 & Semantic x3pcw)
# MODUL 6.38e — Overlapped Continuity Waves (Lexical y4 & Semantic x3pcw)

library(ggplot2)
library(dplyr)

# 1. Load and prepare the shared data
y4_x3pcw_matched_path <- file.path(actual_lsa_output_dir, "modul_6_actual_lsa_y4_x3pcw_matched_transitions.rds")
combined_wave_plot_path <- file.path(actual_lsa_output_dir, "modul_6_actual_lsa_combined_continuity_waves.png")

combined_wave_data <- readRDS(y4_x3pcw_matched_path) |>
  dplyr::mutate(
    text_label = dplyr::recode(model, human = "Human", chatgpt_native = "ChatGPT", claude_native = "Claude", gemini_native = "Gemini"),
    text_label = factor(text_label, levels = c("Human", "ChatGPT", "Claude", "Gemini"))
  ) |>
  dplyr::arrange(text_label, normalized_position)

# 2. Define the sample-relative x3pcw Point d’Or
point_dor_level <- combined_wave_data |>
  dplyr::group_by(text_id) |>
  dplyr::summarise(x3pcw_text_mean = mean(x3pcw_adjacent_paragraph_similarity), .groups = "drop") |>
  dplyr::summarise(point_dor_level = max(x3pcw_text_mean)) |>
  dplyr::pull(point_dor_level)

stopifnot(length(point_dor_level) == 1L, is.finite(point_dor_level))

# 3. Build the combined overlap plot
modul_6_combined_continuity_waves <- ggplot2::ggplot(combined_wave_data, ggplot2::aes(x = normalized_position)) +

  # 3.1 Add the independent sample-relative Point d’Or reference
  ggplot2::geom_hline(yintercept = point_dor_level, linetype = "dotted", colour = "black", linewidth = 0.8, alpha = 0.6) +

  # 3.2 Add exact lexical continuity
  ggplot2::geom_line(ggplot2::aes(y = y4_content_jaccard, colour = text_label, linetype = "Lexical (y4 Jaccard)"), linewidth = 0.9) +
  ggplot2::geom_point(ggplot2::aes(y = y4_content_jaccard, colour = text_label), size = 2.5, alpha = 0.8) +

  # 3.3 Add latent semantic Topic Continuity
  ggplot2::geom_line(ggplot2::aes(y = x3pcw_adjacent_paragraph_similarity, colour = text_label, linetype = "Semantic (x3pcw LSA)"), linewidth = 0.9) +
  ggplot2::geom_point(ggplot2::aes(y = x3pcw_adjacent_paragraph_similarity, colour = text_label), size = 2.5, alpha = 0.8, shape = 17) +

  # 3.4 Separate the four texts and set fixed comparable axes
  ggplot2::facet_wrap(~ text_label, ncol = 2, scales = "fixed") +
  ggplot2::scale_x_continuous(breaks = c(0, 0.5, 1), limits = c(0, 1)) +
  ggplot2::scale_y_continuous(limits = c(-0.1, 1), breaks = seq(0, 1, 0.2)) +

  # 3.5 Set colours, line types, and theme
  ggplot2::scale_color_brewer(palette = "Set1") +
  ggplot2::scale_linetype_manual(values = c("Lexical (y4 Jaccard)" = "solid", "Semantic (x3pcw LSA)" = "dashed")) +
  ggplot2::guides(colour = "none") +
  ggplot2::theme_minimal(base_size = 11) +
  ggplot2::theme(
    legend.position = "bottom", legend.title = ggplot2::element_blank(), panel.grid.minor = ggplot2::element_blank(),
    strip.text = ggplot2::element_text(face = "bold", size = 12),
    plot.caption = ggplot2::element_text(hjust = 0, margin = ggplot2::margin(t = 8)),
    plot.margin = ggplot2::margin(t = 10, r = 15, b = 12, l = 15)
  ) +

  # 3.6 Add methodological labels and definitions
  ggplot2::labs(
    title = "Lexical vs. Latent Semantic Continuity across Paragraph Transitions",
    subtitle = "Solid lines show exact content-lemma recurrence (y4); dashed lines show semantic Topic Continuity (x3pcw)",
    x = "Normalized paragraph-transition position", y = "Observed continuity: cosine/lexical mass handover", linetype = NULL,
    caption = stringr::str_wrap(
      paste0(
        "y4 is the Jaccard proportion of exactly recurring content lemmas; ",
        "x3pcw is cosine similarity in the actual-LSA space. ",
        "Their raw magnitudes are not conceptually equivalent; the overlay ",
        "compares their within-text wave shapes and divergence. ",
        "The black dotted line marks the sample-relative x3pcw Point d’Or ",
        "reference: the highest observed x3pcw text mean (",
        formatC(point_dor_level, digits = 4, format = "f"),
        "), projected with a gradient of 0."
      ),
      width = 110
    )
  )

# 4. Save and display the final combined plot
ggplot2::ggsave(filename = combined_wave_plot_path, plot = modul_6_combined_continuity_waves, width = 9, height = 7, dpi = 300)
modul_6_combined_continuity_waves







 
 
  
   

```


#

```{r}
stopifnot(file.exists(combined_wave_plot_path))
message("Plot saved to: ", combined_wave_plot_path)
```


# # # MODUL 6.39 — Calculate title-based x4 similarities

```{r}
# MODUL 6.39 — Calculate title-based x4 similarities

main_title_id <- all_target_segments |>
  dplyr::filter(segment_level == "main_title") |>
  dplyr::pull(segment_id)

subheading_lookup <- all_target_segments |>
  dplyr::filter(segment_level == "subheading") |>
  dplyr::arrange(segment_order) |>
  dplyr::transmute(
    section_id = dplyr::row_number(),
    subheading_segment_id = segment_id
  )

stopifnot(length(main_title_id) == 1L, nrow(subheading_lookup) == 3L)

segment_cosine <- function(id_1, id_2) {
  vector_1 <- target_segment_vectors[id_1, , drop = TRUE]
  vector_2 <- target_segment_vectors[id_2, , drop = TRUE]
  if (anyNA(vector_1) || anyNA(vector_2)) NA_real_ else sum(vector_1 * vector_2)
}

x4_title_pairs <- all_target_segments |>
  dplyr::filter(segment_level %in% c("sentence", "paragraph", "whole_body_text")) |>
  dplyr::mutate(
    cosine_similarity = purrr::map_dbl(segment_id, ~ segment_cosine(.x, main_title_id))
  )

x4_title_profiles <- x4_title_pairs |>
  dplyr::group_by(text_id, model, segment_level) |>
  dplyr::summarise(similarity = mean(cosine_similarity), .groups = "drop") |>
  dplyr::mutate(
    measure = dplyr::recode(
      segment_level,
      sentence = "x4_sentence_title_similarity_mean",
      paragraph = "x4_paragraph_title_similarity_mean",
      whole_body_text = "x4_text_title_similarity"
    )
  ) |>
  dplyr::select(-segment_level) |>
  tidyr::pivot_wider(names_from = measure, values_from = similarity)

prompt_subheading_title_pairs <- subheading_lookup |>
  dplyr::mutate(
    cosine_similarity = purrr::map_dbl(
      subheading_segment_id,
      ~ segment_cosine(.x, main_title_id)
    )
  )

prompt_subheading_title_baseline <- prompt_subheading_title_pairs |>
  dplyr::summarise(
    prompt_subheading_title_similarity_mean = mean(cosine_similarity)
  )

x4_title_profiles
prompt_subheading_title_baseline
```
# Note: Sentence → title mean
Each body sentence was compared with the shared main-title vector using cosine similarity. The sentence similarities were then averaged separately for each text.
Paragraph → title mean
Each original body paragraph was compared with the same main-title vector. The paragraph similarities were then averaged separately for each text.
Whole text → title
The complete body text was represented by one LSA vector and compared once with the main-title vector. Therefore, this is a single cosine similarity, not a mean of sentence or paragraph values.

#

```{r}
# MODUL 6.39a — Calculate merged-section-to-subheading similarities

x4_section_segments <- paragraph_segment_table |>
  dplyr::arrange(text_id, section_id, paragraph_number) |>
  dplyr::group_by(text_id, model, section_id) |>
  dplyr::summarise(
    n_paragraphs = dplyr::n(),
    lsa_lemma_text = paste(lsa_lemma_text, collapse = " "),
    .groups = "drop"
  ) |>
  dplyr::mutate(
    section_segment_id = paste0(text_id, "_X4_SECTION_", section_id)
  ) |>
  dplyr::left_join(
    subheading_lookup,
    by = "section_id"
  )

section_dfm <- quanteda::dfm(
  quanteda::tokens(
    stats::setNames(
      x4_section_segments$lsa_lemma_text,
      x4_section_segments$section_segment_id
    ),
    what = "fastestword"
  )
) |>
  quanteda::dfm_match(features = lsa_reference_space$vocabulary)

section_matrix <- methods::as(section_dfm, "dgCMatrix")
section_matrix@x <- log1p(section_matrix@x)

x4_section_vectors <- as.matrix(
  section_matrix %*%
    Matrix::Diagonal(x = lsa_reference_space$idf) %*%
    lsa_reference_space$right_singular_vectors %*%
    Matrix::Diagonal(x = 1 / lsa_reference_space$singular_values)
)

section_norms <- sqrt(rowSums(x4_section_vectors^2))
x4_section_vectors <- sweep(x4_section_vectors, 1, section_norms, "/")
x4_section_vectors[!is.finite(x4_section_vectors)] <- NA_real_

section_cosine <- function(section_id, subheading_id) {
  section_vector <- x4_section_vectors[section_id, , drop = TRUE]
  subheading_vector <- target_segment_vectors[subheading_id, , drop = TRUE]
  if (anyNA(section_vector) || anyNA(subheading_vector)) NA_real_
  else sum(section_vector * subheading_vector)
}

x4_section_subheading_pairs <- x4_section_segments |>
  dplyr::mutate(
    cosine_similarity = purrr::map2_dbl(
      section_segment_id,
      subheading_segment_id,
      section_cosine
    )
  )

x4_subheading_profiles <- x4_section_subheading_pairs |>
  dplyr::group_by(text_id, model) |>
  dplyr::summarise(
    n_sections_scored = dplyr::n_distinct(section_id),
    section_similarity_mean = mean(cosine_similarity),
    .groups = "drop"
  ) |>
  dplyr::left_join(
    structural_audit |>
      dplyr::select(text_id, section_mapping_valid),
    by = "text_id"
  ) |>
  dplyr::mutate(
    x4_paragraph_subheading_similarity_mean = dplyr::if_else(
      section_mapping_valid & n_sections_scored == 3L,
      section_similarity_mean,
      NA_real_
    )
  ) |>
  dplyr::select(
    text_id,
    model,
    x4_paragraph_subheading_similarity_mean,
    n_sections_scored,
    section_mapping_valid
  )

x4_section_subheading_pairs
x4_subheading_profiles
```
# Note: Main new measure - x4_paragraph_subheading_similarity_mean -> For each text:

merged Section 1 was compared with Subheading 1;
merged Section 2 was compared with Subheading 2;
merged Section 3 was compared with Subheading 3; the three cosine similarities were averaged equally. 

Section 1 → Subheading 1 = s1
Section 2 → Subheading 2 = s2
Section 3 → Subheading 3 = s3

Each comparison contributes exactly one third to the final value:=> s1+s2+s3/3

Caveat note:The paragraph-to-subheading similarity should be interpreted cautiously because the subheadings may have been added after the body text had already been written. 

(In that case, the measure captures the retrospective semantic fit between each section and its assigned subheading, rather than necessarily showing whether the text was originally developed in response to/toward that subheading.)
#

```{r}
stopifnot(
  all(c(
    "text_id",
    "model",
    "section_id",
    "section_segment_id",
    "subheading_segment_id"
  ) %in% names(x4_section_segments))
)

print(x4_section_segments, n = Inf, width = Inf)
```


# MODUL 6.39b — Assemble, freeze, and validate title and subheading x4

```{r}

# MODUL 6.39b — Assemble, freeze, and validate title and subheading x4

x4_output_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_x4_title_subheading.rds"
)

x4_title_subheading_profiles <- x4_title_profiles |>
  dplyr::left_join(
    x4_subheading_profiles,
    by = c("text_id", "model")
  ) |>
  dplyr::arrange(text_id)

x4_title_subheading_bundle <- list(
  title_pairs = x4_title_pairs,
  section_segments = x4_section_segments,
  section_vectors = x4_section_vectors,
  section_subheading_pairs = x4_section_subheading_pairs,
  profiles = x4_title_subheading_profiles,
  prompt_baseline_pairs = prompt_subheading_title_pairs,
  prompt_baseline = prompt_subheading_title_baseline
)

saveRDS(x4_title_subheading_bundle, x4_output_path, version = 3)
x4_title_subheading_bundle <- readRDS(x4_output_path)

x4_title_subheading_profiles <- x4_title_subheading_bundle$profiles
prompt_subheading_title_baseline <- x4_title_subheading_bundle$prompt_baseline

stopifnot(
  nrow(x4_title_subheading_profiles) == 4L,
  nrow(x4_title_subheading_bundle$title_pairs) == 125L,
  nrow(x4_title_subheading_bundle$prompt_baseline_pairs) == 3L,
  all(
    x4_title_subheading_profiles$section_mapping_valid |
      is.na(
        x4_title_subheading_profiles$
          x4_paragraph_subheading_similarity_mean
      )
  )
)

readr::write_csv(
  x4_title_subheading_profiles,
  file.path(
    actual_lsa_output_dir,
    "modul_6_actual_lsa_x4_title_subheading_profiles.csv"
  )
)

message("MODUL 6.39 title and subheading x4 outputs frozen and reloaded.")

print(x4_title_subheading_profiles, n = Inf, width = Inf)
print(prompt_subheading_title_baseline, width = Inf)
```
# Note: These numbers say, how well each model and the human baseline “sticks to its own headings” at different structural levels. Interpreting the columns as LSA‑based semantic similarity between each unit and its title (higher = more on‑topic):

Whole text → title: All four have the highest values here (0.26–0.46), so at the macro level every text broadly fits its overall title. The human text is strongest (0.4571), followed by Claude, then ChatGPT, with Gemini lowest.

Paragraph → title/ subtitle: Again, the human text is at or near the top (0.3191 paragraph, 0.0317 subheading; note that the subheading value is small for everyone), with Claude and ChatGPT close behind and Gemini consistently lower. This suggests that human paragraphs and sections are more tightly semantically aligned with their titles than the LM ones, especially compared to Gemini.

Sentence → title: At the micro level, ChatGPT (0.1551) and the human text (0.1375) are quite close, with Claude slightly lower and Gemini lowest. That is, individual human and ChatGPT sentences are more semantically related to the text title than Claude’s or Gemini’s are.

Put simply: across all structural levels, the human text shows the strongest and most consistent semantic anchoring to its titles, Claude is generally next, ChatGPT slightly behind but still reasonably aligned, and Gemini is weakest, especially at the paragraph, whole‑text, and sentence levels. This supports the view that humans keep their discourse more tightly organised around the announced topics, while Gemini didn't manage to stick to the prompted headings.
# ===========================================================

# MODUL 6.40a — Load and validate frozen seed-word inputs

```{r}
actual_lsa_output_dir <- file.path("..", "measure_outputs")

expected_seed_words <- readLines(
  file.path("..", "seeds", "seed_words_every_9th_text1_hm.txt"),
  encoding = "UTF-8",
  warn = FALSE
)

all_lemma_annotation_layer <- readRDS(
  file.path("..", "data_prep_CCP_output", "current_texts_parsed.rds")
)

actual_lsa_reference_space <- readRDS(
  file.path(
    actual_lsa_output_dir,
    "modul_6_actual_lsa_reference_space.rds"
  )
)

lsa_pos_keep <- c(
  "NOUN", "PROPN", "VERB", "ADJ",
  "ADV", "CCONJ", "SCONJ"
)

stopifnot(
  length(expected_seed_words) == 61L,
  file.exists(
    file.path(
      actual_lsa_output_dir,
      "modul_6_actual_lsa_reference_space.rds"
    )
  )
)
tibble::tibble(
  frozen_input = c(
    "Expected seed words",
    "All-lemma annotations",
    "Actual-LSA vocabulary",
    "Retained UPOS categories"
  ),
  n = c(
    length(expected_seed_words),
    nrow(all_lemma_annotation_layer),
    length(actual_lsa_reference_space$vocabulary),
    length(lsa_pos_keep)
  )
)
```


# MODUL 6.40b — Prepare expected seed-word and annotation layers

```{r}

# MODUL 6.40b — Prepare expected seed-word and annotation layers

normalize_sw <- function(x) {
  x |>
    as.character() |>
    stringr::str_to_lower() |>
    stringr::str_squish() |>
    stringr::str_replace_all("\\*", "") |>
    stringr::str_replace_all("^[^\\p{L}\\p{N}]+|[^\\p{L}\\p{N}]+$", "")
}

sw_expected <- tibble::tibble(
  sw_position = seq_along(expected_seed_words),
  expected_seed_word = expected_seed_words,
  expected_seed_word_norm = normalize_sw(expected_seed_words)
)

sw_annotation <- all_lemma_annotation_layer |>
  dplyr::arrange(text_id, paragraph_id, sentence_id, as.integer(token_id)) |>
  dplyr::group_by(text_id) |>
  dplyr::mutate(token_position = dplyr::row_number()) |>
  dplyr::ungroup() |>
  dplyr::transmute(
    text_id,
    model = dplyr::case_when(
      text_id == "HUMAN_BASELINE_001" ~ "Human",
      stringr::str_detect(text_id, "CHATGPT") ~ "ChatGPT",
      stringr::str_detect(text_id, "CLAUDE") ~ "Claude",
      TRUE ~ NA_character_
    ),
    paragraph_id,
    sentence_id,
    token_id,
    token_position,
    token,
    token_norm = normalize_sw(token),
    lemma = normalize_sw(lemma),
    pos = stringr::str_to_upper(as.character(pos))
  ) |>
  dplyr::filter(model %in% c("ChatGPT", "Claude", "Human"))

stopifnot(
  nrow(sw_expected) == 61L,
  setequal(unique(sw_annotation$model), c("ChatGPT", "Claude", "Human")),
  all(c("lemma", "pos") %in% names(sw_annotation))
)

tibble::tibble(
  layer = c("Expected seed words", "Retained annotation rows"),
  n = c(nrow(sw_expected), nrow(sw_annotation))
)
    
```


# MODUL 6.40c — Calculate expected seed-word diagnostics

```{r}

# MODUL 6.40c — Calculate expected seed-word diagnostics

sw_expected_match <- sw_expected |>
  dplyr::group_by(expected_seed_word_norm) |>
  dplyr::mutate(sw_occurrence = dplyr::row_number()) |>
  dplyr::ungroup()

sw_annotation_match <- sw_annotation |>
  dplyr::group_by(text_id, token_norm) |>
  dplyr::mutate(sw_occurrence = dplyr::row_number()) |>
  dplyr::ungroup()

expected_seed_word_rows <- tidyr::crossing(
  sw_annotation |> dplyr::distinct(text_id, model),
  sw_expected_match
) |>
  dplyr::left_join(
    sw_annotation_match |>
      dplyr::select(text_id, token_norm, sw_occurrence, token_position, token, lemma, pos),
    by = c("text_id", "expected_seed_word_norm" = "token_norm", "sw_occurrence")
  ) |>
  dplyr::mutate(
    sw_detected = !is.na(token_position),
    sw_annotated = sw_detected & !is.na(lemma) & !is.na(pos),
    sw_lsa_relevant = sw_annotated & pos %in% lsa_pos_keep,
    sw_in_model = sw_lsa_relevant & lemma %in% actual_lsa_reference_space$vocabulary,
    sw_oov = sw_lsa_relevant & !sw_in_model
  ) |>
  dplyr::arrange(text_id, sw_position)

expected_seed_word_diagnostics <- expected_seed_word_rows |>
  dplyr::filter(model %in% c("ChatGPT", "Claude")) |>
  dplyr::group_by(text_id, model) |>
  dplyr::summarise(
    n_sw_expected = dplyr::n(),
    n_sw_detected = sum(sw_detected),
    n_sw_annotated = sum(sw_annotated),
    n_sw_in_model = sum(sw_in_model),
    n_sw_oov = sum(sw_oov),
    .groups = "drop"
  )

human_seed_words_in_model <- expected_seed_word_rows |>
  dplyr::filter(model == "Human", sw_in_model) |>
  dplyr::select(sw_position, expected_seed_word, token_position, lemma, pos)

stopifnot(
  nrow(expected_seed_word_rows) == 183L,
  nrow(expected_seed_word_diagnostics) == 2L,
  all(expected_seed_word_diagnostics$n_sw_expected == 61L),
  sum(expected_seed_word_rows$sw_detected[expected_seed_word_rows$model == "Human"]) == 61L
)

expected_seed_word_diagnostics |>
  dplyr::mutate(
    dplyr::across(
      c(n_sw_detected, n_sw_annotated, n_sw_in_model, n_sw_oov),
      \(x) paste0(x, " / 61")
    )
  )
```

#

```{r}

```



# MODUL 6.40e — Freeze and reload x4 seed-word diagnostics
This freezes the complete reusable MODUL 6.40 bundle. The diagnostics table later enters the x4-associated profile, while human_seed_words_in_model supplies the Human-derived seed associative net component for MODUL 6.41.

```{r}
# MODUL 6.40e — Freeze and reload x4 seed-word diagnostics

expected_seed_word_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_x4_expected_seed_word_diagnostics.rds"
)

expected_seed_word_bundle <- list(
  rows = expected_seed_word_rows |>
    dplyr::filter(model %in% c("ChatGPT", "Claude")),
  diagnostics = expected_seed_word_diagnostics,
  human_seed_words_in_model = human_seed_words_in_model
)

saveRDS(expected_seed_word_bundle, expected_seed_word_path)
expected_seed_word_bundle <- readRDS(expected_seed_word_path)

expected_seed_word_rows <- expected_seed_word_bundle$rows
expected_seed_word_diagnostics <- expected_seed_word_bundle$diagnostics
human_seed_words_in_model <- expected_seed_word_bundle$human_seed_words_in_model

stopifnot(
  nrow(expected_seed_word_rows) == 122L,
  nrow(expected_seed_word_diagnostics) == 2L,
  all(expected_seed_word_diagnostics$n_sw_expected == 61L),
  nrow(human_seed_words_in_model) > 0L,
  all(human_seed_words_in_model$pos %in% lsa_pos_keep),
  all(human_seed_words_in_model$lemma %in% actual_lsa_reference_space$vocabulary)
)

expected_seed_word_diagnostics |>
  dplyr::mutate(
    dplyr::across(
      c(n_sw_detected, n_sw_annotated, n_sw_in_model, n_sw_oov),
      \(x) paste0(x, " / 61")
    )
  ) |>
  print(n = Inf)

message("Expected seed-word diagnostics frozen and reloaded successfully.")
```
# Meaning of the MODUL 6.40 diagnostics
n_sw_expected: all 61 seed-word occurrences prescribed by the prompt.
n_sw_detected: prescribed seed-word occurrences actually found in the text.
n_sw_annotated: detected occurrences with both lemma and POS annotation.
n_sw_in_model: detected and annotated occurrences with an eligible POS whose lemma exists in the actual-LSA vocabulary. These exact rows can form the seed vector.
n_sw_oov: detected and LSA-relevant occurrences whose lemmas are outside the actual-LSA vocabulary.
# Note: 
n_sw_detected is used within MODUL 6.40 as an exploratory x4-associated diagnostic of expected seed-word presence. The retained values are 48/61 for ChatGPT and 50/61 for Claude. Gemini is excluded because retained Markdown emphasis markers systematically corrupted parts of its token and POS annotation. The Human baseline is excluded from the comparative diagnostic because it is the source of the expected seed-word list and supplies the Human-derived seed component for MODUL 6.41. n_sw_detected remains separate from x4_text_prompt_induced_ass_net_similarity and must not be averaged with or added to the cosine score calculated in MODUL 6.42. During final integration, values are available only for ChatGPT and Claude; Gemini and Human should remain missing for this diagnostic, with their respective exclusion reasons documented.
# Because n_sw_detected equals n_sw_annotated for both models, there is no annotation failure among the detected words. Instead:

ChatGPT omitted 13 required seed-word occurrences.
Claude omitted 11 required seed-word occurrences.
Of the detected words, only 25 and 27 respectively can enter the actual-LSA vector because the remaining detected words have excluded POS categories or are OOV.
# ===================================

# MODUL 6.41a — Construct and freeze the Human-derived prompt-induced associative net

```{r}
# MODUL 6.41a — Construct and freeze the Human-derived prompt-induced associative net

prompt_induced_ass_net_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_prompt_induced_associative_net.rds"
)

# 1. Reload the frozen Human seed component

expected_seed_word_bundle <- readRDS(file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_x4_expected_seed_word_diagnostics.rds"
))

human_seed_words_in_model <- expected_seed_word_bundle$human_seed_words_in_model

stopifnot(
  is.data.frame(human_seed_words_in_model),
  "lemma" %in% names(human_seed_words_in_model),
  nrow(human_seed_words_in_model) > 0L
)

# 2. Use the validated heading lemmas

heading_terms <- c(
  "digitalisierung", "arbeitswelt",
  "digitalisierung", "sorge", "arbeitsplatz",
  "fortschritt", "erfordern", "anpassung",
  "menschlich", "qualität", "erfolgreich", "zukunft"
)

human_seed_terms <- human_seed_words_in_model |>
  dplyr::pull(lemma) |>
  as.character() |>
  stringr::str_to_lower() |>
  stringr::str_squish()

stopifnot(
  !anyNA(human_seed_terms),
  all(human_seed_terms != ""),
  all(human_seed_terms %in% actual_lsa_reference_space$vocabulary)
)

# 3. Combine headings and Human seed lemmas before projection

prompt_induced_ass_net_terms <- c(heading_terms, human_seed_terms)

prompt_induced_ass_net_dfm <- quanteda::tokens(c(
  prompt_induced_associative_net =
    paste(prompt_induced_ass_net_terms, collapse = " ")
)) |>
  quanteda::dfm() |>
  quanteda::dfm_match(features = actual_lsa_reference_space$vocabulary) |>
  methods::as("dgCMatrix")

n_prompt_terms_in_model <- as.integer(sum(prompt_induced_ass_net_dfm))

# 4. Apply one log-TF-IDF actual-LSA projection

prompt_induced_ass_net_dfm@x <- log1p(prompt_induced_ass_net_dfm@x)

prompt_induced_ass_net_tfidf <- prompt_induced_ass_net_dfm %*%
  Matrix::Diagonal(x = actual_lsa_reference_space$idf)

prompt_fold_in_matrix <- actual_lsa_reference_space$right_singular_vectors %*%
  Matrix::Diagonal(x = 1 / actual_lsa_reference_space$singular_values)

prompt_induced_associative_net_raw <- as.numeric(
  prompt_induced_ass_net_tfidf %*% prompt_fold_in_matrix
)

prompt_induced_ass_net_raw_norm <- sqrt(
  sum(prompt_induced_associative_net_raw^2)
)

stopifnot(
  is.finite(prompt_induced_ass_net_raw_norm),
  prompt_induced_ass_net_raw_norm > 0
)

prompt_induced_associative_net <- prompt_induced_associative_net_raw /
  prompt_induced_ass_net_raw_norm

# 5. Audit and freeze

prompt_induced_ass_net_audit <- tibble::tibble(
  reference = "Human-derived heading and seed-word associative net",
  t_subheadings_bag = length(heading_terms),
  sw_bag = length(human_seed_terms),
  assoc_net_bag = length(prompt_induced_ass_net_terms),
  assoc_net_lsa = n_prompt_terms_in_model,
  assoc_net_oov = length(prompt_induced_ass_net_terms) - n_prompt_terms_in_model,
  lsa_dimensions = length(prompt_induced_associative_net),
  hum_fix_assoc_net_l2_norm = sqrt(sum(prompt_induced_associative_net^2))
)

prompt_induced_ass_net_bundle <- list(
  heading_terms = heading_terms,
  human_seed_words_in_model = human_seed_words_in_model,
  human_seed_terms = human_seed_terms,
  pooled_terms = prompt_induced_ass_net_terms,
  vector = prompt_induced_associative_net,
  audit = prompt_induced_ass_net_audit
)

saveRDS(prompt_induced_ass_net_bundle, prompt_induced_ass_net_path)

# 6. Reload and validate

prompt_induced_ass_net_bundle <- readRDS(prompt_induced_ass_net_path)
prompt_induced_associative_net <- prompt_induced_ass_net_bundle$vector
prompt_induced_ass_net_audit <- prompt_induced_ass_net_bundle$audit

stopifnot(
  length(prompt_induced_associative_net) == 50L,
  all(is.finite(prompt_induced_associative_net)),
  abs(sum(prompt_induced_associative_net^2) - 1) < 1e-10
)

print(prompt_induced_ass_net_audit, n = Inf, width = Inf)

message(
  "Human-derived prompt-induced associative net frozen and reloaded successfully."
)
```

# MODUL 6.41b — Assemble the ChatGPT and Claude associative-net bags

```{r}
# MODUL 6.41b — Assemble the ChatGPT and Claude associative-net bags



chatgpt_assoc_net_bag <- c(
  heading_terms,
  expected_seed_word_bundle$rows |>
    dplyr::filter(model == "ChatGPT", sw_in_model) |>
    dplyr::pull(lemma) |>
    as.character() |>
    stringr::str_to_lower() |>
    stringr::str_squish()
)

claude_assoc_net_bag <- c(
  heading_terms,
  expected_seed_word_bundle$rows |>
    dplyr::filter(model == "Claude", sw_in_model) |>
    dplyr::pull(lemma) |>
    as.character() |>
    stringr::str_to_lower() |>
    stringr::str_squish()
)

stopifnot(
  length(chatgpt_assoc_net_bag) == 37L,
  length(claude_assoc_net_bag) == 39L,
  all(chatgpt_assoc_net_bag[-seq_along(heading_terms)] %in% actual_lsa_reference_space$vocabulary),
  all(claude_assoc_net_bag[-seq_along(heading_terms)] %in% actual_lsa_reference_space$vocabulary)
)

print(c(
  chatgpt_assoc_net_bag = length(chatgpt_assoc_net_bag),
  claude_assoc_net_bag = length(claude_assoc_net_bag)
))
```

# Inspection =  object names overview

```{r}
names(expected_seed_word_bundle)

utils::str(
  expected_seed_word_bundle,
  max.level = 3,
  list.len = 100
)
```

# MODUL 6.41c — Project and normalize the ChatGPT and Claude associative nets

```{r}
# MODUL 6.41c — Project and normalize the ChatGPT and Claude associative nets

lm_assoc_net_dfm <- quanteda::tokens(c(
  ChatGPT = paste(chatgpt_assoc_net_bag, collapse = " "),
  Claude = paste(claude_assoc_net_bag, collapse = " ")
)) |>
  quanteda::dfm() |>
  quanteda::dfm_match(features = actual_lsa_reference_space$vocabulary) |>
  methods::as("dgCMatrix")

lm_assoc_net_dfm@x <- log1p(lm_assoc_net_dfm@x)

lm_assoc_net_raw <- as.matrix(
  lm_assoc_net_dfm %*%
    Matrix::Diagonal(x = actual_lsa_reference_space$idf) %*%
    prompt_fold_in_matrix
)

lm_assoc_net_raw <- lm_assoc_net_raw / sqrt(rowSums(lm_assoc_net_raw^2))

ChatGPT_assoc_net_vector <- as.numeric(lm_assoc_net_raw["ChatGPT", ])
Claude_assoc_net_vector <- as.numeric(lm_assoc_net_raw["Claude", ])
ChatGPT_assoc_net_l2_norm <- sqrt(sum(ChatGPT_assoc_net_vector^2))
Claude_assoc_net_l2_norm <- sqrt(sum(Claude_assoc_net_vector^2))

stopifnot(
  length(ChatGPT_assoc_net_vector) == 50L,
  length(Claude_assoc_net_vector) == 50L,
  abs(ChatGPT_assoc_net_l2_norm - 1) < 1e-10,
  abs(Claude_assoc_net_l2_norm - 1) < 1e-10
)

print(c(
  ChatGPT_assoc_net_l2_norm = ChatGPT_assoc_net_l2_norm,
  Claude_assoc_net_l2_norm = Claude_assoc_net_l2_norm
))
```

# MODUL 6.41d — Audit, freeze, and reload the three associative nets

```{r}
# MODUL 6.41d — Audit, freeze, and reload the three associative nets

prompt_induced_ass_net_bundle$chatgpt_assoc_net_bag <- chatgpt_assoc_net_bag
prompt_induced_ass_net_bundle$claude_assoc_net_bag <- claude_assoc_net_bag
prompt_induced_ass_net_bundle$ChatGPT_assoc_net_vector <- ChatGPT_assoc_net_vector
prompt_induced_ass_net_bundle$Claude_assoc_net_vector <- Claude_assoc_net_vector
prompt_induced_ass_net_bundle$ChatGPT_assoc_net_l2_norm <- ChatGPT_assoc_net_l2_norm
prompt_induced_ass_net_bundle$Claude_assoc_net_l2_norm <- Claude_assoc_net_l2_norm

prompt_induced_ass_net_bundle$audit <- prompt_induced_ass_net_bundle$audit |>
  dplyr::mutate(
    chatgpt_assoc_net_bag = length(.env$chatgpt_assoc_net_bag),
    claude_assoc_net_bag = length(.env$claude_assoc_net_bag),
    ChatGPT_assoc_net_l2_norm = .env$ChatGPT_assoc_net_l2_norm,
    Claude_assoc_net_l2_norm = .env$Claude_assoc_net_l2_norm
  )

saveRDS(prompt_induced_ass_net_bundle, prompt_induced_ass_net_path)
prompt_induced_ass_net_bundle <- readRDS(prompt_induced_ass_net_path)

prompt_induced_associative_net <- prompt_induced_ass_net_bundle$vector
ChatGPT_assoc_net_vector <- prompt_induced_ass_net_bundle$ChatGPT_assoc_net_vector
Claude_assoc_net_vector <- prompt_induced_ass_net_bundle$Claude_assoc_net_vector
prompt_induced_ass_net_audit <- prompt_induced_ass_net_bundle$audit

stopifnot(
  all(lengths(list(
    prompt_induced_associative_net,
    ChatGPT_assoc_net_vector,
    Claude_assoc_net_vector
  )) == 50L),
  all(abs(c(
    prompt_induced_ass_net_bundle$hum_fix_assoc_net_l2_norm,
    prompt_induced_ass_net_bundle$ChatGPT_assoc_net_l2_norm,
    prompt_induced_ass_net_bundle$Claude_assoc_net_l2_norm
  ) - 1) < 1e-10)
)

print(prompt_induced_ass_net_audit, n = Inf, width = Inf)
message("Human, ChatGPT, and Claude associative nets frozen and reloaded successfully.")
```

# MODUL 6.41e — Calculate, freeze, and reload associative-net similarities

```{r}
# MODUL 6.41e — Calculate, freeze, and reload associative-net similarities

chatgpt_human_assoc_net_similarity <- sum(ChatGPT_assoc_net_vector * prompt_induced_associative_net)
claude_human_assoc_net_similarity <- sum(Claude_assoc_net_vector * prompt_induced_associative_net)
chatgpt_claude_assoc_net_similarity <- sum(ChatGPT_assoc_net_vector * Claude_assoc_net_vector)

prompt_induced_ass_net_bundle$chatgpt_human_assoc_net_similarity <- chatgpt_human_assoc_net_similarity
prompt_induced_ass_net_bundle$claude_human_assoc_net_similarity <- claude_human_assoc_net_similarity
prompt_induced_ass_net_bundle$chatgpt_claude_assoc_net_similarity <- chatgpt_claude_assoc_net_similarity

saveRDS(prompt_induced_ass_net_bundle, prompt_induced_ass_net_path)
prompt_induced_ass_net_bundle <- readRDS(prompt_induced_ass_net_path)

chatgpt_human_assoc_net_similarity <- prompt_induced_ass_net_bundle$chatgpt_human_assoc_net_similarity
claude_human_assoc_net_similarity <- prompt_induced_ass_net_bundle$claude_human_assoc_net_similarity
chatgpt_claude_assoc_net_similarity <- prompt_induced_ass_net_bundle$chatgpt_claude_assoc_net_similarity

tibble::tibble(
  comparison = c("ChatGPT–Human", "Claude–Human", "ChatGPT–Claude"),
  associative_net_similarity = c(
    chatgpt_human_assoc_net_similarity,
    claude_human_assoc_net_similarity,
    chatgpt_claude_assoc_net_similarity
  )
)
```

# MODUL 6.41f — Calculate and freeze seed-word-net similarities

```{r}

# MODUL 6.41f — Calculate and freeze seed-word-net similarities

seed_word_net_dfm <- quanteda::tokens(c(
  Human = paste(prompt_induced_ass_net_bundle$human_seed_terms, collapse = " "),
  ChatGPT = paste(prompt_induced_ass_net_bundle$chatgpt_assoc_net_bag[-seq_along(prompt_induced_ass_net_bundle$heading_terms)], collapse = " "),
  Claude = paste(prompt_induced_ass_net_bundle$claude_assoc_net_bag[-seq_along(prompt_induced_ass_net_bundle$heading_terms)], collapse = " ")
)) |>
  quanteda::dfm() |>
  quanteda::dfm_match(features = actual_lsa_reference_space$vocabulary) |>
  methods::as("dgCMatrix")

stopifnot(
  Matrix::rowSums(seed_word_net_dfm) == c(Human = 35L, ChatGPT = 25L, Claude = 27L)
)

seed_word_net_dfm@x <- log1p(seed_word_net_dfm@x)

seed_word_net_vectors <- as.matrix(
  seed_word_net_dfm %*%
    Matrix::Diagonal(x = actual_lsa_reference_space$idf) %*%
    prompt_fold_in_matrix
)

seed_word_net_vectors <- sweep(
  seed_word_net_vectors, 1,
  sqrt(rowSums(seed_word_net_vectors^2)), "/"
)

seed_word_net_similarities <- tibble::tibble(
  comparison = c("ChatGPT–Human", "Claude–Human", "ChatGPT–Claude"),
  seed_word_net_similarity = c(
    sum(seed_word_net_vectors["ChatGPT", ] * seed_word_net_vectors["Human", ]),
    sum(seed_word_net_vectors["Claude", ] * seed_word_net_vectors["Human", ]),
    sum(seed_word_net_vectors["ChatGPT", ] * seed_word_net_vectors["Claude", ])
  )
)

prompt_induced_ass_net_bundle$seed_word_net_vectors <- seed_word_net_vectors
prompt_induced_ass_net_bundle$seed_word_net_similarities <- seed_word_net_similarities

saveRDS(prompt_induced_ass_net_bundle, prompt_induced_ass_net_path)
prompt_induced_ass_net_bundle <- readRDS(prompt_induced_ass_net_path)

stopifnot(
  all(abs(sqrt(rowSums(prompt_induced_ass_net_bundle$seed_word_net_vectors^2)) - 1) < 1e-10)
)

print(prompt_induced_ass_net_bundle$seed_word_net_similarities, n = Inf)
message("Seed-word-net similarities frozen and reloaded successfully.")

```
# Note: 
The seed-word list was fixed at 61 expected seed words for every text. However, not every LM implemented all 61 seed words, resulting in detected seed-word subsets of 48 for ChatGPT and 50 for Claude. In addition, only seed words belonging to the selected LSA categories—NOUN, PROPN, VERB, ADJ, ADV, CCONJ, and SCONJ—could enter the LSA projection, while many function words were excluded by design. The remaining seed words also had to be represented in the frozen actual-LSA vocabulary. Consequently, the final Seed-word nets are based on 35 seed words in model for Human, 25 for ChatGPT, and 27 for Claude. This reduction is therefore mainly due to incomplete LM implementation, POS-based LSA selection, and OOV cases, rather than an additional annotation loss.

# MODUL 6.41g — Calculate and freeze whole-text similarities

```{r}
# MODUL 6.41g — Calculate and freeze whole-text similarities

whole_text_vectors <- target_segment_vectors[
  whole_body_text_segment_table$segment_id, , drop = FALSE
]
rownames(whole_text_vectors) <- whole_body_text_segment_table$model

whole_text_similarities <- tibble::tibble(
  comparison = c(
    "ChatGPT–Human", "Claude–Human", "Gemini–Human",
    "ChatGPT–Claude", "ChatGPT–Gemini", "Claude–Gemini"
  ),
  whole_text_similarity = c(
    sum(whole_text_vectors["chatgpt_native", ] * whole_text_vectors["human", ]),
    sum(whole_text_vectors["claude_native", ] * whole_text_vectors["human", ]),
    sum(whole_text_vectors["gemini_native", ] * whole_text_vectors["human", ]),
    sum(whole_text_vectors["chatgpt_native", ] * whole_text_vectors["claude_native", ]),
    sum(whole_text_vectors["chatgpt_native", ] * whole_text_vectors["gemini_native", ]),
    sum(whole_text_vectors["claude_native", ] * whole_text_vectors["gemini_native", ])
  )
)

human_similarity_differences <- tibble::tibble(
  comparison = c(
    "ChatGPT–Human vs Claude–Human",
    "ChatGPT–Human vs Gemini–Human",
    "Claude–Human vs Gemini–Human"
  ),
  whole_text_similarity_difference = c(
    abs(whole_text_similarities$whole_text_similarity[whole_text_similarities$comparison == "ChatGPT–Human"] -
          whole_text_similarities$whole_text_similarity[whole_text_similarities$comparison == "Claude–Human"]),
    abs(whole_text_similarities$whole_text_similarity[whole_text_similarities$comparison == "ChatGPT–Human"] -
          whole_text_similarities$whole_text_similarity[whole_text_similarities$comparison == "Gemini–Human"]),
    abs(whole_text_similarities$whole_text_similarity[whole_text_similarities$comparison == "Claude–Human"] -
          whole_text_similarities$whole_text_similarity[whole_text_similarities$comparison == "Gemini–Human"])
  )
)

x4_title_subheading_bundle$whole_text_similarities <- whole_text_similarities
x4_title_subheading_bundle$human_similarity_differences <- human_similarity_differences

saveRDS(x4_title_subheading_bundle, x4_output_path)
x4_title_subheading_bundle <- readRDS(x4_output_path)

whole_text_similarities <- x4_title_subheading_bundle$whole_text_similarities
human_similarity_differences <- x4_title_subheading_bundle$human_similarity_differences

print(whole_text_similarities, n = Inf)
print(human_similarity_differences, n = Inf)
message("Whole-text similarities and LM–Human differences frozen and reloaded successfully.")
```
# Inspect existing MODUL 6 objects and saved bundle fields

```{r}
# Inspect existing MODUL 6 objects and saved bundle fields

cat("\n# Current workspace objects\n")
print(sort(ls(envir = .GlobalEnv)))

cat("\n# Existing MODUL 6 RDS files\n")
modul_6_rds_files <- list.files(
  actual_lsa_output_dir,
  pattern = "\\.rds$",
  full.names = TRUE
)
print(basename(modul_6_rds_files))

cat("\n# Top-level fields in each saved RDS object\n")
for (path in modul_6_rds_files) {
  object <- readRDS(path)

  cat("\n", basename(path), "\n", sep = "")
  cat("Class: ", paste(class(object), collapse = ", "), "\n", sep = "")

  if (!is.null(names(object))) {
    print(names(object))
  } else {
    cat("No named top-level fields.\n")
  }
}
```

#

```{r}
# MODUL 6.42 — Assemble and freeze final x5 and x6 indices families

# 1. Load the existing source objects

expected_seed_word_source <- readRDS(file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_x4_expected_seed_word_diagnostics.rds"
))

prompt_induced_ass_net_source <- readRDS(file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_prompt_induced_associative_net.rds"
))

whole_text_source <- readRDS(file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_x4_title_subheading.rds"
))

stopifnot(
  all(c("n_sw_expected", "n_sw_detected", "n_sw_annotated", "n_sw_in_model", "n_sw_oov") %in%
        names(expected_seed_word_source$diagnostics)),
  all(c(
    "chatgpt_human_assoc_net_similarity",
    "claude_human_assoc_net_similarity",
    "chatgpt_claude_assoc_net_similarity",
    "seed_word_net_similarities"
  ) %in% names(prompt_induced_ass_net_source)),
  all(c("whole_text_similarities", "human_similarity_differences") %in%
        names(whole_text_source)),
  exists("whole_text_vectors")
)

# 2. Assemble the final indices families

x5_prompt_compliance_bundle <- list(
  diagnostics = expected_seed_word_source$diagnostics,
  chatgpt_human_assoc_net_similarity =
    prompt_induced_ass_net_source$chatgpt_human_assoc_net_similarity,
  claude_human_assoc_net_similarity =
    prompt_induced_ass_net_source$claude_human_assoc_net_similarity,
  chatgpt_claude_assoc_net_similarity =
    prompt_induced_ass_net_source$chatgpt_claude_assoc_net_similarity,
  seed_word_net_similarities =
    prompt_induced_ass_net_source$seed_word_net_similarities
)

x6_txts_lsa_sim_bundle <- list(
  whole_text_similarities = whole_text_source$whole_text_similarities,
  human_similarity_differences = whole_text_source$human_similarity_differences,
  whole_text_vectors = whole_text_vectors
)

# 3. Freeze and reload

x5_prompt_compliance_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_x5_prompt_compliance_bundle.rds"
)

x6_txts_lsa_sim_path <- file.path(
  actual_lsa_output_dir,
  "modul_6_actual_lsa_x6_txts_lsa_sim_bundle.rds"
)

saveRDS(x5_prompt_compliance_bundle, x5_prompt_compliance_path, version = 3)
saveRDS(x6_txts_lsa_sim_bundle, x6_txts_lsa_sim_path, version = 3)

x5_prompt_compliance_bundle <- readRDS(x5_prompt_compliance_path)
x6_txts_lsa_sim_bundle <- readRDS(x6_txts_lsa_sim_path)

# 4. Validate the frozen families

stopifnot(
  identical(
    names(x5_prompt_compliance_bundle),
    c(
      "diagnostics",
      "chatgpt_human_assoc_net_similarity",
      "claude_human_assoc_net_similarity",
      "chatgpt_claude_assoc_net_similarity",
      "seed_word_net_similarities"
    )
  ),
  identical(
    names(x6_txts_lsa_sim_bundle),
    c(
      "whole_text_similarities",
      "human_similarity_differences",
      "whole_text_vectors"
    )
  )
)

message("x5_prompt_compliance and x6_txts_lsa_sim frozen and reloaded successfully.")
```


#

```{r}
# MODUL 6.43 — Final audit of frozen actual-LSA outputs

# 1. Load the final measure bundles

final_lsa_paths <- c(
  x3s = file.path(actual_lsa_output_dir, "modul_6_actual_lsa_x3s_sentence_continuity.rds"),
  x3p = file.path(actual_lsa_output_dir, "modul_6_actual_lsa_x3p_paragraph_continuity.rds"),
  x3pcw = file.path(actual_lsa_output_dir, "modul_6_actual_lsa_x3pcw_paragraph_continuity.rds"),
  x3pcw_gradient = file.path(actual_lsa_output_dir, "modul_6_actual_lsa_x3pcw_topic_continuity_gradient.rds"),
  x3co = file.path(actual_lsa_output_dir, "modul_6_actual_lsa_x3co_connector_continuity.rds"),
  x4 = file.path(actual_lsa_output_dir, "modul_6_actual_lsa_x4_title_subheading.rds"),
  x5_prompt_compliance = file.path(actual_lsa_output_dir, "modul_6_actual_lsa_x5_prompt_compliance_bundle.rds"),
  x6_txts_lsa_sim = file.path(actual_lsa_output_dir, "modul_6_actual_lsa_x6_txts_lsa_sim_bundle.rds")
)

stopifnot(all(file.exists(final_lsa_paths)))
final_lsa <- lapply(final_lsa_paths, readRDS)

# 2. Validate the frozen structures

stopifnot(
  all(c("level", "definition", "sentence_pairs", "profiles") %in% names(final_lsa$x3s)),
  all(c("level", "definition", "paragraph_pairs", "profiles") %in% names(final_lsa$x3p)),
  all(c("level", "definition", "pos_keep", "paragraph_segments", "projection", "paragraph_pairs", "profiles") %in% names(final_lsa$x3pcw)),
  all(c("definition", "paragraph_pairs", "gradient_profiles", "profiles_extended") %in% names(final_lsa$x3pcw_gradient)),
  all(c("sentence_segments", "projection", "continuity", "summary", "plot_data") %in% names(final_lsa$x3co)),
  all(c("title_pairs", "section_segments", "section_vectors", "section_subheading_pairs", "profiles", "prompt_baseline_pairs", "prompt_baseline") %in% names(final_lsa$x4)),
  all(c("diagnostics", "chatgpt_human_assoc_net_similarity", "claude_human_assoc_net_similarity", "chatgpt_claude_assoc_net_similarity", "seed_word_net_similarities") %in% names(final_lsa$x5_prompt_compliance)),
  identical(names(final_lsa$x6_txts_lsa_sim), c("whole_text_similarities", "human_similarity_differences", "whole_text_vectors"))
)

# 3. Validate central dimensions and retained values

seed_diagnostics <- final_lsa$x5_prompt_compliance$diagnostics
seed_detected <- stats::setNames(seed_diagnostics$n_sw_detected, seed_diagnostics$model)
seed_expected <- stats::setNames(seed_diagnostics$n_sw_expected, seed_diagnostics$model)

stopifnot(
  nrow(final_lsa$x3s$profiles) == 4L,
  nrow(final_lsa$x3p$profiles) == 4L,
  nrow(final_lsa$x3pcw$profiles) == 4L,
  nrow(final_lsa$x3pcw_gradient$gradient_profiles) == 4L,
  nrow(final_lsa$x4$profiles) == 4L,
  nrow(final_lsa$x4$title_pairs) == 125L,
  nrow(final_lsa$x4$prompt_baseline_pairs) == 3L,
  seed_expected[["ChatGPT"]] == 61,
  seed_expected[["Claude"]] == 61,
  seed_detected[["ChatGPT"]] == 48,
  seed_detected[["Claude"]] == 50,
  nrow(final_lsa$x5_prompt_compliance$seed_word_net_similarities) == 3L,
  nrow(final_lsa$x6_txts_lsa_sim$whole_text_similarities) == 6L,
  nrow(final_lsa$x6_txts_lsa_sim$human_similarity_differences) == 3L,
  nrow(final_lsa$x6_txts_lsa_sim$whole_text_vectors) == 4L,
  ncol(final_lsa$x6_txts_lsa_sim$whole_text_vectors) == 50L
)

# 4. Freeze the final audit

final_lsa_audit <- tibble::tibble(
  indices_family = names(final_lsa_paths),
  frozen_file = basename(final_lsa_paths),
  file_size_bytes = unname(file.info(final_lsa_paths)$size),
  status = "PASS"
)

saveRDS(final_lsa_audit, file.path(actual_lsa_output_dir, "modul_6_actual_lsa_final_audit.rds"), version = 3)
readr::write_csv(final_lsa_audit, file.path(actual_lsa_output_dir, "modul_6_actual_lsa_final_audit.csv"))

print(final_lsa_audit, n = Inf)
message("MODUL 6 final frozen-output audit passed.")
```






